% =====================================================================
\section{Introduction}\label{sec:intro}
% =====================================================================

Theories of consciousness divide into mechanism-driven proposals
(Integrated Information Theory~\citep{Tononi2008,BalduzziTononi2008},
Global Workspace Theory~\citep{Baars1988GWT,Dehaene2014ConsciousAndBrain},
predictive processing~\citep{FristonFreeEnergy2010,ClarkPP2013}) and
structure-driven proposals (geometric or topological substrates,
neural-population dynamics). The mechanism-driven proposals offer
compelling axiomatic stories; we are not aware of prior work that
has yielded the kind of preregistered multi-domain quantitative
benchmark against real EEG and connectomics references used here. The structure-driven
proposals produce numbers but often rely on fitted parameters,
learned weights, or domain-specific calibration.

This paper takes a deliberately constrained third path. Once a
substrate is chosen, we ask which neuroscience phenomena it is
consistent with under \emph{no} shape parameter tuning, no learned
weights, no subject-level measurement fitting, and no
neural-data-fitted shape parameters. The substrate is the
600-cell regular 4-polytope $\Rsixhundred$, treated as a graph with
H$_4$ Coxeter symmetry. It has been studied in pure mathematics for
over a century~\citep{CoxeterRegularPolytopes,Weisstein600Cell}; to
our knowledge it has not been proposed before as an empirical
candidate substrate for consciousness-linked signatures. We construct $\Rsixhundred$, fix its response
operator $\Cph = \Lop + \Ph^{-2} I$ where $\Ph=(1+\sqrt 5)/2$, add a
single condition-dependent self-injection coupling $\eta$ and a
single graph-pinned nonlinearity, and test the resulting witness
against eighteen preregistered correspondences plus six companion
drug/sleep EEG signatures.

\subsection*{What this paper claims}

We claim a single \emph{substrate witness}: that a geometry-fixed
substrate, with no shape parameters tuned to any neural dataset, is
consistent with eighteen preregistered correspondences (frozen
2026-04-18) and six companion drug/sleep EEG signatures of
conscious vs unconscious states.

\begin{enumerate}\itemsep=2pt
\item \textbf{Substrate is fixed once chosen.} Once $\Rsixhundred$
  is selected, the vertex set ($|V|=120$, all on the unit
  $3$-sphere) is forced by the canonical 600-cell construction; H$_4$
  transitivity forces uniform vertex degree (here $12$ on the
  short-edge nearest-neighbour graph); and the Laplacian spectrum is
  computed from the resulting graph and reported as observed, with
  multiplicities matching the expected H$_4$ block sizes
  (\S\ref{sec:substrate}). The response operator
  $\Cph = \Lop + \Ph^{-2} I$ is fully fixed once the graph is
  constructed and the stability shift $\Ph^{-2}$ is chosen as a
  design-level clamp.
\item \textbf{Cortical avalanches.} Wake cascade-event power-law
  exponent $\alpha = 2.252$, $95\%$ CI $[1.82, 2.86]$ ($R^{2}=0.956$),
  three-way overlapping the Sleep-EDFx EEG CI $[2.50, 2.53]$
  (n$=$30 subjects) and ARIA's prior cascade-pipeline CI
  $[2.73, 3.25]$.
\item \textbf{Six drug/sleep signatures.} On a single deterministic
  trajectory at seed $42$: NREM-N3 phenomenal-intensity variance
  collapse to $0.463\!\times$ wake; propofol modality-switching
  $1.83\!\times$ wake; propofol continuity drop $+0.066$; propofol
  $\Phi$ collapse to $0.33\!\times$ wake (IIT direction confirmed);
  recovery deterministically identical to wake; wake cascade-$\alpha$
  in the SOC band.
\item \textbf{Eighteen preregistered correspondences pass.}
  $17/18$ at standard methodology; $18/18$ after a documented
  $N\!=\!20$ deep-dive on the residual high-variance interaction
  test; \emph{no preregistered threshold has been modified}.
\item \textbf{Cross-domain selectivity.} The substrate exhibits
  selective amplification in the two cross-domain tasks tested
  (chess $+40.6$pp leave-one-out lift at depth $n\!=\!25$ ticks;
  conversation $-4.4$pp lift, within preregistered neutrality bounds)
  and serves as an H$_4$-transitive deterministic null reference for
  cortical functional connectivity (HCP full-cohort descriptive
  $n\!=\!1003$: ARIA at $-11.58\sigma$ on degree homogeneity;
  $+79.78\sigma$ on raw participation ratio with the node-count caveat
  of \S\ref{ssec:hcp}).
\end{enumerate}

\subsection*{What this paper does \emph{not} claim}

\begin{itemize}\itemsep=2pt
\item \emph{Not a uniqueness claim.} We do not claim the 600-cell is
  the unique substrate consistent with these signatures. Other regular
  4-polytopes (the 24-cell, the 120-cell) are an explicit ablation
  build, not a discharged comparison. The 600-cell choice is post-hoc
  motivated by the H$_4$ Coxeter cascade structure and biological
  observables; it is not an a-priori derivation from first principles.
\item \emph{Not a derivation of consciousness.} The substrate witness
  shows quantitative agreement with cortical signatures; it does not
  establish that the substrate \emph{is} consciousness, nor that
  its dynamics implement specific phenomenal content.
\item \emph{Not a selection theorem.} The companion adaptive-closure-
  transport preprint~\citep{SmartAdaptiveClosureTransport2026}
  proposes a 4-tuple bridge $(M, L_M, W, R_{\mathrm{hom}})$ in which
  this substrate fills the $L_M$ slot. The selection of the 600-cell
  as the active $M$ is conjectural in that paper and is treated as
  non-load-bearing here. We do not deliver a Lyapunov function on the
  reduced flow, nor a $2I$-equivariance audit of the closure operator,
  nor a formal edge-space decomposition. These are listed as open
  builds in~\S\ref{sec:limitations}.
\item \emph{Not a circuit-level model.} The substrate is at the
  architectural-algorithmic level. We do not identify which neural
  populations implement context rotation or partial emission, only
  that some such mechanisms appear in the substrate's preregistered
  ablation matrix and exhibit strong inter-mechanism coupling.
\item \emph{Not a derivation of the $\Ph^{-2}$ floor.} The shifted-
  Laplacian floor $\Ph^{-2} \approx 0.382$ is a design-level
  stability clamp (it makes $\Cph$ strictly positive definite and
  bounds the Green response). It is not derived as a theorem from a
  closure functional. The companion kernel
  document~\citep{SmartAriaClosureKernel2026} discusses its role.
\end{itemize}

\subsection*{Mapping from numerical results to admissible claims}

To keep this paper inside the substrate-witness scope, we use the
following claim-boundary discipline. Numerical results
$\mathcal R_{\mathrm{numeric}}$ are mapped to admissible claims
$\mathcal C_{\mathrm{admissible}}$ by the rule
\[
q\colon \mathcal R_{\mathrm{numeric}} \longrightarrow \mathcal C_{\mathrm{admissible}},
\qquad
\mathcal C_{\mathrm{admissible}}
\;=\;\{\text{`consistent with', `inside threshold', `direction confirmed'}\}.
\]
We never write `the substrate \emph{is} cortex' or `derives consciousness'.
A result that lands inside its preregistered threshold licenses a
`consistent with' claim. A result that exceeds the preregistered
threshold by an order of magnitude (e.g.\ chess $+40.6$pp vs the
$+15$pp floor) licenses `decisively above prereg', not `proves'. A
$\sigma$-distance result against an external null
(e.g.\ $-11.58\sigma$ on HCP degree homogeneity) licenses `outside
the biological distribution', not `cortex has drifted from an ideal
polytope'. The claim-boundary rule is summarised in the box below
and applied throughout~\S\ref{sec:results}.

\medskip
\begin{center}
\fbox{\parbox{0.92\linewidth}{\small
\textbf{What is tested / what is not claimed.}\par
\emph{Tested:} 18 preregistered correspondences plus 6 drug/sleep EEG
signatures, on a geometry-fixed substrate with one condition-dependent
parameter $\eta$ and one graph-pinned nonlinearity, against published
biological observables.\par
\emph{Not claimed:} substrate uniqueness; derivation of consciousness;
selection theorem on the 4-tuple bridge; circuit-level mechanistic
identification; first-principles derivation of $\Ph^{-2}$ shift;
that cortex \emph{is} the 600-cell.
}}
\end{center}

\subsection*{Layout}

\S\ref{sec:method} gives the provenance ledger (preregistration date,
seeds, scripts, datasets, thresholds). \S\ref{sec:substrate}
constructs $\Rsixhundred$ and the response operator $\Cph$, with the
$\Ph^{-2}$ shift disclosed as a design-level stability clamp.
\S\ref{sec:chain} adds the recurrent self-model layer above the
substrate (single nonlinearity, single self-injection coupling).
\S\ref{sec:results} reports the empirical tables: six drug/sleep
signatures, eighteen preregistered correspondences, three-way
$\alpha$-overlap. \S\ref{sec:stress} presents the C$\times$P
synergy stress test ($N\!=\!3, 5, 10, 20$ trend with bootstrap
$95\%$ CI). \S\ref{sec:cross_domain} reports cross-domain
selectivity (chess, conversation, HCP). \S\ref{sec:discussion}
discusses the substrate witness and proposes a non-load-bearing
ACT bridge (without claiming a selection theorem).
\S\ref{sec:limitations} enumerates limitations and the
hostile-review guard matrix. \S\ref{sec:conclusion} concludes.
