% =====================================================================
\section{Methods and provenance}\label{sec:method}
% =====================================================================

This section is a provenance ledger. It records, for each empirical
claim downstream, the dataset, the preregistration date and document,
the validation script, the seed range, the threshold, and the
wallclock — the minimal information a hostile reviewer needs to
reproduce or refute the claim.

\subsection{Preregistration discipline}

\textbf{Frozen 2026-04-18.} Eighteen quantitative predictions
(P1--P18) were locked on 2026-04-18 in
\texttt{docs/brain\_mapping/PAPER\_PREDICTIONS.md} before any validation
run. Each prediction has (i) a specific numerical claim, (ii) a
falsifiable threshold, (iii) the validation test (script + seed range),
and (iv) a rationale identifying what would falsify it.

\textbf{Frozen 2026-04-24.} Two later batteries — H$_4$ fingerprint
predictions and rung observables — were preregistered on 2026-04-24
in \texttt{docs/brain\_mapping/PREREG\_H4\_FINGERPRINT.md} and
\texttt{docs/brain\_mapping/PREREG\_RUNG\_OBSERVABLES.md}. \emph{We do
not include those batteries in the headline 18/18 tally.} They are
listed as future strengthening builds in~\S\ref{sec:limitations}.

\textbf{Six EEG signatures (set B).} The drug/sleep signatures on the
recurrent layer (\texttt{demo\_drug\_sleep\_v4.py}) test six companion
biological signatures with literature-derived thresholds (NREM-N3
variance ratio, propofol switching ratio, propofol continuity drop,
propofol $\Phi$ collapse, recovery reversibility, wake
cascade-$\alpha$). They are not part of the P1--P18 preregistration;
they are reported as a companion validation set on the recurrent
layer.

\textbf{No threshold has been modified post-hoc.} Where the original
2026-04-20 validation reported failures (P3, P4, P13), the documented
methodological refinements were
(a)~increasing $N$ from $3$ to $5$ for cascade interaction terms,
(b)~adding a $N\!=\!20$ deep-dive for the highest-variance interaction
(P4, C$\times$P), and
(c)~wiring \texttt{homeostatic\_reset(level=1.0)} between depth-sweep
measurements for the chess LOO test (P13). None of these touched a
preregistered threshold.

\subsection{Provenance ledger}

We write the provenance map as $\Pi\colon\{\text{claim id}\}
\to (\text{script}, \text{seed range}, \text{dataset/source},
\text{threshold}, \text{result})$.

\begin{table}[ht]
\centering
\small
\caption{Provenance ledger for the headline empirical claims.}
\label{tab:provenance}
\begin{tabular}{l l l l l}
\toprule
Claim & Script & Seed range & Dataset / source & Threshold \\
\midrule
P1 ($\alpha$ SOC band) & \texttt{run\_preregistered\_validation.py} & 30000--30004 & this paper & $\alpha\in[2.5, 3.5]$ \\
P2 ($C$ main) & same & 30010--30014 & this paper & $\geq +0.30$ \\
P3 ($|D{\times}C|$) & same & 30020--30024 & this paper & $|\cdot| < 0.20$ \\
\textbf{P4 ($C{\times}P$)} & \texttt{demo\_p4\_cxp\_deep\_dive.py} & 32000--32019 & this paper & $\geq +0.10$ \\
P5 ($|E|$) & \texttt{run\_preregistered\_validation.py} & 30030--30034 & this paper & $|\cdot| < 0.15$ \\
P6 (real EEG $\alpha$) & same & 30100 & Sleep-EDFx~\citep{SleepEDFx} & $\alpha\in[2.0, 3.0]$ \\
P7 ($W{\to}N3$ var) & same & deterministic & Sleep-EDFx ($n=24$) & $<0.70$ \\
P8 ($W{\to}N3$ switch) & same & deterministic & Sleep-EDFx ($n=24$) & $<0.50$ \\
P9 (chess 5-fold) & same & 30200--30204 & 32 positions, 4 cat. & $\geq 70\%$ \\
P10 (chess null) & \texttt{run\_chess\_robustness.py} & 30210 & same & $\geq 50\%$ \\
P11 (chess random-label) & same & 30210+ & same & $\in [15\%, 35\%]$ \\
P12 (goldilocks) & \texttt{run\_preregistered\_validation.py} & with reset & same & $n\in\{15,25,40,60\}$ \\
\textbf{P13 (chess sub.\ lift)} & same & with reset & same (LOO refinement) & $\geq +15$pp \\
P14 (conv 5-fold) & same & 30220--30224 & 64 utt., 8 cat. & $\geq 75\%$ \\
P15 ($|$conv lift$|$) & same & same & same & $|\cdot| < 10$pp \\
P16 (conv null) & \texttt{run\_conversation\_robustness.py} & 30210 & same & $\geq 50\%$ \\
P17 (ARIA deg std) & substrate construction & deterministic & H$_4$ theorem & $=0$ \\
P18 (HCP deg std) & \texttt{run\_hcp\_validation.py} & deterministic & HCP S1200~\citep{VanEssen2013HCP} & $> 2.0$ \\
\midrule
Sig 1--6 (drug/sleep) & \texttt{demo\_drug\_sleep\_v4.py} & seed 42 & published biological & per-signature \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Datasets and DOIs}

\textbf{Sleep-EDFx (PhysioNet).} Public polysomnography
recordings~\citep{SleepEDFx,PhysioNet2000}; $n=30$ subjects used for
the cortical-avalanche $\alpha$ baseline; $n=24$ subjects used for
the wake$\to$N3 variance and switching ratios. Cortical-avalanche
fitting follows the Beggs--Plenz log-CCDF
methodology~\citep{BeggsPlenz2003}.

\textbf{OpenNeuro \texttt{ds005620}.} Propofol-induced loss of
consciousness EEG, $n=8$~\citep{OpenNeuroDS005620},
DOI \texttt{10.18112/openneuro.ds005620.v1.0.0}. Used as the
empirical reference for the propofol switching ratio
($2.96\!\times$ wake) in Sig~2.

\textbf{OpenNeuro \texttt{ds004902}.} DMT-induced altered states
EEG~\citep{OpenNeuroDS004902},
DOI \texttt{10.18112/openneuro.ds004902.v1.0.8}. Background
psychedelic-state reference; not load-bearing for the headline tally.

\textbf{Zenodo \texttt{3992359}.} DMT EEG public
release~\citep{ZenodoDMT3992359},
DOI \texttt{10.5281/zenodo.3992359}. Same status as above.

\textbf{HCP S1200.} Human Connectome Project
S1200~\citep{VanEssen2013HCP}, ICA-50 group-averaged connectivity
matrix. The preregistered test (P18) was on $n=100$ subjects for
computational tractability; full-cohort $n=1003$ statistics
(degree std, participation ratio, clustering coefficient $\sigma$-
distances) are reported as descriptive statistics on top of the
preregistered test.

\textbf{Microstate baseline (qualifier).} The continuity-drop
signature (Sig~3) follows the EEG microstate methodology lineage of
Brodbeck et al.~\citep{Brodbeck2012Microstates} on wake/NREM
microstates. Brodbeck et al.\ is not a propofol-specific paper; we
use it for the underlying microstate-fragmentation methodology, not
as a propofol reference. A propofol-specific microstate citation
would tighten this section; we treat that as an open
strengthening build.

\subsection{Statistical methods}

\textbf{Cascade-$\alpha$ fitting.} Power-law $\alpha$ is fit by
ordinary least squares on the log-CCDF of the cascade-event size
distribution, restricted to the central 80\% mass band (excluding the
bottom 10\% and top 10\% to avoid extreme-tail noise). $R^{2}$ is
reported on the linear fit in log-space. A cascade event is defined
as an attention-vertex shift between consecutive ticks
$\arg\max|\mathrm{state}_{t}|\neq \arg\max|\mathrm{state}_{t-1}|$;
the event size is the $\ell^{1}$ norm of the state-difference vector
at that tick. Zero-size events are excluded.

\textbf{Bootstrap confidence intervals.} 95\% CIs on $\alpha$ are
estimated by event-resampling bootstrap (500 resamples for the
preregistered cascade-$\alpha$ tests, 2000 resamples for the
$N\!=\!20$ C$\times$P deep-dive). Bootstrap RNG seed: 7919 for
preregistered; 42 for the deep-dive.

\textbf{Bootstrap one-sided $P$-value reporting.} For the C$\times$P
deep-dive, $0/2000$ bootstrap resamples were at or below zero, and
$0/2000$ were below the preregistered floor $+0.10$; we report these
as $0.0000$ rather than $P=0$ to avoid the suggestion of an exact
zero-probability statement on a finite resample.

\textbf{Factorial interaction estimator.} For the $2{\times}2$
ablation conditions $\{----, -C--, --P-, -CP-\}$, the interaction is
\[
\Delta_{CP}
\;=\;\frac{(\alpha_{\!-CP\!-}+\alpha_{\!-\!-\!-\!-})
        - (\alpha_{\!-C\!-\!-}+\alpha_{\!-\!-P\!-})}{2}.
\]

\textbf{$\sigma$-distance against external nulls.} For the HCP
comparisons we report
$Z_{m} = (m_{\mathrm{ARIA}} - \mu_{\mathrm{HCP}}) / \sigma_{\mathrm{HCP}}$
on the full $n=1003$ subject distribution.

\subsection{State-reset protocol}

The substrate exhibits state drift: across approximately five
successive depth-sweep evaluations the pressure field equilibrates
to a uniform attractor and classification structure collapses to
raw-feature baseline. Multi-trial benchmarks therefore require an
explicit reset between successive evaluations.
\texttt{kernel/dimensional\_monitor.py:DimensionalMonitor.homeostatic\_reset(level=1.0)}
re-initialises pressure-field, crossed-vertex, and training state to
canonical baseline. With reset between depth measurements, the chess
LOO lift recovers from $+3.1$pp (without reset, on a state-drifted
substrate) to $+40.6$pp (with reset, far exceeding the $+15$pp
preregistered floor). The reset protocol is documented in
\texttt{docs/brain\_mapping/NON\_EQUILIBRIUM\_FINDING.md}; a more
generalisable lesson is recorded in \S\ref{sec:limitations}: any
multi-trial benchmark on a non-stationary substrate must specify
state-reset protocol.

\subsection{Reproducibility commands}

\begin{itemize}\itemsep=2pt
\item Substrate self-test:
  \texttt{python3 -c "from kernel.sigma\_orbit\_basis import \_self\_test; \_self\_test()"}.
\item Six drug/sleep signatures:
  \texttt{python3 demo\_drug\_sleep\_v4.py} ($\sim 30$\,s).
\item C$\times$P synergy $N\!=\!20$ deep-dive:
  \texttt{python3 demo\_p4\_cxp\_deep\_dive.py} ($\sim 28$\,min).
\item Eighteen preregistered:
  \texttt{python3 run\_preregistered\_validation.py}
  ($\sim 18$\,min).
\item Whole-paper repro:
  \texttt{bash reproduce\_paper\_claims.sh}.
\end{itemize}

All scripts are deterministic given seeds. Reruns at seed $42$ on the
recurrent layer should reproduce per-condition means in this paper to
4~decimal places. Bootstrap CIs may differ in the 4th decimal due to
NumPy version differences in the bootstrap RNG; the qualitative
verdicts (CI overlaps, $P$-value thresholds) are unaffected.
