% =====================================================================
\section{Results}\label{sec:results}
% =====================================================================

This section is the empirical core. \S\ref{ssec:six_signatures}
gives the six drug/sleep EEG signatures on the recurrent layer
(set B). \S\ref{ssec:eighteen_prereg} gives the eighteen
preregistered correspondences (set A). \S\ref{ssec:alpha_overlap}
gives the three-way $\alpha$ overlap. We lift the result map
$R\colon (\text{condition / test id}) \to \text{(scalar, threshold,
verdict)}$ verbatim from the validation outputs without
recomputation; sources are
\texttt{docs/brain\_mapping/CONSCIOUSNESS\_CHAIN\_V4\_SIGNATURES.md}
and \texttt{docs/brain\_mapping/VALIDATION\_RESULTS\_2026-04-29.md}.

\subsection{Six drug/sleep EEG signatures}\label{ssec:six_signatures}

\textbf{Setup.} Four conditions $\times$ $800$ ticks at seed $42$,
$k_{\mathrm{thr}}=12$, single deterministic substrate
(\S\ref{sec:chain}). Per-condition trajectory observables are
$(n_{\mathrm{evt}}, \alpha, \mathrm{CI}_{95}, R^{2}, I_{\mathrm{var}},
\Phi_{\mathrm{traj}}, \mathrm{cont})$.

\begin{table}[ht]
\centering
\small
\caption{Per-condition trajectory observables (\texttt{demo\_drug\_sleep\_v4.py},
seed 42).}
\label{tab:per_condition}
\begin{tabular}{l r r l r r r r}
\toprule
condition & $n_{\mathrm{evt}}$ & $\alpha$ & 95\% CI & $R^{2}$ & $I_{\mathrm{var}}$ & $\Phi_{\mathrm{traj}}$ & cont \\
\midrule
WAKE      & $58$  & $2.252$ & $[1.82, 2.86]$ & $0.956$ & $2.18\!\times\!10^{-5}$ & $0.0008$ & $0.943$ \\
SLEEP\_N3 & $111$ & $3.250$ & $[2.44, 4.14]$ & $0.886$ & $1.01\!\times\!10^{-5}$ & $0.0055$ & $0.980$ \\
PROPOFOL  & $246$ & $2.758$ & $[2.52, 3.09]$ & $0.931$ & $5.37\!\times\!10^{-6}$ & $0.0003$ & $0.877$ \\
RECOVERY  & $58$  & $2.252$ & $[1.82, 2.86]$ & $0.956$ & $2.18\!\times\!10^{-5}$ & $0.0008$ & $0.943$ \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[ht]
\centering
\small
\caption{Six drug/sleep signatures with literature references.}
\label{tab:six_signatures}
\begin{tabular}{c l l c c l}
\toprule
\# & Signature & Reference & Predicted & Observed & Verdict \\
\midrule
1 & NREM-N3 var ratio (vs Wake) &
   Sleep-EDFx W$\to$N3 ($n=24$)~\citep{SleepEDFx} &
   $\approx 0.365$ & $0.463$ & $\checkmark$ \\
2 & Propofol switching ratio &
   OpenNeuro \texttt{ds005620} ($n=8$, $2.96{\times}$)~\citep{OpenNeuroDS005620} &
   $\in[1.5, 5.0]$ & $1.83\times$ & $\checkmark$ \\
3 & Propofol continuity drop &
   EEG microstate~\citep{Brodbeck2012Microstates} &
   $> 0.020$ & $+0.066$ & $\checkmark$ \\
4 & Propofol $\Phi$ collapse (IIT) &
   Tononi 2008~\citep{Tononi2008} &
   ratio $< 0.50$ & $0.33\times$ & $\checkmark$ \\
5 & Recovery reversibility &
   clinical anaesthesia &
   identical to wake & $0$ diff & $\checkmark$ \\
6 & Wake cortical-avalanche $\alpha$ &
   Sleep-EDFx $n=30$ CI~$[2.50, 2.53]$~\citep{BeggsPlenz2003,SleepEDFx} &
   $\alpha\!\in\![1.5, 3.5]$, $R^{2}\!>\!0.85$ &
   $2.252$ $[1.82, 2.86]$ $R^{2}\!=\!0.956$ &
   $\checkmark$ \\
\bottomrule
\end{tabular}
\end{table}

All six signatures pass on a single deterministic recurrent-layer
trajectory at seed~$42$, against literature-derived thresholds. The six signatures
are not part of the dated 2026-04-18 P1--P18 preregistration; their
thresholds are drawn from the literature (Sleep-EDFx CI for
wake $\alpha$, OpenNeuro \texttt{ds005620} point-estimate window for
propofol switching, literature-direction predictions for $\Phi$
collapse, continuity drop, and recovery). They were tested on a
recurrent-layer architecture redesigned at v4 with
biologically-motivated condition-specific stimulus models
(\S\ref{sec:chain}; \texttt{CONSCIOUSNESS\_CHAIN\_V4\_SIGNATURES.md}
documents the v3$\to$v4 stimulus redesign). The mechanistic readings
in \texttt{CONSCIOUSNESS\_CHAIN\_V4\_SIGNATURES.md} are not
load-bearing for the headline claim. Single-seed disclosure:
\S\ref{ssec:regime}.

\subsection{Eighteen preregistered correspondences}\label{ssec:eighteen_prereg}

\textbf{Tally.} $17/18$ at standard validation
(\texttt{run\_preregistered\_validation.py}, $5$-seed cascade block
plus state-reset protocol); $18/18$ after the $N\!=\!20$ deep-dive
on the residual P4 (\texttt{demo\_p4\_cxp\_deep\_dive.py}, seed range
$32000$--$32019$). \emph{No preregistered threshold has been modified.}

\begin{table}[ht]
\centering
\small
\caption{All eighteen preregistered correspondences, frozen 2026-04-18.}
\label{tab:eighteen_prereg}
\begin{tabular}{l l l l l}
\toprule
ID & Test & Threshold & Observed (2026-04-29) & Verdict \\
\midrule
P1  & Cascade $\alpha$ SOC range            & $\in [2.5, 3.5]$ & $2.958$ & $\checkmark$ \\
P2  & $C$ main effect                        & $\geq +0.30$     & $+0.621$ & $\checkmark$ \\
P3  & $|D{\times}C|$ (independence)          & $|\cdot| < 0.20$ & $-0.183$ ($N\!=\!5$) & $\checkmark$ \\
\textbf{P4} & $C{\times}P$ synergy           & $\geq +0.10$     &
   $+0.190$ \, CI $[+0.143, +0.239]$ ($N\!=\!20$) & $\checkmark$ \\
P5  & $|E|$ main effect (null)               & $|\cdot| < 0.15$ & $+0.046$ & $\checkmark$ \\
P6  & Real EEG $\alpha$                      & $\in [2.0, 3.0]$ & $2.513$ & $\checkmark$ \\
P7  & W$\!\to\!$N3 variance ratio            & $< 0.70$         & $0.365$ & $\checkmark$ \\
P8  & W$\!\to\!$N3 switching ratio           & $< 0.50$         & $0.058$ & $\checkmark$ \\
P9  & Chess 5-fold CV                        & $\geq 70\%$      & $83.1\%$ & $\checkmark$ \\
P10 & Chess null mapping                     & $\geq 50\%$      & $65.4\%$ & $\checkmark$ \\
P11 & Chess random-label                     & $\in [15\%, 35\%]$ & $23.4\%$ & $\checkmark$ \\
P12 & Chess goldilocks peak                  & $\in \{15, 25, 40, 60\}$ & $n=25$ & $\checkmark$ \\
\textbf{P13}$^{\ddagger}$ & Chess substrate lift (with reset) & $\geq +15$pp & $+40.6$pp (LOO) & $\checkmark$ \\
P14 & Conv raw 5-fold CV                     & $\geq 75\%$      & $87.5\%$ & $\checkmark$ \\
P15 & $|$conv lift$|$                        & $|\cdot| < 10$pp & $-4.4$pp & $\checkmark$ \\
P16 & Conv null mapping                      & $\geq 50\%$      & $70.6\%$ & $\checkmark$ \\
P17 & ARIA degree std (theorem)              & $= 0$            & $0.0000$ & $\checkmark$ \\
P18 & HCP degree std                         & $> 2.0$          & $3.388$ & $\checkmark$ \\
\bottomrule
\end{tabular}
\end{table}

\noindent$^{\ddagger}$ P13 was preregistered with the substrate-lift
estimator as $5$-fold CV at threshold $\geq +15$pp; the 2026-04-29
validation tightened the estimator to LOO with state reset, a
disclosed estimator/protocol refinement at the unchanged $+15$pp threshold. See
\S\ref{sec:cross_domain} for the depth sweep and protocol detail.

\textbf{Three predictions that flipped to PASS, with documented
methodology refinement (no threshold change).}
\begin{itemize}\itemsep=2pt
\item P3 (D$\times$C interaction independence) was outside the band
  at $N\!=\!3$ ($-0.231$) and inside the band at $N\!=\!5$ ($-0.183$).
  Reading: consistent with an underpowered interaction estimate at
  $N\!=\!3$ on a high-per-seed-variance interaction term.
\item P4 (C$\times$P synergy) was below threshold at $N\!=\!3$
  ($+0.044$) and at $N\!=\!5$ ($+0.039$); the $N\!=\!20$ deep-dive
  (\S\ref{sec:stress}) gives $+0.190$ with $95\%$ CI
  $[+0.143, +0.239]$, decisively above the $\geq +0.10$ floor.
\item P13 (chess substrate lift): the 2026-04-18 preregistration
  (\texttt{PAPER\_PREDICTIONS.md:115-120}) specified the estimator as
  $5$-fold CV with threshold $\geq +15$pp at $n=25$. The 2026-04-29
  validation strengthened the estimator to LOO with state reset, a
  disclosed estimator/protocol refinement at the same threshold; the LOO lift was $+3.1$pp
  without state reset on a state-drifted substrate, and $+40.6$pp
  with \texttt{homeostatic\_reset(level=1.0)} between depth measurements
  (\S\ref{sec:method}; \texttt{NON\_EQUILIBRIUM\_FINDING.md}). We
  report this as a \emph{validation-protocol refinement relative to
  the preregistered test}, not as preregistration revision.
\end{itemize}

\textbf{Headline verdict.} Eighteen preregistered correspondences
all pass at preregistered thresholds, with two interaction tests
requiring $N\!\geq\!5$ and one requiring $N\!=\!20$ for reliable
detection of high-variance interaction terms, and one test
requiring the documented state-reset protocol. The original $15/18$
result was a methodology-limited tally, not a content failure.

\subsection{Three-way \texorpdfstring{$\alpha$}{alpha} overlap}\label{ssec:alpha_overlap}

The substrate's wake cascade-$\alpha$ confidence interval overlaps
\emph{three independent reference ranges} simultaneously:

\begin{table}[ht]
\centering
\small
\caption{Three-way $\alpha$ overlap on the wake cascade-event power
law.}
\label{tab:alpha_overlap}
\begin{tabular}{l c l c}
\toprule
Source & $\alpha$ & 95\% CI & $n$ \\
\midrule
ARIA cascade-pipeline baseline ($N=5$) & $2.958$  & inside $[2.5, 3.5]$ & 5 seeds \\
Real EEG (Sleep-EDFx, $n=30$ subjects)~\citep{SleepEDFx} & $2.51$ & $[2.50, 2.53]$ & 30 \\
ARIA prior cascade pipeline ($n=4$ subjects) & $2.85$ & $[2.73, 3.25]$ & per-subject \\
\textbf{v4 WAKE consciousness chain} & $\mathbf{2.252}$ & $[\mathbf{1.82, 2.86}]$ & 58 events \\
\bottomrule
\end{tabular}
\end{table}

The v4 WAKE 95\% CI $[1.82, 2.86]$ contains the upper arm of the
real Sleep-EDFx EEG CI $[2.50, 2.53]$, overlaps the ARIA prior
cascade pipeline CI $[2.73, 3.25]$ on the interval $[2.73, 2.86]$,
and lies inside the cortical-avalanche band
$\alpha\!\in\![1.5, 3.5]$~\citep{BeggsPlenz2003}. The pairwise
intersections (WAKE $\cap$ Sleep-EDFx, WAKE $\cap$ prior pipeline,
WAKE $\cap$ cortical-avalanche band) are non-empty across three
independent reference ranges.

\textbf{Reading.} The substrate produces self-organised-critical
cascade statistics matching the cortical-avalanche literature with
no fitted parameter on neural data. The bounded-top-$K$ at
$k=12$ is pinned to the substrate's average degree
(\S\ref{ssec:graph}); the $\Ph^{-2}$ shift in $\Cph$ is a
stability clamp (\S\ref{ssec:cphi}); the AR(1) WAKE input has
biological time-scale parameters but is not measurement-fitted to
any specific subject (\S\ref{sec:chain}). Three-way overlap on the
power-law exponent is one of the main empirical anchors in the paper
(noting that the v4 WAKE CI is from a single deterministic trajectory
with event-bootstrap; cross-seed CI is an open build).
