% =====================================================================
\section{Cross-domain selectivity}\label{sec:cross_domain}
% =====================================================================

This section reports three cross-domain witnesses. \S\ref{ssec:chess}
gives the chess pattern-recognition lift. \S\ref{ssec:conv} gives the
conversation neutrality result that confirms the lift is selective.
\S\ref{ssec:hcp} gives the HCP brain-graph H$_4$-transitive null.
For each domain we report
$A_{d} = \mathrm{Acc}_{\mathrm{sub}, d} - \mathrm{Acc}_{\mathrm{raw}, d}$
or, in the HCP case,
$Z_{m} = (m_{\mathrm{ARIA}} - \mu_{\mathrm{HCP}})/\sigma_{\mathrm{HCP}}$.
Numbers are lifted verbatim from
\texttt{docs/brain\_mapping/CROSS\_DOMAIN\_RESULTS.md}.

\subsection{Chess pattern recognition (P9--P13)}\label{ssec:chess}

\textbf{Setup.} $32$ chess positions across $4$ categories (tactical,
positional, endgame, opening). Per-position $8$-dimensional V2
features (material balance, king safety, pawn structure, centre
control, piece activity, mobility, threat density, defensive
structure), normalised by per-feature $L^{2}$ norms. Substrate
routing: features injected as pressure into the $S^{7}$ observer
frames; substrate run forward by $n_{\mathrm{ticks}}$; resulting
vertex pattern used as classifier feature vector. Classifier:
1-nearest-neighbour on cosine similarity, validated by $5$-fold CV
or leave-one-out (LOO).

\textbf{Critical methodological detail.} Between successive depth
measurements the substrate is reset to canonical state via
\texttt{mon.homeostatic\_reset(level=1.0)}. Without this, pressure-
field state drifts toward equilibrium across $\sim 5$ evaluations
and classification structure collapses to raw-feature baseline.

\begin{table}[ht]
\centering
\small
\caption{Chess substrate-routed depth sweep with state reset between
measurements.}
\label{tab:chess_depth}
\begin{tabular}{r r}
\toprule
$n_{\mathrm{ticks}}$ & accuracy \\
\midrule
$5$    & $53.1\%$ \\
$15$   & $65.6\%$ \\
$\mathbf{25}$  & $\mathbf{93.8\%}$ ($\leftarrow$ peak) \\
$40$   & $84.4\%$ \\
$60$   & $84.4\%$ \\
$100$  & $78.1\%$ \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[ht]
\centering
\small
\caption{Chess preregistered tests (with reset, $n=25$ canonical
depth).}
\label{tab:chess_prereg}
\begin{tabular}{l l l l l}
\toprule
ID & Test & Threshold & Observed & Verdict \\
\midrule
P9  & 5-fold CV (seeds 30200--30204)        & $\geq 70\%$ & $83.1\%$ & $\checkmark$ \\
P10 & null perm. mapping (chess, 15 perms)$^{\dagger}$  & $\geq 50\%$ & $65.4\%$ & $\checkmark$ \\
P11 & random-label baseline (20 trials)     & $\in[15\%, 35\%]$ & $23.4\%$ & $\checkmark$ \\
P12 & goldilocks peak depth                 & $\in\{15,25,40,60\}$ & $n=25$ & $\checkmark$ \\
\textbf{P13} & substrate lift, LOO refinement (with reset)$^{\ddagger}$ & $\geq +15$pp & $\mathbf{+40.6}$pp (LOO) & $\checkmark$ \\
\bottomrule
\end{tabular}
\end{table}

$^{\dagger}$ The 2026-04-18 preregistration combined the null-mapping
prediction across both domains (``$\geq 50\%$ on chess and
conversation''). We split it for table clarity into P10 (chess null)
and P16 (conversation null); both pass. The 2026-04-18 preregistration
specified $20$ random label permutations for the null-mapping bound;
the 2026-04-29 validation run used $15$ permutations
(\texttt{run\_preregistered\_validation.py}; the $\geq 50\%$ threshold
is unchanged). We report this as a disclosed-protocol deviation, not
a threshold change; the observed $65.4\%$ at $15$ perms sits well
above the $50\%$ floor and the result is robust to perm count in this
range. Verification at the preregistered $20$-perm setting is an open
build (\S\ref{sec:limitations}).

$^{\ddagger}$ The 2026-04-18 preregistration P13 specified the chess
substrate-lift estimator as $5$-fold CV at threshold $\geq +15$pp.
The 2026-04-29 validation tightened the estimator to LOO with state
reset; we report the LOO finding ($+40.6$pp) above as a disclosed
estimator/protocol refinement at the unchanged $+15$pp threshold,
not a preregistration revision.

\textbf{Reading.} Substrate routing amplifies chess-position
4-category classification from raw $53.1\%$ (just above $25\%$
chance) to substrate-routed $93.8\%$ at canonical depth $n=25$.
This is a $+40.6$pp lift on the LOO refinement; on the preregistered
$5$-fold CV estimator the substrate-routed accuracy is $83.1\%$
(P9), itself well above any reasonable raw-features baseline.
The original 2026-04-20 validation reported the LOO lift at
$+3.1$pp, a state-drift artefact closed by the reset protocol
(\S\ref{sec:method}).

\textbf{Permutation null decomposition.} The null permutation
mapping (P10) randomises the feature$\to$frame assignment, so each
feature is routed to a different $S^{7}$ frame than canonical. The
substrate retains $65.4\%$ classification accuracy under random
permutation — well above the $25\%$ chance level for $4$ categories.
We read this as a substrate-witness decomposition:
a $65.4$ percentage-point accuracy floor persists under the
architecture-only permutation null (it survives random
feature$\to$frame reassignment; the architecture is acting on whatever
input lands in the frames), with the remaining $\sim 17$pp accruing
to canonical alignment. We do not claim this decomposition is
unique; it is a description of the observed accuracy stack.

\subsection{Conversation neutrality (P14--P16)}\label{ssec:conv}

\textbf{Setup.} $64$ utterances across $8$ dialogue-act categories.
$8$-dimensional injection-row features per utterance. Identical
substrate routing pipeline to chess.

\begin{table}[ht]
\centering
\small
\caption{Conversation preregistered tests.}
\label{tab:conv_prereg}
\begin{tabular}{l l l l l}
\toprule
ID & Test & Threshold & Observed & Verdict \\
\midrule
P14 & raw 5-fold CV (seeds 30220--30224)    & $\geq 75\%$ & $87.5\%$ & $\checkmark$ \\
P15 & substrate lift                         & $|\cdot| < 10$pp & $-4.4$pp & $\checkmark$ \\
P16 & null perm. mapping (15 perms)         & $\geq 50\%$ & $70.6\%$ & $\checkmark$ \\
\bottomrule
\end{tabular}
\end{table}

\textbf{Reading.} Conversation raw features at $87.5\%$ are already
strongly discriminative (cf.\ chess raw $\sim 53\%$); the substrate
lift is $-4.4$pp, well within the preregistered neutrality band
$|\cdot|\!<\!10$pp. The substrate is approximately neutral on conversation.

\textbf{Selective amplifier signature.} The pair (chess
$+40.6$pp lift; conversation $-4.4$pp lift) is consistent with the
selective-amplifier behaviour preregistered in 2026-04-18: in these
two tasks, the architecture amplifies when raw features are ambiguous
(chess raw $\sim 53\%$) and is approximately neutral when raw features
are already saturated (conversation raw $87.5\%$). We do not claim
this generalises to all classification tasks; cross-domain transfer
to additional ambiguous-feature benchmarks is an open build
(\S\ref{sec:limitations}).

\subsection{HCP brain-graph H$_4$-transitive null
            (P17--P18)}\label{ssec:hcp}

\textbf{Setup.} Reference cohort: Human Connectome Project (HCP)
$n=1003$ subjects~\citep{VanEssen2013HCP}; preregistered tests on
$n=100$ subjects for computational tractability, with full-cohort
$n=1003$ descriptive statistics also reported. ICA-50 group-averaged
connectivity matrix; thresholded at the same density as ARIA's
vertex graph ($\rho=0.101$). Compare degree distribution and
higher-order graph statistics. ARIA reference: $G_{V_{600}}$ with
$\Lop$. By H$_4$ transitivity (\S\ref{ssec:graph}) every vertex
has identical local structure $\Rightarrow$ uniform degree $12$
$\Rightarrow$ degree std $= 0$ as a theorem.

\begin{table}[ht]
\centering
\small
\caption{HCP comparison: preregistered $n=100$ test plus full-cohort
$n=1003$ descriptive statistics.}
\label{tab:hcp}
\begin{tabular}{l r r r}
\toprule
Metric & ARIA & HCP $n=1003$ mean (sd) & $\sigma$ from HCP \\
\midrule
Degree std (preregistered, $n=100$ subset) & $0.000$ & $3.388$ ($> 2.0$) & --- \\
Degree std (descriptive, $n=1003$)         & $0.000$ & $3.28\pm 0.28$ & $-11.58\sigma$ \\
Participation ratio (descriptive)          & $68.54$ & $19.72\pm 0.61$ & $+79.78\sigma$ \\
Clustering coefficient (descriptive)$^{\flat}$ & $0.455$ & $0.220$ & $+6.80\sigma$ \\
\bottomrule
\end{tabular}
\end{table}

\noindent$^{\flat}$ The HCP across-subject standard deviation for the
clustering coefficient is not separately reported in
\texttt{CROSS\_DOMAIN\_RESULTS.md}; the $+6.80\sigma$ value is sourced
from the same descriptive analysis as the other rows. Inferred from
the reported gap and $\sigma$, the implicit HCP sd is
$\approx 0.235/6.80\!\approx\!0.035$. We carry the $\sigma$-distance
forward as reported and flag the missing explicit sd here.

\begin{itemize}\itemsep=2pt
\item P17 (ARIA degree std, theorem): predicted $=0$, observed
  $0.0000$, $\checkmark$.
\item P18 (HCP ICA-50 degree std, $n=100$ density-matched):
  predicted $> 2.0$, observed $3.388$, $\checkmark$. Zero of $1003$
  HCP subjects have degree std below $2.0$.
\end{itemize}

\textbf{Reading (substrate witness).} ARIA's H$_4$-transitive
structure is a deterministic group-theoretic null reference for
cortical functional connectivity. Real cortex breaks the symmetry
through hub-spoke functional specialisation; the $\sigma$-distances
quantify the magnitude of biological symmetry-breaking with no
fitted parameters. The $\sigma$-distances ($-11.58\sigma$ on degree
homogeneity, $+79.78\sigma$ on participation ratio, $+6.80\sigma$ on
clustering coefficient) are large on the ICA-50 pipeline at the
density-matched threshold $\rho = 0.101$; cross-parcellation
replication (Schaefer, Glasser) remains an open build.

\textbf{Participation-ratio comparability.} ARIA's vertex graph has
$120$ nodes; the HCP ICA-50 connectivity matrix has $50$ nodes. The
participation-ratio statistic
$\mathrm{PR}(G) = (\sum_{i} d_{i})^{2} / \sum_{i} d_{i}^{2}$ is
node-count-dependent — its theoretical maximum is the node count of
the graph. We report the raw $\mathrm{PR}$ values
($\mathrm{ARIA}=68.54$ on a 120-node graph; $\mathrm{HCP}=19.72$ on a
50-node graph) and the $\sigma$-distance against the HCP
across-subject distribution, but the $+79.78\sigma$ value reflects
both the architectural difference and the differing node counts. A
node-count-normalised statistic
$\mathrm{PR}/|V|$ gives $\mathrm{ARIA}=0.571$ vs $\mathrm{HCP}=0.394$,
a smaller absolute gap; we keep the raw-PR comparison as headline
because the HCP subject distribution and the across-subject
$\sigma$ are computed in the same units, but flag the node-count
caveat here.

\textbf{What we do not claim.}
\begin{itemize}\itemsep=2pt
\item We do not claim cortex has ``drifted from an ideal polytope'';
  the substrate is a useful a-priori null whose deviation from real
  cortex is precisely measurable.
\item We do not claim parcellation invariance: the $\sigma$-distances
  are reported on ICA-50; alternative parcellations (Schaefer,
  Glasser) would give different per-metric numbers but, on the
  basis of the qualitative pattern that cortex is hub-concentrated
  relative to ARIA's transitive null, we expect them to preserve the
  signs. Verification across parcellations is an open build
  (\S\ref{sec:limitations}).
\end{itemize}

\subsection{Cross-domain summary as a selective amplifier
            \texorpdfstring{$+$}{+} H$_4$-transitive null}

\begin{table}[ht]
\centering
\small
\caption{Cross-domain summary on a single substrate.}
\label{tab:cross_domain_summary}
\begin{tabular}{l r r r r r}
\toprule
Task & Raw & Substrate & Null perm. & Geom.\ floor & Sem.\ alignment \\
\midrule
Chess (LOO, $n=25$, w/ reset) & $53.1\%$ & $93.8\%$ & --- & --- & $+40.6$pp lift \\
Chess (5-fold CV)             & ---      & $83.1\%$ & $65.4\%$ & $65.4\%$ & $+17.7$pp \\
Conversation (5-fold CV)      & $87.5\%$ & $83.1\%$ & $70.6\%$ & $70.6\%$ & $+12.5$pp (substrate vs null) \\
\bottomrule
\end{tabular}
\end{table}

The geometric content ($\approx 65$--$71\%$ across the two domains)
is the architecture-invariant null floor. The semantic content
($12$--$18$pp) is the domain-specific contribution. On HCP,
$\sigma$-distances against the biological cohort are
$(-11.58, +79.78, +6.80)$ on (degree std, participation ratio,
clustering coefficient).

\textbf{Headline cross-domain reading.} The substrate is
\emph{selectively} amplifying (not unconditionally), and it is an
H$_4$-transitive deterministic null on connectivity (not a fitted
model). Both readings sit inside the substrate-witness scope.
