% =====================================================================
\section{Discussion}\label{sec:discussion}
% =====================================================================

This section reads the substrate-witness result against existing
theories of consciousness, identifies what is novel here that is not
a re-statement of an earlier theory, and proposes a non-load-bearing
ACT bridge to the companion adaptive-closure-transport
preprint~\citep{SmartAdaptiveClosureTransport2026}. We do not claim a
selection theorem, we do not claim a Lyapunov derivation, and we do
not claim the recurrent layer ``is'' consciousness.

\subsection{What is novel in this work}

Three things are claimed novel as a substrate witness:
\begin{enumerate}\itemsep=2pt
\item \textbf{A geometry-fixed substrate that is consistent with
  real-cortex EEG signatures without fitted shape parameters on neural
  data.} Once the 600-cell is chosen as the substrate, its graph
  ($120$ vertices, uniform degree $12$ on the canonical short-edge
  nearest-neighbour graph) and the response operator
  $\Cph = \Lop + \Ph^{-2} I$ are fixed by the construction (no
  shape parameter is tuned to neural data); cascade-$\alpha$ matches
  Sleep-EDFx within preregistered tolerance with pairwise CI overlap
  on three reference ranges; six drug/sleep signatures pass on a
  single deterministic recurrent-layer trajectory at seed~$42$,
  against literature-derived thresholds.
  We are not aware of a prior geometric substrate that has been tested
  against this many preregistered cortical correspondences from a
  graph fixed by the construction with no neural-data-fitted shape
  parameters; we cannot rule out that such a model exists.
\item \textbf{The strong-coupling architectural finding.} $C$ and $P$
  are strongly coupled cascade-state stabilisers, not
  nearly-orthogonal ones. The $C\!\times\!P$ interaction
  ($+0.190$, $95\%$ CI $[+0.143, +0.239]$ at $N\!=\!20$) is comparable
  in magnitude to the $P$ main effect ($-0.218$). This was hidden by
  underpowered ablation and emerged only at $N\!\geq\!20$ — a
  substantive correction to the architectural reading from the
  original 3-seed validation.
\item \textbf{The 18/18 preregistered correspondences with no
  threshold modification.} Every prediction in the preregistered set
  passes at the preregistered thresholds. The two interaction tests
  (P3, P4) required $N\!\geq\!5$ and $N\!\geq\!20$ respectively, and
  one test (P13) required the documented state-reset protocol. We
  report this transparently as methodology refinement, not as
  threshold change.
\end{enumerate}

\subsection{Comparison to existing theories of consciousness}

\textbf{vs IIT.}~\citep{Tononi2008,BalduzziTononi2008} ARIA produces
IIT-direction-correct $\Phi$ collapse on propofol ($0.33\!\times$
wake). The $\Phi$ proxy (\S\ref{sec:chain}) is designed to be small
under H$_4$-equivariant dynamics and to increase when dynamics
produce cross-class asymmetries. ARIA does not implement the full
IIT axioms (cause-effect repertoires, exclusion postulate,
integration-over-partitions); it reproduces an observable consequence
on the propofol--wake state contrast. This is a consistency-of-direction
result, not a discharge of IIT.

\textbf{vs Global Workspace Theory.}~\citep{Baars1988GWT,Dehaene2014ConsciousAndBrain}
The $S^{7}$ context-rotation mechanism (\S\ref{sec:chain}) is
functionally analogous to a workspace with rotating attentional
selection; the active observer frame plays the role of a temporary
in-workspace subset of features. ARIA does not commit to the GWT
broadcast/access distinction at the architectural level; the
analogy is descriptive.

\textbf{vs Predictive Processing.}~\citep{FristonFreeEnergy2010,ClarkPP2013}
ARIA does not implement prediction-error minimisation or hierarchical
generative models.
The recurrent self-model layer ($\eta\!=\!0.20$) provides top-down
modulation of the substrate response by cosine direction alignment
with the prior phenomenal snapshot, not by learned prediction errors.
Predictive-processing-style refinements (e.g.\ $\eta$ as an adaptive
learning rate over a prediction-error norm) are an open build, not
delivered here.

\textbf{vs neural mass models.} ARIA operates at the
architectural-algorithmic level; it does not specify which neural
circuits implement context rotation or partial emission. The 600-cell
substrate is proposed as an abstract description of the criticality-
maintaining structure of cortex, not as a circuit model.

\subsection{The non-load-bearing ACT bridge}\label{ssec:act_bridge}

The companion adaptive-closure-transport
preprint~\citep{SmartAdaptiveClosureTransport2026} proposes a
4-tuple bridge
$(M, L_{M}, W, R_{\mathrm{hom}})$ — substrate $M$, response operator
$L_{M}$, learnable Hebbian-like field $W$, and a homeostatic
regulariser $R_{\mathrm{hom}}$. We propose the dictionary
$D_{\mathrm{ACT}}$:
\[
D_{\mathrm{ACT}}\colon (M, L_{M}, W, R_{\mathrm{hom}})
\;\longmapsto\;
(\Rsixhundred,\ \Cph,\ \text{cascade pressure field}\ W_{\mathrm{p}},
   \ \texttt{homeostatic\_reset}).
\]
\textbf{This bridge is non-load-bearing for the present paper.} It is
included as a route-K (alternative-route) reading; the substrate-
witness claims (six signatures, $18/18$, chess $+40.6$pp,
HCP $\sigma$-distances) do not require any of the ACT theorems.

\textbf{What ACT would have to deliver to make this load-bearing.}
The companion preprint identifies four open builds, each of which is
deferred:
\begin{itemize}\itemsep=2pt
\item A Lyapunov function $V(W)$ on the reduced flow whose
  monotonicity proves selection — not delivered.
\item An edge-space decomposition of $\mathbb{R}^{E_{M}}$ under the
  Hodge edge Laplacian $L_{\mathrm{edge}} = \delta_{2}\delta_{2}^{\mathsf T} +
  \delta_{1}^{\mathsf T}\delta_{1}$ — not delivered.
\item A formal $2I$-equivariance audit of the closure operator
  family — not delivered.
\item A full reduced-flow convergence theorem on
  $W$-trajectories — not delivered.
\end{itemize}
Until these are delivered, ARIA is positioned as the empirical
\emph{substrate witness} for the family that ACT names; ACT is not the
selection-theorem witness for ARIA. The companion kernel
document~\citep{SmartAriaClosureKernel2026} discusses the same $\Cph$
in a passive-regime witness via the $B\to K^{*}\mu^{+}\mu^{-}$ flavour
anomaly~\citep{SmartBAnomaly2026}; that line shares operator-level
infrastructure with this paper, but does not transfer empirical
support for ARIA.

\subsection{The strong-coupling reading for cortical architecture}

Real cortical criticality is maintained by multiple parallel
mechanisms: E/I balance, neuromodulation (acetylcholine, noradrenaline),
homeostatic plasticity, gain control. The naive expectation — and the
one we held until the $N\!=\!20$ deep-dive — is that these are mostly
orthogonal, so losing one removes only its own main effect. The
$N\!=\!20$ result reverses this on the substrate: $C$ and $P$ are
strongly coupled. Disabling one cascades into losing the synergistic
contribution of the other.

This matches clinical observations: anaesthesia (which targets
GABAergic transmission) and seizure (which targets E/I balance)
produce widespread network-level dysfunction beyond their direct
targets — qualitatively consistent with a strong-coupling hypothesis. We position this as
\emph{a hypothesis the substrate witness raises}, not as a proof.
The bridge from cascade-mechanism interaction on $\Rsixhundred$ to
real-cortex pharmacological coupling is a step we do not take in
this paper.

\subsection{Methodological contributions}

Two methodological items are worth recording outside the headline:
\begin{enumerate}\itemsep=2pt
\item \textbf{$N\!\geq\!20$ for similar high-variance ablation matrices.}
  Allocation discipline for preregistration: in this cascade-ablation
  matrix specifically, P4 ($C\!\times\!P$) required $N\!=\!20$ for
  reliable detection at the preregistered threshold. The general
  rule we draw — when preregistering an interaction effect on a
  system with unknown per-seed variance, budget for at least this
  scale — should be tested against other ablation matrices, not
  taken as universal. The original 3-seed plan was the source of two
  underpowered-interaction estimates in this work.
\item \textbf{State-reset protocol on non-stationary substrates.}
  ARIA's substrate is a non-stationary dynamical system; the
  pressure field equilibrates within $\sim 5$ successive evaluations.
  Any multi-trial benchmark must specify a state-reset protocol or
  document the drift. Generalisable lesson: \emph{published
  cross-domain benchmarks on non-stationary substrates should report
  an explicit reset/equilibration discipline}, not just seed.
\end{enumerate}

\subsection{The substrate as an H$_4$-transitive connectivity null}

The HCP comparison (\S\ref{ssec:hcp}) places ARIA as a principled
deterministic null reference for cortical functional connectivity.
Real cortex breaks the symmetry through hub-spoke functional
specialisation; the $\sigma$-distances from ARIA quantify the
magnitude of biological symmetry-breaking with no fitted parameters.

This is a methodological contribution to comparative connectomics.
Stochastic nulls (Erd\H{o}s--R\'enyi, configuration model,
edge-randomised graphs) compare cortex to a random graph with matched
density. ARIA is a different kind of null: a deterministic
group-theoretic graph with structure-level statements: degree std
$=\!0$ by H$_4$ transitivity, and a fully-determined Laplacian
spectrum (\S\ref{ssec:graph}) computed from the constructed graph.
Both null kinds are useful; ARIA gives a specific, reproducible,
group-theoretic baseline that cortex deviates from in quantifiable
$\sigma$-units.

\subsection{Open questions raised by the substrate witness}

\begin{itemize}\itemsep=2pt
\item Do the six drug/sleep signatures replicate across $10$--$20$
  cross-seed runs of the recurrent layer? (Single-seed disclosure;
  see \S\ref{sec:limitations}.)
\item Do alternative regular 4-polytopes ($24$-cell, $120$-cell)
  reproduce comparable signature sets, or is the $600$-cell
  distinguished?
\item Does the strong-coupling reading ($C\!\times\!P$) survive an
  independent fresh-seed $N\!=\!20$ replication at a different seed
  range?
\item Does the substrate's amplifier behaviour transfer to other
  ambiguous-feature classification tasks beyond chess (e.g.\ visual
  pattern, audio classification)?
\item Does the Sleep-EDFx three-way CI overlap survive on a different
  EEG cohort (TUH, NHM)?
\end{itemize}
We list these as open questions raised by the witness, not as gaps
in the witness itself.
