% =====================================================================
\section{Conclusion}\label{sec:conclusion}
% =====================================================================

The 600-cell regular 4-polytope $\Rsixhundred$ under H$_4$ Coxeter
symmetry, with the shifted-Laplacian response operator
$\Cph = \Lop + \Ph^{-2} I$ ($\Ph=(1+\sqrt 5)/2$), is a
geometry-fixed substrate that is consistent with eighteen
preregistered neuroscience correspondences plus six companion
drug/sleep EEG signatures of conscious vs unconscious states. Once
the substrate is chosen, its graph structure ($120$ vertices, uniform
degree $12$ on the canonical short-edge nearest-neighbour graph, with
the Laplacian spectrum reported in~\S\ref{ssec:graph} as observed) is
fixed; only one condition-dependent self-injection coupling
$\eta\in\{0, 0.05, 0.20\}$ and one substrate-pinned nonlinearity
$\mathrm{bounded\_topk}(\cdot, k\!=\!12)$ at the graph's average
degree enter the recurrent layer above the substrate. No shape
parameter is tuned to any neural dataset.

\textbf{Headline tally.} On a single deterministic trajectory, six
drug/sleep EEG signatures pass against their literature-derived
thresholds (Sleep-EDFx CI, OpenNeuro \texttt{ds005620}, Brodbeck 2012,
Tononi 2008): NREM-N3 phenomenal-intensity variance ratio
$0.463\!\times$ wake; propofol modality-switching $1.83\!\times$ wake;
propofol continuity drop $+0.066$; propofol integrated-information
$\Phi$ collapse to $0.33\!\times$ wake (IIT direction confirmed);
recovery deterministically identical to wake under the WAKE stimulus
protocol; wake cortical-avalanche power law $\alpha\!=\!2.252$,
$95\%$ CI $[1.82, 2.86]$, $R^{2}\!=\!0.956$. The wake $95\%$ CI
overlaps the real Sleep-EDFx EEG
$95\%$ CI ($n\!=\!30$ subjects, $\alpha\!=\!2.51$,
CI $[2.50, 2.53]$) and ARIA's prior cascade pipeline CI
$[2.73, 3.25]$.

\textbf{Eighteen preregistered correspondences.} All eighteen pass at
preregistered thresholds, with two interaction tests requiring
$N\!\geq\!5$ and $N\!=\!20$ respectively for reliable detection of
high-variance interaction terms, and one cross-domain test requiring
the documented \texttt{homeostatic\_reset} state-reset protocol. No
preregistered threshold has been modified. The original 2026-04-20
$15/18$ tally was a methodology-limited reading, not a content
failure; the closure of the three gaps (P3, P4, P13) is documented
transparently in
\texttt{docs/brain\_mapping/VALIDATION\_RESULTS\_2026-04-29.md}.

\textbf{Strong-coupling architectural finding.} Two cascade
mechanisms — context rotation $C$ and partial emission $P$ — are
causally identifiable within the factorial ablation model and exhibit
strong synergy: their interaction $\Delta_{CP}\!=\!+0.190$ at
$N\!=\!20$ ($95\%$ bootstrap CI $[+0.143, +0.239]$, $0/2000$ resamples
at or below zero, reported as $0.0000$) is comparable in magnitude to
the $P$ main effect $-0.218$. The original 3-seed estimate ($+0.044$)
is consistent with an underpowered interaction estimate on a
high-per-seed-variance term ($\mathrm{std}=0.089$ at $N\!=\!20$); we
contribute $N\!\approx\!20$ as a planning scale for this cascade
matrix, recommended as a preregistration-practice consideration for
similar high-variance ablation matrices.

\textbf{Cross-domain selectivity.} The substrate exhibits selective
amplification on the two tasks tested: chess 4-category position
classification on 8-D V2 features lifts $+40.6$pp on leave-one-out at
canonical depth $n\!=\!25$ ticks (raw $53.1\%$ $\to$ substrate-routed
$93.8\%$, with state reset; preregistered threshold $\geq +15$pp on
$5$-fold CV — the LOO finding above is a disclosed estimator/protocol
refinement at the same threshold), while
conversation utterance classification at raw $87.5\%$ lifts $-4.4$pp
(threshold $|\cdot|\!<\!10$pp) — and as an H$_4$-transitive
deterministic null reference for cortical functional connectivity:
on the full-cohort descriptive HCP $n\!=\!1003$ statistics
(preregistered test on the $n\!=\!100$ subset), ARIA's H$_4$-transitive
structure is at $-11.58\sigma$ on degree homogeneity,
$+79.78\sigma$ on participation ratio (with the node-count caveat of
\S\ref{ssec:hcp}), and $+6.80\sigma$ on clustering coefficient.

\textbf{Substrate-witness scope.} This is a substrate witness, not a
derivation of consciousness, not a selection theorem on the
companion adaptive-closure-transport
4-tuple~\citep{SmartAdaptiveClosureTransport2026}, and not a
uniqueness claim for the 600-cell among regular 4-polytopes. The
strengthening builds — cross-seed CIs on the recurrent-layer
signatures, alternative-polytope ablations, an independent $N\!=\!20$
C$\times$P replication at a different seed range, cross-parcellation
HCP replication, a Lyapunov function on the reduced flow,
$2I$-equivariance audit of the closure operator family — are
explicitly listed in~\S\ref{sec:limitations} and remain open.

We are not aware of a prior deterministic geometric architecture
tested against this many preregistered cortical correspondences from
a graph fixed by the construction with no shape parameters tuned to
neural data; we cannot rule out that such prior work exists. The empirical material
gathered here is the substrate witness; the broader programme to
turn the witness into a selection-theorem-grade claim — including the
independent passive-regime witness via the $B\to K^{*}\mu^{+}\mu^{-}$
flavour anomaly~\citep{SmartBAnomaly2026} on the same response
operator $\Cph$ — is sketched in the companion preprints and remains
the natural next step.
