\documentclass[11pt]{article}
\usepackage{amsmath,amssymb,amsthm,mathtools}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\newtheorem{theorem}{Theorem}
\newtheorem{proposition}[theorem]{Proposition}
\theoremstyle{definition}
\newtheorem{result}[theorem]{Certified Result}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\zetaK}{\zeta_{K}}

\title{A Verification-Gated Engine for Geometry$\to$Arithmetic\\ Correspondences,
and the $(O2)$ Negative}
\author{(VFD programme --- working draft)}
\date{2026}

\begin{document}
\maketitle

\begin{abstract}
We describe a reproducible, gate-first engine for testing proposed correspondences
between geometric objects and arithmetic $L$-functions, and we use it to settle one
specific question in the negative. A proposed correspondence (``expression'') is
accepted only if it passes an eight-check verification gate---chief among them a
\emph{provenance} check that rejects any map fitted to its target, and a battery of
spectral discriminators (zero positions, density, nearest-neighbour spacing, number
variance) that separate one $L$-function from another. Applied to the VFD programme,
the engine \emph{certifies} that the icosian substrate realizes $\zetaK$ (and
$L(\chi_5)$, $\zeta(s)\zeta(s-1)$), and \emph{rejects}, on five independent grounds,
the stronger claim that it realizes the Riemann $\zeta$ itself. We call this the
$(O2)$ negative. The engine issues verification certificates only and makes no
claim about the Riemann Hypothesis.
\end{abstract}

\section{Motivation: the hardcoding trap}
Claims of the form ``geometric object $X$ realizes arithmetic object $Y$'' are easy
to make and easy to fake: one may fit a map from $X$'s data to $Y$'s known invariants
and then ``recover'' $Y$. Such a map encodes nothing---it is circular. A prior
internal audit of this programme (the ``$W5$'' audit) caught exactly such a fitted
substrate$\to$eigenvalue map. The purpose of the present engine is to make the
verification, rather than the proposal, the primary and reproducible step: a
correspondence is a \emph{result} only if it survives an explicit gate that a fitted
or overclaimed map cannot.

\section{The engine}
An \emph{expression} is a tuple $(\,\text{geometry},\ \text{claim},\
\text{predict},\ \text{meta},\ \text{provenance}\,)$, where \texttt{predict} is a
parameter-free procedure that produces the on-line zeros of the claimed $L$-function
\emph{without taking the target as input}. The gate evaluates eight checks:
\begin{enumerate}\itemsep1pt
\item \textbf{no-hardcoding}: the target is not an input to \texttt{predict};
\item \textbf{provenance}: no constant is fitted or tuned to the target;
\item \textbf{fingerprint}: nearest-neighbour spacing (fraction of gaps $<0.3$),
which separates a single $L$-function ($\approx 0$, GUE repulsion) from a
superposition;
\item \textbf{pointwise}: agreement of the first $k$ zeros;
\item \textbf{degree} and \textbf{pole}: structural metadata match;
\item \textbf{density}: zeros-per-unit-height (Weyl rate);
\item \textbf{rigidity}: number variance $\Sigma^2(L)$.
\end{enumerate}
An expression is \textbf{verified} iff it passes all applicable checks; a
certificate (a hashed record of every check and its evidence) is emitted either way.
A proposer interface (``ARIA'') may suggest expressions; it never certifies---the
gate does. The engine is deliberately a \emph{certifier}, not a search: most honest
outputs are negatives.

\section{The certified atlas}
Running the engine across the programme's correspondences yields:
\begin{result}\label{res:pos}
The following are \textbf{verified} (all eight checks): the icosian/$V_{600}$
substrate $\leftrightarrow \zetaK(s)\zetaK(s-1)$; the $24$-cell
$\leftrightarrow \zeta(s)\zeta(s-1)$; the $\varphi$-shell $\leftrightarrow L(\chi_5)$.
\end{result}
\begin{result}[the $(O2)$ negative]\label{res:o2}
The proposed correspondence ``substrate $\leftrightarrow \zeta$'' is \textbf{rejected}
on five independent checks: fingerprint ($0.077$ vs.\ $0.0$), pointwise (first-zero
displacement $\approx 22$), density, rigidity, and degree ($4$ vs.\ $1$). The
substrate realizes $\zetaK=\zeta\cdot L(\chi_5)$, which \emph{contains} $\zeta$ as a
factor but does not \emph{isolate} it.
\end{result}
\begin{result}[$(O2)$, independent second route]\label{res:o2b}
The negative is confirmed by a route disjoint from the spacing discriminators: a
direct \emph{count} of on-line zeros. Since $\zetaK=\zeta\cdot L(\chi_5)$, the
substrate's $L$-function carries, on the critical line, the union of the zeros of
$\zeta$ and of $L(\chi_5)$; its on-line zero density therefore strictly exceeds that
of $\zeta$ (empirically a factor $\approx 3$--$4.7$ up to height $50$, tending to $2$
asymptotically as each factor contributes $\sim\tfrac1{2\pi}\log\tfrac{T}{2\pi}$).
A spectrum strictly denser than $\zeta$'s is not $\zeta$'s.
\end{result}
The two routes---spacing/fingerprint (Result~\ref{res:o2}) and counting/density
(Result~\ref{res:o2b})---are independent and agree.

Two controls confirm the gate's discrimination:
\begin{result}[fitted-map control]
A map declared as least-squares fitted to the zeros is rejected on \emph{provenance}
alone, regardless of how well its output matches---reproducing the $W5$ catch
mechanically.
\end{result}
\begin{result}[class-vs-individual control]
A generic GUE operator (random Hermitian, no arithmetic) \emph{passes} fingerprint
and rigidity (it has the right universality class) but \emph{fails} pointwise and
density (it is not the right individual). Thus GUE-class membership is certified to
be insufficient to be $\zeta$.
\end{result}

\begin{center}\small
\begin{tabular}{llp{4.6cm}}
\hline
expression & verdict & failing checks (cert id)\\
\hline
$24$-cell $\to \zeta(s)\zeta(s-1)$ & \textsc{verified} & --- \quad(\texttt{2c9f2a5a})\\
icosian $\to \zetaK$ & \textsc{verified} & --- \quad(\texttt{0ae81ff8})\\
$\varphi$-shell $\to L(\chi_5)$ & \textsc{verified} & --- \quad(\texttt{080d0d3c})\\
substrate $\to \zeta$ (bridge) & \textsc{rejected} & fingerprint, pointwise, density, rigidity, degree \quad(\texttt{f5c303f3})\\
fitted map $\to \zeta$ & \textsc{rejected} & provenance \quad(\texttt{50a16461})\\
generic GUE $\to \zeta$ & \textsc{rejected} & pointwise, density, rigidity \quad(\texttt{051fff64})\\
\hline
\end{tabular}
\end{center}
\noindent The certified atlas (hashes abbreviated). Each \textsc{rejected} entry
fails for a reason intrinsic to the discriminator named, not by adjudication.

\section{Reproducibility and scope}
All certificates are content-hashed and emitted to a machine-readable atlas; the
engine, gate, and scripts accompany this note. We claim only what the certificates
state: which correspondences pass the gate and which do not. In particular,
Result~\ref{res:o2} is a statement that a specific \emph{construction} fails to
realize $\zeta$; it is \textbf{not} a statement about the zeros of $\zeta$, and
implies nothing about the Riemann Hypothesis. No proof certificate is defined or
issued.

\end{document}
