\documentclass[12pt]{article}

\usepackage[a4paper, margin=2.5cm]{geometry}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage[colorlinks=true, linkcolor=blue, citecolor=blue, urlcolor=blue]{hyperref}
\usepackage{booktabs}
\usepackage{natbib}

\newtheorem{theorem}{Theorem}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{remark}[theorem]{Remark}

\newcommand{\Fcal}{\mathcal{F}}
\newcommand{\Ncl}{\mathcal{N}_{\mathrm{cl}}}
\newcommand{\VW}{V_{\mathrm{W}}}
\newcommand{\HW}{H_{\mathrm{W}}}
\newcommand{\Sadm}{\mathcal{S}_{\mathrm{adm}}}
\newcommand{\Rst}{R_{\mathrm{st}}}
\newcommand{\dUrel}{\delta U_{\mathrm{rel}}}

\title{From Closure Dynamics to Quantum Structure:\\[6pt]
\large A Structural Synthesis of the Dissipative-to-Schr\"odinger Programme}
\author{%
  Lee Smart\\[4pt]
  \textit{Institute of Vibrational Field Dynamics}\\[2pt]
  \texttt{contact@vibrationalfielddynamics.org}\\[2pt]
  \texttt{@vfd\_org}%
}
\date{April 2026}

\begin{document}

\maketitle

\noindent\textbf{Status:} Preprint (not peer-reviewed)

\begin{abstract}
Papers~XVI--XX of the VFD closure programme, building on the closure dynamics developed in Papers~XII--XV, have developed a layered dynamical framework beginning from dissipative gradient-flow systems and culminating in a nonlinear Schr\"odinger-type equation with an exact equilibrium linear limit. The present paper provides a structural synthesis of those results.

The central outcome is that the Schr\"odinger equation is not introduced axiomatically within this framework, but emerges as the equilibrium tangent limit of a paired forward/backward stochastic process constructed from the closure dynamics. The underlying paired dynamics is nonlinear, with a modulus-curvature residual term that vanishes exactly at equilibrium and is perturbatively small near it.

This paper clarifies what has been established, what has not been established, and how the resulting framework relates to the Madelung, Nelson, and nonlinear Schr\"odinger formulations. The goal is not to replace or reinterpret quantum theory, but to identify a mathematically consistent pathway by which Schr\"odinger-type evolution can arise from a broader dynamical system without being postulated.
\end{abstract}

%==================================================================
\section{Introduction}\label{sec:intro}
%==================================================================

The Schr\"odinger equation occupies a central role in modern physics. It is typically introduced as a fundamental postulate governing wavefunction evolution. A long-standing question in mathematical physics is whether such dynamics can instead arise from deeper structural principles, without quantum axioms as the starting point.

The closure programme (Papers~XII--XX) approaches this question from a specific direction: starting from dissipative gradient-flow systems and stochastic dynamics, and asking what happens when the resulting stochastic structure is taken seriously as a paired forward/backward system.

The purpose of this paper is to synthesise the technical results of Papers~XVI--XX~\citep{SmartXVI, SmartXVII, SmartXVIII, SmartXIX, SmartXX} into a coherent structural statement, identify the precise relationship to known quantum formulations, and clearly delineate what has and has not been achieved.

\subsection*{Core question}

\begin{quote}
Can Schr\"odinger-type dynamics emerge from a broader, non-unitary dynamical system without being postulated?
\end{quote}

The programme supports a qualified positive answer: Schr\"odinger-type evolution emerges as the equilibrium tangent limit of a nonlinear closure-paired dynamics. The emergence is exact at equilibrium and perturbative near it; away from equilibrium, the dynamics is genuinely nonlinear.

%==================================================================
\section{The Five-Stage Construction}\label{sec:stages}
%==================================================================

The programme proceeds through five stages, each building on the previous:

\subsection{Stage 1: Closure evolution equation (Paper~XVI)}

Starting from the Paper~XIV~\citep{SmartXIV} Langevin equation $d\psi = -\nabla\Fcal\,dt + \sigma\,dW$ and the Paper~XV~\citep{SmartXV} complexification ansatz, a differential evolution equation is derived:
\begin{equation}\label{eq:XVI}
    \partial_t\Psi = \frac{\sigma^2}{2}\Delta\Psi - \nabla\Fcal\cdot\nabla\Psi + \frac{i}{\sigma^2}\Fcal\,\Psi
\end{equation}

This equation has the operator structure of a Fokker--Planck--Schr\"odinger hybrid: the backward Kolmogorov generator of the stochastic closure dynamics plus an imaginary (phase-generating) potential. Crucially, the kinetic term $(\sigma^2/2)\Delta$ is \textit{real diffusion}, not the imaginary propagation required for quantum evolution.

\subsection{Stage 2: Dissipative obstruction (Paper~XVII)}

Paper~XVII proves that the kinetic-term gap identified in Stage~1 is structural, not accidental:

\begin{quote}
\textit{Any generator of the form $L_{\mathrm{BK}} + iM$, where $L_{\mathrm{BK}}$ is a backward Kolmogorov generator from a real Langevin process, is dissipative in the equilibrium-weighted inner product --- regardless of the choice of imaginary potential $iM$.}
\end{quote}

This is the dissipative obstruction theorem. It establishes that phase complexification alone (Paper~XV) cannot produce unitary evolution: the obstruction sits in the kinetic sector, not the phase sector. Future progress must target a reversible kinetic construction.

\subsection{Stage 3: Nelson pairing (Paper~XVIII)}

Paper~XVIII constructs a paired forward/backward system using the closure Langevin process and its time-reversal. Defining the current velocity $v = (b_+ + b_-)/2$ and the osmotic velocity $u = (b_+ - b_-)/2$, the paper derives:
\begin{itemize}
    \item an exact continuity equation $\partial_t\rho + \nabla\cdot(\rho v) = 0$,
    \item an automatically irrotational phase function $S = -\Fcal - \sigma^2\log\sqrt{\rho}$,
    \item an exact closure Madelung pair,
    \item a nonlinear wave equation with the correct imaginary kinetic structure:
\end{itemize}
\begin{equation}\label{eq:XVIII}
    i\sigma^2\partial_t\Psi = -\frac{\sigma^4}{2}\Delta\Psi + \left(-\sigma^2\VW + \sigma^4\frac{\Delta|\Psi|}{|\Psi|}\right)\Psi
\end{equation}

The imaginary kinetic term $-(\sigma^4/2)\Delta$, absent in the single-generator framework, emerges from the forward/backward pairing. This evades the XVII obstruction because the paired dynamics is not of the $L_{\mathrm{BK}} + iM$ form.

\subsection{Stage 4: Residual isolation (Paper~XIX)}

Paper~XIX rewrites the dynamics in equilibrium-centred normal form:
\begin{equation}\label{eq:XIX}
    i\,\partial_t\Psi = \HW\Psi + \dUrel[\Psi]\,\Psi
\end{equation}
where $\HW = -(\sigma^2/2)\Delta + \VW$ is the Witten Hamiltonian and:
\begin{equation}\label{eq:residual}
    \dUrel[\Psi] = \sigma^2\left(\frac{\Delta|\Psi|}{|\Psi|} - \frac{\Delta\Rst}{\Rst}\right)
\end{equation}

The residual is proved to be: phase-invariant, modulus-dependent, local, amplitude-homogeneous, exactly zero at equilibrium, and $O(\varepsilon)$ near equilibrium. The full nonlinear equation conserves the $L^2$ norm exactly.

\subsection{Stage 5: Structural classification (Paper~XX)}

Paper~XX identifies the closure equation as a derived modulus-curvature nonlinear Schr\"odinger structure:
\begin{itemize}
    \item the nonlinear term belongs to the same broad class as Doebner--Goldin-type modulus-curvature corrections,
    \item the coefficient $\sigma^2$ (in Schr\"odinger-normalised form) and the equilibrium subtraction $2\VW$ are fixed by the closure geometry, not introduced phenomenologically,
    \item the linear Schr\"odinger equation is the Fr\'echet tangent dynamics at the equilibrium state.
\end{itemize}

%==================================================================
\section{The Central Structural Result}\label{sec:result}
%==================================================================

The combined outcome of Papers~XVI--XX is:

\begin{quote}
\textbf{Structural synthesis.} The closure programme establishes an exact nonlinear paired-process wave equation whose equilibrium state has vanishing residual, and whose Fr\'echet linearisation at that equilibrium is the Schr\"odinger equation with Witten Hamiltonian $\HW$.
\end{quote}

More precisely:
\begin{itemize}
    \item The full dynamics is an exact nonlinear Schr\"odinger-type equation with a state-dependent modulus-curvature term.
    \item The nonlinear term vanishes exactly when the amplitude profile matches the equilibrium profile.
    \item Near equilibrium, the dynamics reduces to:
    \begin{equation}
        i\,\partial_t\Psi = \HW\Psi = \left(-\frac{\sigma^2}{2}\Delta + \VW\right)\Psi
    \end{equation}
    \item The effective parameters ($m_{\mathrm{eff}} = 1/\sigma^2$, $V_{\mathrm{eff}} = \VW$) are determined by the closure functional $\Fcal$ and noise scale $\sigma$.
\end{itemize}

Within this framework, the Schr\"odinger equation appears not as a starting axiom, but as the \textit{distinguished equilibrium tangent sector} of a broader nonlinear dynamics.

\begin{remark}[On non-triviality of the tangent limit]
While any nonlinear system admits a linearisation at a fixed point, the non-trivial content here is that the linearised generator is explicitly the Witten Hamiltonian $\HW$, derived from the closure functional $\Fcal$ and noise scale $\sigma$, rather than an arbitrary linear operator.
\end{remark}

%==================================================================
\section{Relation to Known Formulations}\label{sec:relation}
%==================================================================

\subsection{Madelung / Bohmian mechanics}

The closure construction produces the same modulus-curvature functional $\Delta\sqrt\rho/\sqrt\rho$ that appears as the quantum potential in the Madelung formulation. The structural difference is:
\begin{itemize}
    \item In standard quantum mechanics, the quantum potential is absorbed entirely into the linear Schr\"odinger kinetic term.
    \item In the closure framework, a sign reversal (inherited from the dissipative origin of the forward process) prevents full absorption, leaving a nonlinear residual with an equilibrium subtraction.
\end{itemize}

The closure residual is therefore the non-cancelled remainder of the same object that, in linear QM, is hidden inside the Schr\"odinger kinetic structure.

\subsection{Nelson stochastic mechanics}

The pairing construction is structurally analogous to Nelson's stochastic mechanics, which also derives Schr\"odinger dynamics from forward/backward stochastic processes. The differences are:
\begin{itemize}
    \item Nelson postulates a stochastic Newton equation ($ma = -\nabla V$) that produces the correct quantum-potential sign.
    \item The closure framework derives the dynamics from the Fokker--Planck structure of a gradient-flow Langevin equation, which produces the reversed sign.
    \item The closure construction does not require postulating an acceleration law; it uses only the existing stochastic process and its time-reversal.
\end{itemize}

The price of derivation (rather than postulation) is the nonlinear residual. The benefit is that the construction uses no quantum axioms.

\subsection{Nonlinear Schr\"odinger equations}

The closure equation belongs to the broad class of modulus-curvature nonlinear Schr\"odinger equations. Within this class, the closure equation is distinguished by:
\begin{itemize}
    \item a coefficient ($\sigma^2$ in normalised form) fixed by the noise scale,
    \item an equilibrium subtraction ($2\sigma^2\VW$) fixed by the closure geometry,
    \item a distinguished linear limit selected by the equilibrium state.
\end{itemize}

No exact identification with the full Doebner--Goldin family or any specific existing model is claimed. The closure equation is a \textit{derived member} of a recognised class, not a phenomenological addition.

%==================================================================
\section{What Has Been Established}\label{sec:established}
%==================================================================

\begin{table}[ht]
\centering
\caption{Status of results across the programme.}
\label{tab:status}
\begin{tabular}{@{}llp{6cm}@{}}
\toprule
Paper & Result & Status \\
\midrule
XVI & Closure evolution equation & Derived generator-level evolution equation \\
XVII & Dissipative obstruction & Proved (Theorem~1 of XVII) \\
XVIII & Continuity equation & Exact \\
XVIII & Irrotational phase function & Exact (structural consequence) \\
XVIII & Nonlinear wave equation & Exact \\
XVIII & Near-equilibrium Schr\"odinger & Leading order in perturbation \\
XIX & Residual structural properties & Proved (phase invariance, locality, etc.) \\
XIX & Norm conservation & Exact for full nonlinear equation \\
XIX & Gauge non-removability & Proved for phase and linear-unitary transforms \\
XX & Modulus-curvature classification & Structural placement \\
XX & Equilibrium tangent limit & Proved (Fr\'echet linearisation) \\
\bottomrule
\end{tabular}
\end{table}

%==================================================================
\section{What Has Not Been Established}\label{sec:not}
%==================================================================

The following are \textbf{not} claimed by the programme:

\begin{itemize}
    \item \textbf{Full quantum mechanics.} No measurement theory, no Hilbert-space axioms, no Born rule as a derived physical law.
    \item \textbf{Exact global Schr\"odinger recovery.} The full dynamics is nonlinear; the Schr\"odinger equation holds only in the equilibrium-centred regime.
    \item \textbf{Unitarity.} The full nonlinear equation conserves norm but is not generated by a skew-adjoint operator.
    \item \textbf{Physical identification of $\sigma^2$ with $\hbar$.} The structural correspondence is established; the physical identification is not.
    \item \textbf{Experimental predictions.} No observable consequences of the nonlinear residual have been derived or bounded.
    \item \textbf{Physical status of the residual.} Whether the nonlinear correction represents genuine beyond-QM physics, an effective-theory artefact, or a gauge-removable structure remains open.
    \item \textbf{No uniqueness claim.} The programme identifies one coherent route from closure dynamics to Schr\"odinger-type structure; it does not prove that this route is unique or that the forward/backward pairing is the only viable construction.
\end{itemize}

\subsection*{Limitations of the construction}

The results establish a structural pathway from closure dynamics to Schr\"odinger-type evolution, but do not constitute a derivation of full quantum mechanics. In particular: measurement theory is not derived, Hilbert-space axioms are not introduced, the Born rule is not obtained as a physical law, and the nonlinear residual is not assigned physical interpretation. The results should therefore be understood as a mathematical construction linking dissipative dynamics to quantum structure, rather than as a complete physical theory.

%==================================================================
\section{The Logical Structure}\label{sec:logic}
%==================================================================

The programme's logical chain is:
\begin{enumerate}
    \item \textbf{Gradient-flow closure dynamics} (Papers~XII--XIV): deterministic descent, stochastic noise, Boltzmann equilibrium.
    \item \textbf{Phase complexification} (Paper~XV): imaginary potential added to path integral $\rightarrow$ interference and phase structure.
    \item \textbf{Evolution equation} (Paper~XVI): differential generator derived from the complexified path integral.
    \item \textbf{Kinetic obstruction} (Paper~XVII): single-generator unitarity is impossible $\rightarrow$ reversible kinetic sector needed.
    \item \textbf{Forward/backward pairing} (Paper~XVIII): Nelson-type construction produces imaginary kinetic term $\rightarrow$ nonlinear Schr\"odinger-type equation.
    \item \textbf{Residual classification} (Papers~XIX--XX): nonlinear remainder is isolated, classified as modulus-curvature structure with equilibrium subtraction.
    \item \textbf{Equilibrium tangent limit}: linearisation at equilibrium recovers exact Schr\"odinger equation with Witten Hamiltonian.
\end{enumerate}

Each step either proves a result, identifies an obstruction, or introduces a construction that resolves the previous obstruction. The chain is structurally constructive: each paper directly addresses the gap left by its predecessor.

%==================================================================
\section{The Reframing}\label{sec:reframing}
%==================================================================

The programme reframes the logical status of the Schr\"odinger equation within the closure framework:

\begin{quote}
The Schr\"odinger equation is not the full dynamics. It is the \textit{distinguished equilibrium tangent sector} of an exact nonlinear closure-paired dynamics.
\end{quote}

This reframing has several consequences:
\begin{itemize}
    \item Linear quantum mechanics is not contradicted: it is recovered exactly in the equilibrium-centred regime, which is the regime most directly aligned with standard linear quantum phenomenology.
    \item The nonlinear residual defines a regime beyond linear QM that may or may not be physically accessible.
    \item The Witten Hamiltonian $\HW$ --- which connects the closure functional $\Fcal$ to quantum structure via supersymmetric quantum mechanics --- is identified as the effective quantum Hamiltonian, not by postulate, but by the equilibrium tangent construction.
\end{itemize}

%==================================================================
\section{Open Questions}\label{sec:open}
%==================================================================

\begin{enumerate}
    \item Does the nonlinear residual produce observable deviations from standard QM? If so, at what energy or density scales?
    \item Can the nonlinear equation be constrained by existing experimental bounds on nonlinear quantum mechanics?
    \item Does the closure equation possess a conserved energy functional beyond the $L^2$ norm?
    \item Can the nodal divergence of the residual be regularised, or does it indicate a physical boundary condition?
    \item Is there a variational principle for the full nonlinear closure dynamics?
    \item Can the construction be extended to field-theoretic or multi-particle settings?
    \item Can the closure residual be related to existing experimentally constrained nonlinear Schr\"odinger frameworks in a quantitatively testable way?
\end{enumerate}

%==================================================================
\section{Conclusion}\label{sec:conclusion}
%==================================================================

The closure programme demonstrates a mathematically consistent pathway:
\begin{equation*}
    \text{Dissipative dynamics} \;\rightarrow\; \text{Stochastic closure} \;\rightarrow\; \text{Paired processes} \;\rightarrow\; \text{Nonlinear wave equation} \;\rightarrow\; \text{Schr\"odinger limit}
\end{equation*}

The central conclusion is:

\begin{quote}
\textit{Linear quantum evolution can arise as the equilibrium tangent limit of a broader nonlinear dynamical system, without being postulated as a fundamental axiom.}
\end{quote}

The significance of this result lies not in replacing quantum mechanics --- which remains the correct description of the experimentally established linear quantum regime --- but in identifying a mathematically coherent route by which it may emerge from a deeper dynamical structure. Whether that deeper structure has physical content beyond the linear quantum limit is the defining open question for the next stage of the programme.

%==================================================================
\bibliographystyle{plainnat}
\bibliography{references}

\end{document}
