\documentclass[11pt]{article}
\usepackage[margin=1.1in]{geometry}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{booktabs}
\usepackage[hidelinks]{hyperref}
\usepackage{microtype}

\newcommand{\KK}{\mathbb{Q}(\sqrt5)}
\newcommand{\RH}{\mathrm{RH}}
\newtheorem{prop}{Proposition}
\theoremstyle{remark}
\newtheorem*{scope}{Scope}

\title{The Prime Field Through One Instrument:\\
a Ten-Row Phenomena Ledger}
\author{VFD Programme\thanks{Code, gates, and machine-readable artifacts:
\texttt{github.com/vfd-org/closure-positivity-lab},
\texttt{lab/prime\_ledger/}, \texttt{out/prime\_ledger.json}.
Work order WO-VFD-PRIME-LEDGER-001.}}
\date{June 2026}

\begin{document}
\maketitle

\begin{abstract}\noindent
We run ten named prime-distribution phenomena through one computational
instrument --- the explicit formula read as prime waves, together with
per-place admissibility counting --- and publish the results as a ledger.
The organising observation is a factorisation: nine of the ten phenomena
split into a \emph{local layer} (an all-finite-places product: decidable,
classical, verified here numerically) and a \emph{global layer} (a statement
about the zero distribution of an explicit $L$-function, reached through the
explicit formula --- open in every conjectural case, and open in structurally
parallel ways); the tenth (Mersenne primes) is included precisely because the
factorisation \emph{fails} there, and the ledger marks its own boundary
rather than stretching past it. Headline
verifications: the twin-prime count below $5\times10^7$ matches the
Hardy--Littlewood all-places integral to $0.003\%$; eight Polignac gap ratios
match their local products to $\le 0.25\%$ (gap $210\to 3.1994$ vs $16/5$);
two hundred Goldbach counts match per-$N$ predictions to $0.65\%$ mean; the
first Chebyshev lead change reproduces Leech's $x=26{,}861$ exactly; the
maximal gap below $5\times10^7$ reproduces the record $(220,\,47{,}326{,}693)$;
and three theorem-grade anchor rows (Dirichlet, Chebotarev in $\KK$,
Sato--Tate for the icosian cuspidal $L$-function) calibrate the instrument
against proven mathematics, the last doubling as an audit of our own geometric
Brandt eigenvalues ($663$ angles, all moments within pre-set $3\sigma$ gates).
\textbf{Layer-1 results are classical --- verified, not discovered. Nothing
here proves RH, any GRH, or any infinitude conjecture.} The ledger's sole
novel claim is organisational: one local motif, one instrument, and the open
parts located --- each in the zero distribution of a named $L$-function.
\end{abstract}

\section{The thesis, and what is not claimed}

Fix the convention of \cite{Flagship,Diffraction}: a ``closure'' condition at
a finite place $p$ is a congruence-admissibility condition, and the
\emph{local product} of a pattern is the product over all finite places of its
admissible-density ratios. The ledger tests one statement:

\begin{quote}
\textbf{Factorisation thesis.} Every prime-distribution phenomenon below
(row 9, the marked boundary case, excepted) factors as (i) a local layer: an
all-finite-places product, decidable and exactly computable, which fixes
\emph{local admissibility} and the singular-series constant --- conjecturally,
the density; and (ii) a global layer: a statement about correlations
among zeros of the relevant $L$-functions, reached through the explicit
formula, which fixes \emph{infinitude and error terms} and is open in every
conjectural row.
\end{quote}

\begin{scope}
The local results verified here are classical --- Hardy--Littlewood
\cite{HL1923}, Dirichlet, Chebotarev, Barnet-Lamb--Gee--Geraghty
\cite{BLGHT} --- and the ledger claims no priority over any of them. The
global statements (twin-prime and Polignac infinitude, binary Goldbach,
Cram\'er's conjecture, full GUE, RH itself) remain open and are not claimed,
individually or in aggregate. The contribution is the uniform decomposition,
the pre-registered gates, and the single reproducible artifact.
\end{scope}

Three design rules make the ledger an instrument rather than an
advertisement. \emph{Gates are set before the data are inspected} (for
statistical rows, at $3\sigma$ under the null; for deterministic rows, on
exact literature constants). \emph{Anchor rows} --- phenomena that are
theorems --- run alongside conjectural rows on the same code path; the
instrument must reproduce proven mathematics exactly or stand indicted.
\emph{Every row carries a falsifier}, stored in the machine-readable artifact.

\section{The rows}

All computations: sieve to $X=5\times10^7$, native Python/NumPy;
$\zeta$-zeros to height $2500$ by PARI/GP; total runtime $\sim$11\,s.
One command: \texttt{python3 -m lab.prime\_ledger}.

\subsection{Multiplicative patterns (rows 1--3)}

\textbf{Row 1, twin primes.} The pattern $\{0,2\}$ is admissible at every
finite place; the local product is $2C_2$ with
$C_2=\prod_{p>2}\bigl(1-(p-1)^{-2}\bigr)$, and Hardy--Littlewood's
conjecture B predicts $\pi_2(x)\sim 2C_2\int_2^x dt/\log^2 t$. Verified:
$239{,}101$ twins counted below $5\times10^7$ against $239{,}107$ predicted
(ratio $0.99997$), with the approach visible across checkpoints
($1.0038 \to 1.0017 \to 0.99997$). \emph{Wall:} infinitude open; its
sharpest analytic connection is to the pair correlation of zeros
\cite{Goldston}; unconditional frontier: bounded gaps $\le 246$
\cite{Maynard,Polymath8b}.

\textbf{Row 2, Polignac gaps.} The local product predicts \emph{relative}
abundances: pairs at even gap $h$ occur
$\prod_{p\mid h,\,p>2}\frac{p-1}{p-2}$ times as often as twins. Verified at
eight gaps to $\le0.25\%$: gap $6\to1.9950$ (predicted $2$), gap
$10\to1.3318$ ($4/3$), gap $30\to2.6633$ ($8/3$), gap $210\to3.1994$
($16/5$). \emph{Wall:} same as row 1, for every fixed $h$.

\textbf{Row 3, Goldbach.} The same machinery in additive form: ordered
representations $N=p+q$ track
$2C_2\prod_{p\mid N,\,p>2}\frac{p-1}{p-2}\int_2^{N-2}\frac{dt}{\log t\,\log(N-t)}$,
a \emph{per-$N$} prediction. Verified on $200$ consecutive even $N$ near
$10^6$: mean deviation $0.65\%$, maximum $2.2\%$; the $3\mid N$ doubling is
conspicuous ($N=10^6{+}2$: $16{,}400$ representations vs $\sim$$8{,}300$--$10{,}800$
for its neighbours). \emph{Wall:} positivity for every even $N$ (binary
Goldbach) open; ternary case is a theorem \cite{Helfgott}.

\subsection{Equidistribution anchors (rows 4, 6, 7)}

\textbf{Row 4, primes in arithmetic progressions (theorem).} Admissible
classes mod $q$ are exactly the residues coprime to $q$ --- the closure
condition at the places dividing $q$. Verified: equidistribution deviation
$<0.04\%$ across $q\in\{4,5,12\}$; inadmissible classes exactly empty.
\emph{Wall:} the error term --- square-root cancellation is GRH for Dirichlet
$L$-functions.

\textbf{Row 6, splitting in $\KK$ (theorem; the programme's base field).}
$p$ splits iff $p\equiv\pm1\pmod 5$. Verified: split fraction $0.49993$;
and one \emph{exact} closure fact --- a twin pair $(p,p+2)$ with
$p\equiv3\pmod5$ forces $5\mid p+2$, so exactly one such pair exists,
$(3,5)$; the sieve finds exactly one. The Brandt prime ideals of
\cite{Diffraction} lie over precisely these split/inert primes.
\emph{Wall:} error term $=$ zeros of $\zeta_{\KK}$.

\textbf{Row 7, Sato--Tate for the icosian cuspidal $L$ (theorem
\cite{BLGHT}; self-audit).} The normalized Brandt eigenvalue angles of the
non-CM Hilbert modular form behind \cite{Diffraction} must equidistribute
under $(2/\pi)\sin^2\theta$. Tested on the $663$ good-prime angles (the
norm-$31$ Steinberg prime dropped; $31$ splits, and only the level ideal is
removed): moments $(0.0005,\,0.2515,\,-0.0000,\,0.1265)$ against
$(0,\,\tfrac14,\,0,\,\tfrac18)$, all inside pre-set $3\sigma$ gates;
Kolmogorov--Smirnov distance $0.034$ ($5\%$ critical value
$1.36/\sqrt{663}=0.053$). Because Sato--Tate is a \emph{theorem} for this
object, the row audits our geometric eigenvalues, not the conjecture: a
persistent deviation would indict the Brandt data. \emph{Wall:} the
\emph{rate} (effective Sato--Tate) $=$ symmetric-power $L$-functions;
$\sqrt{\phantom{n}}$-cancellation is the GRH-regime signature.

\subsection{Interference rows (rows 5, 10)}

\textbf{Row 5, Chebyshev bias.} In explicit-formula terms the bias is
interference: the prime-\emph{square} wave loads one residue class,
displacing counts by $\sim\sqrt x/\log x$ --- an asymmetric fringe on the
same instrument as the prime-wave reconstruction of \cite{Diffraction}.
Verified: first mod-$4$ lead change at exactly $x=26{,}861$ \cite{Leech};
lead fraction $99.93\%$ of primes below $5\times10^7$; mod-$3$ bias never
reverses below the sieve bound (first reversal known near $6\times10^{11}$).
\emph{Wall:} the logarithmic density $0.9959\ldots$ of the lead set is a
theorem only under GRH $+$ linear independence \cite{RubinsteinSarnak};
unconditionally, even existence of the density is open.

\textbf{Row 10, zero statistics --- the central wall, measured.} The
global object behind rows 1, 2 and 8 --- $\zeta$'s own zero statistics,
the prototype for the parallel Dirichlet and symmetric-power walls of the
other rows --- is measured directly on $1{,}983$ $\zeta$-zeros to height
$2500$: unfolded spacing mean $1.0011$; level repulsion
$P(s<0.2)=0.0025$ where Poisson statistics would give $0.181$ --- a
$70\times$ suppression, the decisive qualitative GUE signature at this
height; spacing variance $0.1502$ and pair-correlation deviation $0.094$
from $1-\mathrm{sinc}^2$, both inside the documented finite-height
correction scale $1/\log(T/2\pi)\approx0.167$ \cite{Berry,Odlyzko}.
Montgomery's theorem \cite{Montgomery} covers restricted test functions
only; the full GUE law is conjectural. This row has no deeper layer:
proving these correlations --- or RH, which they refine --- \emph{is} the
prototype of the walls the other rows point at.

\subsection{Extremes and the boundary (rows 8, 9)}

\textbf{Row 8, Cram\'er gaps.} Theorem-grade part verified: mean gap equals
$\log x$ to $\le0.12\%$ in three windows; deterministic part: the maximal
gap below $5\times10^7$ is exactly $220$, after $47{,}326{,}693$, reproducing
the record tables; Cram\'er ratio $220/\log^2(47{,}326{,}693)=0.704$. The
pure exponential model's tail is reported \emph{with} its known defect:
$P(u>2)=0.113$ vs the model's $0.135$ --- the Hardy--Littlewood correlations
of rows 1--2 are exactly the corrections the random model misses, a
cross-row consistency check. \emph{Wall:} Cram\'er's $O(\log^2 p)$ open;
RH gives only $O(\sqrt p\,\log p)$; unconditional
$\ll x^{0.525}$ \cite{BHP}.

\textbf{Row 9, Mersenne primes --- the honest boundary.} Shape-constrained
primes have \emph{no $L$-function handle}: the instrument does not apply,
and the ledger says so rather than stretching. What survives is rigid local
structure (every factor $q$ of $2^p-1$ obeys $q\equiv1\bmod 2p$ and
$q\equiv\pm1\bmod 8$ --- verified on the actual smallest factors) and a
count heuristic (Lenstra--Pomerance--Wagstaff
$\approx(e^\gamma/\log2)\log y$: predicts $20.8\pm4.6$ below $y=3300$;
observed $18$) that is \emph{not} a singular-series consequence.
Deterministic anchor: a from-scratch Lucas--Lehmer sweep reproduces the
literature's eighteen exponents exactly. \emph{Wall, doubly open:}
infinitude of Mersenne primes \emph{and} of Mersenne composites.

\section{The ledger}

\begin{center}
\small
\begin{tabular}{@{}llll@{}}
\toprule
row & phenomenon & local layer verified & global wall (open unless noted) \\
\midrule
1 & twin primes & $0.003\%$ vs HL integral & pair corr.\ of zeros \\
2 & Polignac gaps & $8$ ratios $\le0.25\%$ & same, each fixed $h$ \\
3 & Goldbach & $200$ $N$, mean $0.65\%$ & minor arcs / Dirichlet zeros \\
4 & primes in APs & $<0.04\%$; exact emptiness & error term $=$ GRH
(\emph{law: theorem}) \\
5 & Chebyshev bias & Leech's $26{,}861$ exact & density needs GRH$+$LI \\
6 & splitting in $\KK$ & $0.49993$; $(3,5)$ unique & error $=\zeta_\KK$ zeros
(\emph{law: theorem}) \\
7 & Sato--Tate (our $L$) & $663$ angles, $3\sigma$ & rate $=$ sym-power $L$s
(\emph{law: theorem}) \\
8 & Cram\'er gaps & mean exact; record $(220)$ & $O(\log^2p)$ vs
$O(\sqrt p\log p)$ \\
9 & Mersenne & factors exact; LL $18/18$ & beyond the instrument \\
10 & zero statistics & GUE at height scale & measures $\zeta$'s wall directly \\
\bottomrule
\end{tabular}
\end{center}

Three rows are theorems (4, 6, 7): they are the calibration. One row (9) is
the deliberate boundary marker. Each of the remaining six has its open part
in the zero distribution of an explicit $L$-function --- $\zeta$'s zero
correlations for rows 1, 2, 8 and 10; Dirichlet $L$-functions for rows 3
and 5; a family of parallel walls, not literally one. That parallel
structure --- checkable row by row, falsifiable row by row --- is the
ledger's entire claim.

\paragraph{Falsifier for the thesis itself.} A prime-distribution phenomenon
whose local layer cannot be written as an all-finite-places product, or whose
global layer does not reduce to zero-correlations through the explicit
formula, breaks the factorisation thesis; by the ledger's rules it would be
published as a counterexample row, not omitted.

\begin{thebibliography}{99}\small
\bibitem{HL1923} G.~H. Hardy, J.~E. Littlewood,
\emph{Some problems of `Partitio Numerorum' III}, Acta Math.\ 44 (1923).
\bibitem{Maynard} J.~Maynard, \emph{Small gaps between primes},
Ann.\ of Math.\ 181 (2015).
\bibitem{Polymath8b} D.H.J. Polymath, \emph{Variants of the Selberg
sieve, and bounded intervals containing many primes}, Res.\ Math.\
Sci.\ 1:12 (2014) (the bound 246).
\bibitem{Helfgott} H.~Helfgott, \emph{The ternary Goldbach conjecture is
true}, arXiv:1312.7748 (2013).
\bibitem{Leech} J.~Leech, \emph{Note on the distribution of prime numbers},
J.~London Math.\ Soc.\ 32 (1957).
\bibitem{RubinsteinSarnak} M.~Rubinstein, P.~Sarnak,
\emph{Chebyshev's bias}, Experiment.\ Math.\ 3 (1994).
\bibitem{BLGHT} T.~Barnet-Lamb, T.~Gee, D.~Geraghty, \emph{The Sato--Tate
conjecture for Hilbert modular forms}, J.~Amer.\ Math.\ Soc.\ 24 (2011)
411--469.
\bibitem{Montgomery} H.~Montgomery, \emph{The pair correlation of zeros of
the zeta function}, Proc.\ Sympos.\ Pure Math.\ 24 (1973).
\bibitem{Odlyzko} A.~Odlyzko, \emph{The $10^{20}$-th zero of the Riemann
zeta function and 175 million of its neighbors} (1992).
\bibitem{Berry} M.~Berry, \emph{Semiclassical formula for the number
variance of the Riemann zeros}, Nonlinearity 1 (1988).
\bibitem{Goldston} D.~Goldston, \emph{On the pair correlation conjecture
for zeros of the Riemann zeta-function}, J.~reine angew.\ Math.\ 385 (1988).
\bibitem{BHP} R.~Baker, G.~Harman, J.~Pintz, \emph{The difference between
consecutive primes II}, Proc.\ London Math.\ Soc.\ 83 (2001).
\bibitem{Flagship} VFD Programme, \emph{Total positivity as the closure
law}, companion note, this repository.
\bibitem{Diffraction} VFD Programme, \emph{Closure diffraction and the
icosian $L$-function}, this repository (2026).
\end{thebibliography}

\end{document}
