% =====================================================================
\section{Introduction}\label{sec:intro}
% =====================================================================

A response operator on a fixed graph, with no shape parameters
tuned to any dataset, that simultaneously gives (i) a structural
fit consistent with the $q^{2}$ shape of the
$b\to s\mu^{+}\mu^{-}$ angular anomaly across five public
flavour-physics datasets in their passive linear-response regime,
and (ii) eighteen preregistered substrate/neuroscience
correspondences plus six drug/sleep EEG signatures in the active
dynamical regime of a
recurrent self-model layer above the same graph, deserves a separate
preprint that names the operator, gives its construction in full,
and threads the relationship between the two empirical landings
without inheriting either's load-bearing claims. That is what this
paper does.

The operator is
\begin{equation}\label{eq:cphi_intro}
\Cph \;=\; L_M + \Ph^{-2} I,
\qquad \Ph \;=\; (1+\sqrt{5})/2,
\end{equation}
where $M$ is a closure substrate (graph, simplicial complex, or
projected coordinate) and $L_M$ is its Laplacian. The shift
$\Ph^{-2} \approx 0.382$ regularises the inverse: for self-adjoint
non-negative $L_M$ on a connected finite graph, $\Cph$ is strictly
positive definite, the smallest eigenvalue is $\Ph^{-2}$, and the
operator-norm bound is
\begin{equation}\label{eq:opnorm_intro}
\|\Cph^{-1}\| \;=\; 1/\Ph^{-2} \;=\; \Ph^{2} \;\approx\; 2.618.
\end{equation}
The continuum projection in one coordinate $x$ has a closed-form
Green's function $G(x) = (\Ph/2)\,e^{-|x|/\Ph}$ with decay scale
$\Ph$ (\S\ref{sec:definition}).

The discrete substrate used by the two empirical witnesses is
the 600-cell graph $\Rsixhundred$: $120$ unit vectors on $S^{3}$,
generated by three standard coordinate families ($8$ axis vertices,
$16$ half-integer vertices, $96$ $\Ph$-mixed vertices), connected by
short edges $\langle v, w\rangle = \Ph/2$. The choice of this
polytope is post-hoc motivated by the empirical landings
(\S\ref{sec:limitations}); the construction itself is theorem-level
rigorous. The graph has $|E|=720$ edges, uniform degree by H$_4$
transitivity (with the value $12$ from the short-edge
construction), a $9$-shell H$_3$ partition
$\{1,12,20,12,30,12,20,12,1\}$, and antipodal symmetry
$s(-v) = 8 - s(v)$. The Laplacian spectrum has nine eigenvalue
classes in $\mathbb{Z}[\Ph]$ with multiplicities summing to $120$.
All of these facts are reproduced numerically by
\texttt{repro/verify\_kernel.py} from the canonical generators
alone — no external data input.

\subsection*{What this paper claims}

We claim a single \emph{operator witness}: that one geometry-fixed
operator, on one fixed graph, with no shape-parameter retuning
between regimes, appears as the shared response primitive
underneath two
empirical works covering qualitatively distinct physical settings.

\begin{enumerate}\itemsep=2pt
\item \textbf{Operator definition is fixed by the construction.}
  Once $\Rsixhundred$ is selected and the stability shift
  $\Ph^{-2}$ is chosen, $\Cph$ is fully determined. No shape
  parameter, no fitted threshold, no learned weight enters the
  operator. The Laplacian spectrum, the operator-norm bound, and
  the discrete-to-continuum agreement are computed (not fitted)
  from the construction and reproduced in
  \texttt{repro/verify\_kernel.py}.
\item \textbf{Discrete-to-continuum agreement is empirical, not
  postulated.} For a localised source at any vertex, the discrete
  response $\psi = \Cph^{-1} f$ is shell-constant to machine
  precision, and the multiplicity-weighted per-vertex Pearson
  correlation between $\psi(v)$ and $G(\|v - v_{\mathrm{src}}\|)$
  is $\rho = 0.976$ on the unweighted graph Laplacian. (This is a
  multiplicity-weighted shell-radius correlation, not $119$
  independent radial samples; the within-shell variance is at
  machine precision; see \S\ref{ssec:result_unweighted}.) It is
  numerical agreement between two independently-defined objects
  (a 120-dimensional discrete inverse and a continuum exponential
  kernel), not a definitional identity.
\item \textbf{Variant comparison among the three tested variants.}
  Two $\Ph$-cocycle weighted Laplacian variants ($\Ph$-geometric,
  $\Ph$-arithmetic edge weights via shell-grade exponents
  $\omega_{+}(v) = \Ph^{\kappa(v)}$) score lower per-vertex
  correlation: $0.888$ and $0.884$ respectively. Within the three
  tested variants, the unweighted Laplacian ranks highest on the
  geometry-only criterion. This reproduces,
  on a different test, the qualitative ranking established
  independently by the b-anomaly paper's data-$\chi^{2}$ comparison
  (\S\ref{sec:passive_witness}).
\item \textbf{Two domain-disjoint empirical landings, same operator.}
  (a)~The b-anomaly preprint~\citep{SmartBAnomaly2026} uses the
  same fixed $\Cph$ on the same $\Rsixhundred$ to describe the
  $q^{2}$ shape of the $b\to s\mu^{+}\mu^{-}$ anomaly across five
  public datasets, with one fitted dimensionless amplitude per
  dataset and the kernel held fixed; sign uniformity holds in
  $5/5$ datasets ($A>0$, $\Delta C_{9}^{\mathrm{eff}} < 0$).
  (b)~The aria-chess preprint~\citep{SmartAriaChess2026} uses the
  same fixed $\Cph$ on the same $\Rsixhundred$, augmented by a
  recurrent self-model layer above the substrate, to pass eighteen
  preregistered substrate/neuroscience correspondences (frozen
  2026-04-18) plus six drug/sleep EEG signatures.
\end{enumerate}

\subsection*{What this paper does \emph{not} claim}

\begin{itemize}\itemsep=2pt
\item \emph{Not a derivation of the $\Ph^{-2}$ floor.} The shift
  $\Ph^{-2}$ is a design-level stability clamp that bounds
  $\|\Cph^{-1}\|$ at $\Ph^{2}$. It is not derived from a closure
  functional or a symmetry argument; we report its role as a
  regularisation-of-mass scale.
\item \emph{Not a uniqueness claim for $\Rsixhundred$.} Other
  regular 4-polytopes (the $24$-cell, the $120$-cell), other
  highly symmetric graphs, and continuum substrates are all
  candidate $M$ for $\Cph = L_M + \Ph^{-2} I$. The 600-cell choice
  is post-hoc motivated by the empirical landings; a formal
  ablation against alternative substrates is an open build.
\item \emph{Not a kernel-uniqueness claim on either empirical
  landing.} The b-anomaly's free-width Gaussian alternative shows
  that a free-width Gaussian charm-loop proxy fits the same five
  datasets comparably in $\chi^{2}$ at the cost of one extra shape
  parameter; the b-anomaly AIC comparison against
  $\mathrm{FREE\_C9}$ (a constant Wilson-coefficient shift) is a
  statistical tie on current data
  ($w_{\mathrm{VFD}}=0.348$ vs $w_{\mathrm{FREE\_C9}}=0.652$).
  Both caveats are inherited verbatim from the b-anomaly preprint.
\item \emph{Not a selection theorem on the
  ACT 4-tuple.} The companion adaptive-closure-transport
  preprint~\citep{SmartAdaptiveClosureTransport2026} proposes a
  selection layer $(M, L_M, W, R_{\mathrm{hom}})$ in which $\Cph$
  fills the response slot. Selection — Lyapunov $V(W)$ on the
  reduced flow, edge-space decomposition under $2I$-equivariance
  (where $2I$ is the binary icosahedral group, the $\mathrm{SU}(2)$
  preimage of the icosahedral symmetry of $\Rsixhundred$),
  full reduced-flow convergence — is left open in that paper and
  is not delivered here.
\item \emph{Not a circuit-level model on the active-regime side.}
  The aria-chess preprint operates at the architectural-algorithmic
  level above $\Cph$. We import its empirical results verbatim and
  do not relitigate them here; their substrate-witness scope
  applies.
\end{itemize}

\subsection*{Mapping from numerics to admissible claims}

To keep this paper inside the operator-witness scope, we use the
same claim-boundary discipline as the aria-chess
preprint~\citep{SmartAriaChess2026}: numerical results
$\mathcal R_{\mathrm{numeric}}$ are mapped to admissible claims
$\mathcal C_{\mathrm{admissible}}$ by the rule
\[
q\colon \mathcal R_{\mathrm{numeric}} \longrightarrow
        \mathcal C_{\mathrm{admissible}},
\qquad
\mathcal C_{\mathrm{admissible}}
   \;=\;\{\text{`computed', `consistent with', `direction confirmed'}\}.
\]
A numerically computed quantity (the Laplacian spectrum, the
operator-norm bound, the per-vertex correlation $0.976$) licenses
a `computed'-type claim. Downstream empirical agreement using the
fixed operator (sign uniformity in $5/5$ b-anomaly datasets, the
aria-chess $17/18$ standard plus $18/18$ after the documented
$N\!=\!20$ P4 deep-dive) licenses a `consistent
with'-type claim. We never write
`derives the kernel', `proves uniqueness', or `establishes
selection'.

\medskip
\begin{center}
\fbox{\parbox{0.92\linewidth}{\small
\textbf{What is claimed / what is not.}\par
\emph{Claimed:} a geometry-fixed response operator $\Cph$ on the
600-cell graph, with computed spectrum, operator-norm bound, and
discrete-to-continuum correlation; the same fixed operator appearing
in two domain-disjoint empirical works (b-anomaly five-dataset
structural fit, AIC tie + Mode-B drift caveats inherited;
aria-chess $17/18$ preregistered correspondences under the
standard validation protocol and $18/18$ after the documented P4
$N\!=\!20$ deep-dive (thresholds unchanged) with the documented
P3/P13 methodology refinements and the open prereg-exact P10
$20$-permutation rerun inherited verbatim, plus $6/6$
literature-thresholded companion EEG signatures) without
shape-parameter retuning between regimes.\par
\emph{Not claimed:} derivation of $\Ph^{-2}$; uniqueness of
$\Rsixhundred$; uniqueness of the kernel shape on either empirical
landing; a selection theorem on the ACT 4-tuple; that either
empirical landing settles the underlying physics (flavour anomaly
or consciousness) by the operator alone.
}}
\end{center}

\subsection*{Layout}

\S\ref{sec:definition} gives the operator definition, the positivity
properties of $\Cph$, the operator-norm bound, and the continuum
projection. \S\ref{sec:substrate} constructs $\Rsixhundred$ from
canonical generators, gives the graph facts, and the $9$-shell
decomposition. \S\ref{sec:spectrum} reports the Laplacian spectrum
in $\mathbb{Z}[\Ph]$ with multiplicities, computed numerically.
\S\ref{sec:agreement} runs the discrete-to-continuum agreement test
across three Laplacian variants. \S\ref{sec:passive_witness} and
\S\ref{sec:active_witness} thread the two domain-disjoint empirical
witnesses (b-anomaly and aria-chess) at the operator level.
\S\ref{sec:programme_home} positions $\Cph$ relative to the
published adaptive-closure-transport companion as a member of the
same polynomial-in-$L$ family, and identifies the open selection
layer. \S\ref{sec:limitations} enumerates limitations in a
five-move guard matrix. \S\ref{sec:conclusion} concludes.
