% =====================================================================
\section{Active-regime witness: aria-chess}\label{sec:active_witness}
% =====================================================================

This section threads the second of the two domain-disjoint empirical landings of
$\Cph$. The full preprint is~\citep{SmartAriaChess2026}; we
summarise here only what the operator-witness narrative requires
and inherit the preprint's substrate-witness scope verbatim.

\subsection{What aria-chess tests}\label{ssec:aria_setup}

The aria-chess preprint adds a recurrent self-model layer above
the same $\Cph$ on the same $\Rsixhundred$. The architecture
introduces:
\begin{itemize}\itemsep=2pt
\item One \emph{condition-dependent} self-injection coupling
  $\eta \in \{0, 0.05, 0.20\}$ (PROPOFOL, SLEEP\_N3,
  WAKE/RECOVERY) that controls the strength of the recurrent
  feedback;
\item One \emph{substrate-pinned} nonlinearity
  $\mathrm{bounded\_topk}(\cdot, k=12)$ at the graph's average
  degree (\S\ref{ssec:graph}: degree $12$ uniform). The choice
  $k=12$ is not a free hyperparameter; it is the substrate's
  average degree.
\item Condition-specific \emph{biologically-motivated} stimulus
  models (slow oscillation + spindles + K-complexes for SLEEP\_N3,
  AR(1) noise + tonic shell + attention episodes for WAKE,
  low-amplitude tonic noise for PROPOFOL). These are
  biologically-motivated design choices, not measurement-fits to
  subject-level EEG data.
\end{itemize}
The kernel parameter $\Ph^{-2}$ is \emph{not retuned} between
b-anomaly and aria-chess; the same fixed shift used in the
flavour-physics structural fit is used in the cortical substrate
witness.

\subsection{Eighteen preregistered correspondences}

Eighteen quantitative predictions (P1--P18) were locked on
2026-04-18 in the aria-chess preprint's
\texttt{docs/brain\_mapping/PAPER\_PREDICTIONS.md} before any
validation run. Each prediction has a specific numerical claim, a
falsifiable threshold, and a named validation script. The
preregistered tally as reported in~\citep{SmartAriaChess2026}:

\begin{itemize}\itemsep=2pt
\item $17/18$ at standard validation methodology ($5$-seed
  cascade block plus state-reset protocol);
\item $18/18$ after a documented $N\!=\!20$ deep-dive on the
  residual high-variance interaction $C\!\times\!P$ (P4: bootstrap
  point estimate $+0.190$, $95\%$ CI $[+0.143, +0.239]$,
  $0/2000$ resamples at-or-below zero, reported as $0.0000$).
\item No preregistered threshold has been modified.
\end{itemize}
The aria-chess preprint reports this as methodology refinement
(documented seed-count increase on a high-per-seed-variance
interaction term), \emph{not} as a threshold change. Two further
predictions in the aria-chess set carry their own documented
methodology refinements that we also inherit: P3
($D\!\times\!C$ independence) closes at $N\!=\!5$ within
its preregistered band, and P13 (chess substrate lift) was
preregistered with $5$-fold cross-validation but the
landed value $+40.6$pp uses the documented leave-one-out /
state-reset protocol; both refinements are reported
in~\citep{SmartAriaChess2026} with no threshold modification, and
we inherit those caveats verbatim. The aria-chess preprint also
flags that the 2026-04-29 P10 (chess null mapping) validation used
$15$ permutations rather than the preregistered $20$ (threshold
$\geq 50\%$ unchanged; observed $65.4\%$ robust in the $15$--$20$
range; the prereg-exact $20$-permutation rerun is left open in the
aria-chess limitations); we inherit this caveat verbatim alongside
the $18/18$ summary.

\subsection{Six drug/sleep EEG signatures}

On a single deterministic substrate trajectory at seed~$42$, the
aria-chess preprint reports six biological signatures testing
against literature-derived thresholds:

\begin{itemize}\itemsep=2pt
\item \textbf{Wake cortical-avalanche $\alpha$}: $\alpha = 2.252$,
  $95\%$ CI $[1.82, 2.86]$, $R^{2}=0.956$ — the WAKE confidence
  interval overlaps both the Sleep-EDFx EEG CI $[2.50, 2.53]$
  ($n=30$ subjects) and aria-chess's prior cascade pipeline CI
  $[2.73, 3.25]$ pairwise (the Sleep-EDFx and prior-pipeline
  intervals do not overlap each other; the WAKE interval is the
  pairwise common ground).
\item \textbf{NREM-N3 phenomenal-intensity variance ratio}:
  $0.463\!\times$ wake (predicted $\sim 0.365$, threshold $<0.70$).
\item \textbf{Propofol modality-switching ratio}: $1.83\!\times$
  wake (threshold $\in [1.5, 5.0]$, empirical reference
  $2.96\!\times$ from OpenNeuro \texttt{ds005620}).
\item \textbf{Propofol continuity drop}: $+0.066$
  (threshold $> 0.020$).
\item \textbf{Propofol $\Phi$ collapse}: $0.33\!\times$ wake (IIT
  direction confirmed; $\Phi$-proxy not full IIT).
\item \textbf{Recovery deterministic identity to wake}: under the
  WAKE stimulus protocol, the RECOVERY trajectory is bit-identical
  to the WAKE trajectory.
\end{itemize}

\subsection{Cross-domain selectivity}

\begin{itemize}\itemsep=2pt
\item \textbf{Chess pattern recognition (P9--P13)}: $32$ chess
  positions across $4$ categories on $8$-D V2 features; under the
  disclosed leave-one-out / state-reset refinement of the
  preregistered $\geq +15$pp substrate-lift test (the original
  prereg used $5$-fold cross-validation; the LOO/state-reset
  protocol is documented in the aria-chess
  preprint~\citep{SmartAriaChess2026}), substrate routing lifts
  classification at canonical depth $n=25$ ticks from raw
  $53.1\%$ to substrate-routed $93.8\%$ ($+40.6$pp), above the
  preregistered $\geq +15$pp floor.
\item \textbf{Conversation pattern recognition (P14--P16)}:
  $64$ utterances, $8$ categories; raw classification $87.5\%$,
  substrate-routed lift $-4.4$pp (within the preregistered
  neutrality band $|\cdot| < 10$pp). The substrate is selectively
  amplifying in tasks where raw features are ambiguous and
  approximately neutral when raw features are already
  discriminative.
\item \textbf{HCP brain functional connectivity (P17--P18)}:
  full-cohort descriptive statistics on $n=1003$ subjects show
  ARIA's $H_4$-transitive structure at $-11.58\sigma$ on degree
  homogeneity, $+79.78\sigma$ on raw participation ratio (with
  node-count caveat: ARIA $|V|=120$ vs HCP ICA-50 $|V|=50$), and
  $+6.80\sigma$ on clustering coefficient. ARIA's degree std is
  $0$ by H$_4$ transitivity (a theorem), $11.58$ standard
  deviations below the HCP biological cohort.
\end{itemize}

\subsection{Reading at the operator level}

The aria-chess result is the \emph{active-regime} empirical
witness for $\Cph$ on $\Rsixhundred$. The recurrent self-model
layer above $\Cph$ uses one condition-dependent coupling and one
substrate-pinned nonlinearity at the graph's average degree
$k=12$; no other \emph{kernel-shape} parameter enters. Above the
operator, aria-chess inherits its own dynamical and stimulus
constants (e.g.\ a fixed dynamical decay, fixed cascade gains,
condition-specific biologically-motivated stimulus models); these
are documented in the aria-chess preprint and are not retuned in
this paper. The kernel shift $\Ph^{-2}$ is not retuned between
b-anomaly and aria-chess. Under those design choices, the same
fixed $\Cph$ on the same $\Rsixhundred$ is consistent with the
aria-chess active-regime tally (frozen 2026-04-18): $17/18$
preregistered cortical correspondences under the standard
validation protocol, $18/18$ after the documented $N\!=\!20$ P4
deep-dive (thresholds unchanged), plus six literature-thresholded
EEG drug/sleep signatures.

The aria-chess preprint stays inside substrate-witness scope: it
does not claim the substrate \emph{is} consciousness, does not
claim 600-cell uniqueness among regular 4-polytopes, and does not
deliver a selection theorem on the ACT 4-tuple. We inherit the
scope verbatim. What we add at the operator level is the
observation that the same fixed $\Cph$ — under no shape-parameter
retuning between regimes — is the shared response primitive
underneath both empirical landings (the b-anomaly $q^{2}$
projection above $\Cph$ and the aria-chess recurrent self-model
layer above $\Cph$ are distinct above-operator constructions; the
operator below them is the same).

\subsection{Two-witness structure}

\begin{table}[ht]
\centering
\small
\caption{Two domain-disjoint empirical landings of $\Cph$ on
$\Rsixhundred$, with no shape retuning between regimes.}
\label{tab:two_witness}
\begin{tabular}{p{0.22\linewidth} p{0.36\linewidth} p{0.36\linewidth}}
\toprule
& Passive regime & Active regime \\
\midrule
Preprint           & b-anomaly~\citep{SmartBAnomaly2026} & aria-chess~\citep{SmartAriaChess2026} \\
Domain             & flavour physics                    & cortical neuroscience \\
Datasets           & 5 public ($n_{\mathrm{tot}}=138$ bins) & prereg + EEG (HCP $n=1003$, Sleep-EDFx $n=30$, etc.) \\
Operator           & same fixed $\Cph$                  & same fixed $\Cph$ \\
Substrate          & same $\Rsixhundred$                & same $\Rsixhundred$ \\
Shift $\Ph^{-2}$   & not retuned                        & not retuned \\
Headline           & $5/5$ sign uniformity              & $17/18$ standard, $18/18$ after P4 $N\!=\!20$ deep-dive (P3 $N\!=\!5$ + P13 LOO/state-reset documented refinements; P10 prereg-exact $20$-perm rerun open); $6/6$ EEG \\
Kernel-shape free params & 0                            & 0 \\
Other free params  & 1 amplitude $A$ per dataset        & 1 coupling $\eta$ per condition; v4 stimulus protocols \\
Caveat             & AIC tie; free-width Gaussian alt.; Mode-B linearised-to-non-linear refit drift & single-seed EEG; no polytope ablation; P10 prereg-exact rerun open \\
\bottomrule
\end{tabular}
\end{table}

The two witnesses share, by design, exactly the geometry-fixed
operator: the same $\Cph$, the same substrate $\Rsixhundred$, and
the same shift $\Ph^{-2}$. They share no fitted parameter,
threshold, dataset, or methodological choice above the operator
level. The b-anomaly amplitude $A$ is fitted to flavour-physics
$q^{2}$ shape, dataset by dataset; the aria-chess coupling $\eta$
is fixed by experimental condition (PROPOFOL/SLEEP/WAKE), not by
neural data; the v4 stimulus protocols are biologically motivated,
not subject-fit. Independence in this paper is independence at
the empirical layer above the shared operator.
