% =====================================================================
\section{Limitations and hostile-review guard matrix}\label{sec:limitations}
% =====================================================================

This section enumerates limitations transparently, organised as a
five-move guard matrix following the b-anomaly preprint
template~\citep{SmartBAnomaly2026}: regime, post-hoc, interpretation,
test/claim, state-drift. For each guard we record
$G\colon \mathrm{risk} \to (\mathrm{disclosure}, \mathrm{evidence},
\mathrm{strengthening\ build})$.

\subsection{Regime}\label{ssec:regime}

\textbf{Single substrate (the 600-cell).} We have not tested
whether $\Cph$ on the $24$-cell, the $120$-cell, or other H$_n$
Coxeter graphs would give comparable per-vertex correlations on
the discrete-to-continuum agreement test, or comparable structural
fits on either empirical landing. The 600-cell choice is post-hoc
motivated by the empirical landings, not from an a-priori
derivation. \emph{Disclosure:} \S\ref{sec:intro},
\S\ref{sec:substrate}, \S\ref{sec:programme_home}.
\emph{Evidence:} per-vertex correlation $0.976$ on $\Rsixhundred$;
empirical landings of \S\ref{sec:passive_witness} and
\S\ref{sec:active_witness}. \emph{Strengthening build:}
\texttt{repro/verify\_kernel.py} extension to the $24$-cell and
$120$-cell, with the same per-vertex correlation criterion
applied; verbatim re-run of the b-anomaly fit on alternative
substrates from~\citep{SmartBAnomaly2026}; the aria-chess
preprint's regime section already lists alternative-polytope
ablation as an open build.

\textbf{Single shift ($\Ph^{-2}$).} We have not tested whether
nearby shifts ($\Ph^{-2} \pm \epsilon$ for small $\epsilon$) give
comparable per-vertex correlation, or whether the empirical
landings tolerate a shift sweep. The shift is held fixed across
both regimes; perturbation analysis is an open build.
\emph{Strengthening build:} sweep $\Ph^{-2} \cdot (1 \pm \delta)$
for $\delta \in \{0.01, 0.05, 0.10, 0.25\}$ on the discrete-to-
continuum correlation; report sensitivity envelope.

\subsection{Post-hoc}\label{ssec:posthoc}

\textbf{The 600-cell choice is post-hoc justified by empirical
observables.} While the construction of $\Rsixhundred$ is
theorem-level rigorous (H$_4$ Coxeter group theory), the choice
of \emph{this} polytope as the discrete substrate instance is
motivated by the empirical landings observed, not by an a-priori
geometric or algebraic argument selecting it over the $24$-cell
or $120$-cell. \emph{Disclosure:} \S\ref{sec:intro}.
\emph{Evidence:} two domain-disjoint empirical witnesses on
$\Rsixhundred$. \emph{Strengthening build:} formal substrate
ablation (above).

\textbf{The geometry-first variant agreement is not historically
blind on b-anomaly.} Per the b-anomaly preprint's limitations
section, the LHCb 2025 data was inspected first and the
pure-geometry ranking was verified afterward to agree with the
data-$\chi^{2}$ ranking. The two-criterion agreement is
\emph{criterion-independent} (geometry-only correlation here is a
different test from b-anomaly's data $\chi^{2}$) but not
historically pre-registered. \emph{Disclosure:} we inherit the
caveat verbatim. \emph{Strengthening build:} a future blind variant
test would freeze the variant choice before observing the data
$\chi^{2}$.

\textbf{$\Ph^{-2}$ floor not derived.} The shifted-Laplacian floor
$\Ph^{-2} \approx 0.382$ is a stability clamp making $\Cph$
strictly positive definite; it is not derived from a closure
functional or symmetry argument. \emph{Disclosure:}
\S\ref{ssec:opnorm}, \S\ref{sec:definition}. \emph{Evidence:} the
same operator with the same shift serves as the basis for two
domain-disjoint empirical witnesses across qualitatively distinct
regimes (\S\ref{sec:passive_witness},
\S\ref{sec:active_witness}). \emph{Strengthening build:} derive
or justify a distinguished shift under a named criterion (e.g.\
minimum-amplitude amplification on a specified function class);
uniqueness is not assumed and is itself an open problem.

\subsection{Interpretation}\label{ssec:interpretation}

\textbf{The discrete-to-continuum agreement is descriptive, not
causal.} The per-vertex correlation $0.976$ between $\psi$ on
$\Rsixhundred$ and the continuum kernel
$G(x) = (\Ph/2)\,e^{-|x|/\Ph}$ at the same chord radii is a
\emph{computed agreement} between two independently-defined
objects, not a derivation that the discrete operator equals the
continuum kernel. \emph{Disclosure:} \S\ref{sec:agreement} marks
this explicitly. \emph{Evidence:} the agreement is at machine
precision in the operator-norm bound and at $\rho = 0.976$ in the
per-vertex correlation. \emph{Strengthening build:} a formal
discrete-continuum convergence proof under a specified large-graph
limit; a continuum limit theorem on H$_n$ Coxeter substrates as
$n \to \infty$.

\textbf{Variant ranking is criterion-dependent.} The unweighted
variant wins on both the geometry-only criterion of this paper
and b-anomaly's data $\chi^{2}$, but neither criterion is the
\emph{unique} natural ranking. Edge-weighted variants outside the
$\Ph$-cocycle family ($\sqrt{\deg}$-weighted, normalised
Laplacian) are not tested here. \emph{Strengthening build:}
extend \texttt{repro/verify\_kernel.py} to a wider variant
catalogue.

\subsection{Test/claim}\label{ssec:testclaim}

\textbf{Two domain-disjoint empirical landings, not formal physics.}
The b-anomaly result is a structural fit (kernel shape held fixed)
under an AIC tie with $\mathrm{FREE\_C9}$ on current data; the
aria-chess result is a substrate witness (no claim that the
substrate \emph{is} consciousness). Neither lands a theorem-grade
physics claim on its own domain; both are appropriately
hedged in their own preprints, and we inherit those hedges
verbatim. \emph{Disclosure:} \S\ref{sec:passive_witness},
\S\ref{sec:active_witness}. \emph{Evidence:} the witnesses pass
their own preregistered or literature-derived thresholds.
\emph{Strengthening build:} more flavour-physics datasets
(LHCb Run~3, Belle~II) for the passive-regime witness;
cross-cohort EEG (TUH, NHM) and cross-parcellation HCP
(Schaefer, Glasser) for the active-regime witness; both already
listed in the respective preprints.

\textbf{Discrete-to-continuum correlation reported on a single pole,
verified across all $|V|$ on the unweighted operator.} The
headline per-vertex correlation $0.976$ is reported with the
canonical pole ($+x_{0}$ axis) as source on the unweighted
Laplacian. H$_4$ transitivity predicts the correlation is
invariant under choice of source vertex. The
\texttt{multi\_source\_sweep} routine in
\texttt{repro/verify\_kernel.py} runs over all $120$ vertices on
the unweighted operator only (the two $\Ph$-cocycle weighted
variants are not swept here, an explicit ablation build); every
unweighted source returns the same per-vertex correlation
$0.976202$ to within $\sim 10^{-15}$
(\texttt{multi\_source\_sweep.max\_minus\_min} in
\texttt{results.json}). The single-pole headline is therefore a
representative, not a sample, of the unweighted operator's
behaviour under H$_4$ transitivity; the two $\Ph$-cocycle weighted
variants are not covered by this sweep.

\subsection{State-drift / out-of-scope}\label{ssec:scope}

\textbf{Selection layer not delivered.} As enumerated in
\S\ref{sec:programme_home}, the selection-layer constructions
(Lyapunov $V(W)$, edge-space decomposition under $2I$-equivariance,
full reduced-flow convergence) are open in the companion ACT
preprint~\citep{SmartAdaptiveClosureTransport2026} and are not
delivered here. The two empirical witnesses strengthen confidence
in the response primitive without addressing the selection gap.

\textbf{Broader cascade programme is out of scope.} Other
preprints in the broader cascade programme share operator-level
infrastructure with this paper but are not load-bearing here.
Deliberately out of scope.

\textbf{Out-of-scope items NOT delivered (correctly).}
\begin{itemize}\itemsep=2pt
\item Lyapunov function $V(W)$ on the reduced flow — open build
  of the ACT companion paper~\citep{SmartAdaptiveClosureTransport2026}.
\item $2I$-equivariance audit of the closure operator family —
  open build of ACT.
\item Edge-space decomposition of $\mathbb{R}^{E_{M}}$ under the
  Hodge edge Laplacian — open build of ACT.
\item Selection theorem identifying $\Rsixhundred$ over alternative
  regular 4-polytopes — see \S\ref{ssec:regime}.
\item Formal discrete-to-continuum convergence theorem under a
  specified large-graph limit — see \S\ref{ssec:interpretation}.
\item First-principles derivation of the $\Ph^{-2}$ shift — see
  \S\ref{ssec:posthoc}.
\item Family-membership classification theorem for the
  polynomial-in-$L$ Lyapunov family — see \S\ref{sec:programme_home}.
\end{itemize}

\subsection{The honest verdict}

The result is an \emph{operator witness}: a geometry-fixed response
operator on a fixed graph, with no shape parameters tuned to any
dataset, is consistent with two domain-disjoint empirical landings in
qualitatively distinct regimes. We do not claim the operator is
the unique kernel for either landing. We do not claim selection is
delivered. We do not claim 600-cell uniqueness. The strengthening
builds for stronger claims are listed above and remain open.
