% =====================================================================
\section{Conclusion}\label{sec:conclusion}
% =====================================================================

The closure-response operator $\Cph = L_M + \Ph^{-2} I$ on the
600-cell graph $\Rsixhundred$, with $\Ph = (1+\sqrt{5})/2$, is a
geometry-fixed response primitive: positive definite under
(H1)--(H3) on the substrate $(M, L_M)$, and on the connected
finite graph $\Rsixhundred$ where $\inf \sigma(L_M) = 0$ it has
smallest eigenvalue $\Ph^{-2}$ and operator-norm identity
$\|\Cph^{-1}\| = \Ph^{2} \approx 2.618$ (general substrates with
$\inf \sigma(L_M) > 0$ give the strict inequality
$\|\Cph^{-1}\| < \Ph^{2}$).
The 600-cell instance has $|V|=120$, $|E|=720$, uniform degree
$12$, $9$-shell partition $\{1,12,20,12,30,12,20,12,1\}$, and a
Laplacian spectrum that numerically resolves to the closed-form
$\mathbb{Z}[\Ph]$ values listed in Table~\ref{tab:spectrum}. The
discrete-to-continuum agreement between $\psi = \Cph^{-1} f$ and
the continuum kernel $G(x) = (\Ph/2)\,e^{-|x|/\Ph}$ at per-vertex
chord distances (non-source vertices) is Pearson $\rho = 0.976$ on
the unweighted Laplacian, above the two $\Ph$-cocycle weighted
variants tested ($0.888$ geometric, $0.884$ arithmetic). All numbers are reproduced from canonical
generators by \texttt{repro/verify\_kernel.py}; no parameter is
fitted.

\textbf{Two domain-disjoint empirical landings.} The same fixed
$\Cph$ on the same fixed $\Rsixhundred$, with no shape-parameter
retuning between regimes, appears as the shared response
primitive underneath:
\begin{enumerate}\itemsep=2pt
\item \textbf{Passive regime}~\citep{SmartBAnomaly2026}: a single
  fitted dimensionless amplitude $A$ per dataset (kernel shape
  held fixed) gives a sign-uniform
  $\Delta C_{9}^{\mathrm{eff}} < 0$ description of the
  $b\to s\mu^{+}\mu^{-}$ angular anomaly across five public
  datasets (LHCb 2015, LHCb 2021, CMS 2025, LHCb 2025, LHCb 2015
  $B_s\to\phi$).
\item \textbf{Active regime}~\citep{SmartAriaChess2026}: a
  recurrent self-model layer above the same operator (one
  condition-dependent self-injection coupling
  $\eta\in\{0,0.05,0.20\}$, one substrate-pinned nonlinearity
  $\mathrm{bounded\_topk}(\cdot, k=12)$ at the graph's average
  degree) is consistent with the aria-chess active-regime tally
  (frozen 2026-04-18): $17/18$ preregistered
  substrate/neuroscience correspondences under the standard
  validation protocol, $18/18$ after the documented $N\!=\!20$
  P4 deep-dive (thresholds unchanged; with the documented P3
  $N\!=\!5$ and P13 LOO/state-reset refinements, and the open
  prereg-exact P10 $20$-permutation rerun, all inherited verbatim
  from~\citep{SmartAriaChess2026}), plus six
  literature-thresholded drug/sleep EEG signatures.
\end{enumerate}
By design, the two witnesses share exactly the geometry-fixed
operator: the same $\Cph$, substrate $\Rsixhundred$, and shift
$\Ph^{-2}$. Above that operator level, they share no fitted
parameter, threshold, or dataset, and the empirical claims are
tested on disjoint physical domains (flavour physics vs cortical
neuroscience). Independence here is
independence at the empirical layer above the shared operator,
not statistical independence of the operator itself.

\textbf{Operator-witness scope.} This is an operator witness, not
a derivation of physics on either landing. We do not derive the
$\Ph^{-2}$ shift; it is a design-level stability clamp. We do not
claim 600-cell uniqueness; alternative regular 4-polytopes are an
explicit ablation build. We do not claim kernel uniqueness on
either empirical landing; the b-anomaly's free-width Gaussian
alternative, AIC tie ($w_{\mathrm{VFD}}=0.348$ vs
$w_{\mathrm{FREE\_C9}}=0.652$), and Mode-B linearised-to-non-
linear refit drift caveats are inherited verbatim, and the
aria-chess substrate-witness scope is inherited verbatim. We do
not deliver a selection theorem on the ACT
4-tuple~\citep{SmartAdaptiveClosureTransport2026}; that paper's
open builds (Lyapunov $V(W)$, edge-space decomposition under
$2I$-equivariance, full reduced-flow convergence) remain open and
are not delivered here.

\textbf{Programme position.} \S\ref{sec:programme_home} positions
$\Cph$ relative to the published adaptive-closure-transport
companion as a member of the same polynomial-in-$L$ family;
family-membership is \emph{programme-positioned}, not formally
classified. The two empirical landings provide external
consistency checks on the \emph{response} primitive without
reducing or addressing the selection gap.

The empirical material gathered here is the operator witness; the
broader programme to convert the witness into a selection-theorem-
grade claim is sketched in the companion preprints and remains
the natural next step.
