\documentclass[12pt]{article}

\usepackage[a4paper, margin=2.5cm]{geometry}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage[colorlinks=true, linkcolor=blue, citecolor=blue, urlcolor=blue]{hyperref}
\usepackage{booktabs}
\usepackage{natbib}

\newtheorem{theorem}{Theorem}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{example}[theorem]{Example}
\newtheorem{lemma}[theorem]{Lemma}

\newcommand{\Ecal}{\mathcal{E}}
\newcommand{\Ocal}{\mathcal{O}}

\title{Observer Frames and Relativity\\from Event-Order Geometry}
\author{%
  Lee Smart\\[4pt]
  \textit{Institute of Vibrational Field Dynamics}\\[2pt]
  \texttt{contact@vibrationalfielddynamics.org}\\[2pt]
  \texttt{@vfd\_org}%
}
\date{April 2026}

\begin{document}

\maketitle

\noindent\textbf{Status:} Preprint (not peer-reviewed)

\begin{abstract}
Paper~XXIII derived a strict partial order on admissible overlap events between two phase-rotated 600-cell copies, interpreting the ordering as an emergent temporal structure. The present paper shows that this event poset carries observer-dependent structure analogous to the kinematic backbone of special relativity.

An observer is defined as a linear extension of the event partial order equipped with a sampling measure over events. Different observers assign different time functions to the same event set. The paper proves: (1) relativity of simultaneity --- incomparable events can be ordered differently by different observers; (2) that differing observer measures yield non-unique elapsed-time assignments along shared event chains, providing a time-dilation analogue; (3) worldlines arise as maximal chains in the poset. A worked example on the dual-pentagon system demonstrates all constructions explicitly.

The paper does not derive Lorentz invariance, the metric tensor, or the Einstein equations. It establishes that the kinematic phenomena of frame-dependent time and simultaneity emerge from the event-order geometry without assuming spacetime.
\end{abstract}

%==================================================================
\section{Introduction}\label{sec:intro}
%==================================================================

Paper~XXIII~\citep{SmartXXIII} established that the phase-overlap geometry of two conjugate 600-cell copies produces a strict partial order $\prec$ on an event set $\Ecal$, with no canonical global clock. The present paper converts this mathematical structure into observer-dependent physics.

The central question is: given a partial order on events with no preferred total ordering, how do different observers reconstruct temporal experience, and do the resulting differences reproduce features of relativistic kinematics?

\subsection*{Scope}

This paper derives observer-dependent time functions, proves relativity of simultaneity, and exhibits a time-dilation analogue. It does not derive Lorentz invariance, construct a metric, or claim equivalence with special relativity. It establishes that relativity-like kinematic structure emerges from the event poset without spacetime assumptions.

%==================================================================
\section{Observer Definition}\label{sec:observer}
%==================================================================

\begin{definition}[Observer]\label{def:observer}
An \textit{observer} on the event poset $(\Ecal, \prec)$ is a pair
\begin{equation}
    O = (\leq_O,\; \mu_O)
\end{equation}
where:
\begin{itemize}
    \item $\leq_O$ is a linear extension of $\prec$: a total order on $\Ecal$ such that $e_1 \prec e_2 \Rightarrow e_1 \leq_O e_2$,
    \item $\mu_O: \Ecal \to \mathbb{R}_{>0}$ is a \textit{sampling measure}: a positive function assigning a ``duration weight'' to each event.
\end{itemize}
\end{definition}

\begin{remark}
The linear extension $\leq_O$ determines \textit{what order} the observer assigns to events. The sampling measure $\mu_O$ determines \textit{how much time} the observer associates with each event. Together they produce a complete observer time function.
\end{remark}

\begin{remark}
No physical velocity, mass, or energy is assumed. The sampling measure is a structural degree of freedom intrinsic to the observer's relationship with the event poset.
\end{remark}

%==================================================================
\section{Observer Time Functions}\label{sec:time}
%==================================================================

\begin{definition}[Observer time]\label{def:time}
For an observer $O = (\leq_O, \mu_O)$, the \textit{observer time function} is:
\begin{equation}\label{eq:time}
    t_O(e) = \sum_{\substack{e' \in \Ecal \\ e' \leq_O e}} \mu_O(e')
\end{equation}
That is, the time assigned to event $e$ is the cumulative measure of all events up to and including $e$ in the observer's ordering.
\end{definition}

\begin{proposition}[Order preservation]\label{prop:preserve}
For any observer $O$:
\begin{equation}
    e_1 \prec e_2 \;\Rightarrow\; t_O(e_1) < t_O(e_2)
\end{equation}
\end{proposition}

\begin{proof}
If $e_1 \prec e_2$, then $e_1 \leq_O e_2$ (since $\leq_O$ extends $\prec$) and $e_1 \neq e_2$. The sum defining $t_O(e_2)$ includes all terms in $t_O(e_1)$ plus at least $\mu_O(e_2) > 0$. Therefore $t_O(e_2) > t_O(e_1)$.
\end{proof}

\begin{remark}
The time function $t_O$ is not unique: it depends on both the choice of linear extension and the choice of sampling measure. Different observers generically assign different times to the same events, even when they agree on the partial order.
\end{remark}

%==================================================================
\section{Relativity of Simultaneity}\label{sec:simultaneity}
%==================================================================

\begin{theorem}[Relativity of simultaneity]\label{thm:simultaneity}
If events $e_1, e_2 \in \Ecal$ are incomparable under $\prec$ (neither $e_1 \prec e_2$ nor $e_2 \prec e_1$), then there exist observers $O_1, O_2$ such that:
\begin{equation}
    t_{O_1}(e_1) < t_{O_1}(e_2) \qquad\text{and}\qquad t_{O_2}(e_2) < t_{O_2}(e_1)
\end{equation}
\end{theorem}

\begin{proof}
Since $e_1$ and $e_2$ are incomparable, there exist linear extensions $\leq_1$ placing $e_1$ before $e_2$, and $\leq_2$ placing $e_2$ before $e_1$. (This follows from the standard extension property of finite partial orders: for any two incomparable elements, linear extensions exist placing them in either relative order.) Choose any positive measure $\mu > 0$ and set $O_1 = (\leq_1, \mu)$, $O_2 = (\leq_2, \mu)$. Then $t_{O_1}(e_1) < t_{O_1}(e_2)$ and $t_{O_2}(e_2) < t_{O_2}(e_1)$ by construction.
\end{proof}

\begin{remark}
This is the geometric analogue of the relativity of simultaneity in special relativity: events that are spacelike-separated (incomparable in the causal order) can be ordered differently by different inertial frames (linear extensions). The VFD framework derives this from the event poset; special relativity derives it from Lorentz invariance. The structural role is analogous.
\end{remark}

%==================================================================
\section{Time Dilation Analogue}\label{sec:dilation}
%==================================================================

\begin{definition}[Elapsed time along a chain]\label{def:elapsed}
For a chain $C = (e_1 \prec e_2 \prec \cdots \prec e_n) \subset \Ecal$ and observer $O$, the elapsed time is:
\begin{equation}\label{eq:elapsed}
    \Delta t_O(C) = t_O(e_n) - t_O(e_1) = \sum_{\substack{e' \in \Ecal \\ e_1 <_O e' \leq_O e_n}} \mu_O(e')
\end{equation}
\end{definition}

\begin{theorem}[Time-dilation analogue]\label{thm:dilation}
Let $O_1 = (\leq, \mu_1)$ and $O_2 = (\leq, \mu_2)$ be observers sharing the same linear extension but with different sampling measures. If $\mu_1$ and $\mu_2$ differ on the set of events lying between $e_1$ and $e_n$ in the shared ordering, then:
\begin{equation}
    \Delta t_{O_1}(C) \neq \Delta t_{O_2}(C)
\end{equation}
for the chain $C = (e_1 \prec \cdots \prec e_n)$. Both observers agree on the ordering of events but disagree on the elapsed time.
\end{theorem}

\begin{proof}
Since $O_1$ and $O_2$ share the ordering $\leq$, the set $S$ of events between $e_1$ and $e_n$ is identical for both. By hypothesis, $\mu_1$ and $\mu_2$ differ on $S$, so:
\begin{equation}
    \sum_{e' \in S} \mu_1(e') \neq \sum_{e' \in S} \mu_2(e')
\end{equation}
giving $\Delta t_{O_1}(C) \neq \Delta t_{O_2}(C)$.
\end{proof}

\begin{remark}
This is the event-order analogue of time dilation: two observers traversing the same sequence of events measure different elapsed times. In special relativity, this arises from relative velocity; here it arises from different sampling measures. The structural effect --- same events, different durations --- is the same.
\end{remark}

\begin{remark}
The sampling measure $\mu_O$ is not yet connected to physical velocity or energy. Establishing that connection requires metric structure (Paper~XXV). The present paper demonstrates only that the \textit{kinematic possibility} of time dilation exists in the event-order framework.
\end{remark}

%==================================================================
\section{Worldlines}\label{sec:worldlines}
%==================================================================

\begin{definition}[Worldline]\label{def:worldline}
A \textit{worldline} is a maximal chain in $(\Ecal, \prec)$: a totally ordered subset $W \subset \Ecal$ that cannot be extended by inclusion.
\end{definition}

\begin{remark}
A worldline is a complete sequence of events each following the previous in the partial order. Different worldlines may share some events (intersect) or be entirely disjoint.
\end{remark}

\begin{definition}[Chain time]\label{def:chaintime}
For an observer $O$ and a chain $W = (e_1 \prec \cdots \prec e_n)$, the \textit{chain time} (or worldline time) is:
\begin{equation}
    \tau_O(W) = \sum_{k=1}^{n} \mu_O(e_k)
\end{equation}
This is the cumulative measure restricted to events on the chain itself.
\end{definition}

\begin{remark}
Chain time $\tau_O(W)$ differs in general from the observer elapsed time $\Delta t_O(W) = t_O(e_n) - t_O(e_1)$, since $\Delta t_O$ includes contributions from all events between $e_1$ and $e_n$ in the observer's total ordering, not just those on the chain. The chain time $\tau_O$ is the intrinsic duration along the worldline; $\Delta t_O$ is the global clock reading.
\end{remark}

%==================================================================
\section{Relative Velocity (Proto-Concept)}\label{sec:velocity}
%==================================================================

A full notion of velocity requires spatial structure, which this paper does not construct. However, a proto-concept can be defined.

\begin{definition}[Relative sampling rate]\label{def:velocity}
For two observers $O_1, O_2$, their \textit{relative sampling rate} along a chain $C$ is:
\begin{equation}
    \rho(O_1, O_2; C) = \frac{\tau_{O_1}(C)}{\tau_{O_2}(C)} = \frac{\sum_{e \in C} \mu_1(e)}{\sum_{e \in C} \mu_2(e)}
\end{equation}
\end{definition}

When $\rho \neq 1$, the two observers experience the same chain at different rates. This is the structural precursor to relative velocity: two observers ``moving'' through the event structure at different ``speeds'' (sampling rates).

\begin{remark}
This proto-velocity does not have units of distance/time. It is a dimensionless ratio of sampling measures. Connecting it to physical velocity requires the spatial metric structure of Paper~XXV.
\end{remark}

%==================================================================
\section{Worked Example: Dual Pentagons}\label{sec:example}
%==================================================================

\begin{example}[Extending the pentagon system]\label{ex:pentagon}
Using the dual-pentagon event set from Paper~XXIII: 5 events at $\theta_b = 0$ and 5 events at $\theta_b = 2\pi/5$.

\textbf{Two observers.} Define:
\begin{itemize}
    \item $O_1 = (\leq_1, \mu_1)$: orders the $\theta_b = 0$ events as $(0,0) < (1,1) < (2,2) < (3,3) < (4,4)$, with uniform measure $\mu_1(e) = 1$ for all $e$.
    \item $O_2 = (\leq_2, \mu_2)$: orders the $\theta_b = 0$ events as $(2,2) < (4,4) < (1,1) < (3,3) < (0,0)$ (a different permutation), with non-uniform measure $\mu_2(e) = 2$ for $\theta_b = 0$ events and $\mu_2(e) = 1$ for $\theta_b = 2\pi/5$ events.
\end{itemize}

Both observers agree: all $\theta_b = 0$ events precede all $\theta_b = 2\pi/5$ events.

\textbf{Simultaneity.} $O_1$ assigns $t_{O_1}(0,0) = 1 < t_{O_1}(1,1) = 2$. But $O_2$ assigns $t_{O_2}(0,0) = 10 > t_{O_2}(1,1) = 6$. The events $(0,0)$ and $(1,1)$ are ordered differently by the two observers: \textit{relativity of simultaneity demonstrated}.

\textbf{Time dilation.} Consider the chain $C$: any $\theta_b = 0$ event followed by any $\theta_b = 2\pi/5$ event (e.g.\ $(0,0) \prec (0,4)$).
\begin{itemize}
    \item $O_1$: $\Delta t_{O_1}(C) = t_{O_1}(0,4) - t_{O_1}(0,0)$. With uniform $\mu_1 = 1$, if $(0,0)$ is the 1st event and $(0,4)$ is the 6th: $\Delta t = 5$.
    \item $O_2$: with $\mu_2 = 2$ for the first layer, if $(0,0)$ is the 5th event in layer 1: $t_{O_2}(0,0) = 10$. Then $(0,4)$ in layer 2 at position 6: $t_{O_2}(0,4) = 10 + 1 = 11$. So $\Delta t = 1$.
\end{itemize}
Same chain, different elapsed times: $\Delta t_{O_1} = 5 \neq \Delta t_{O_2} = 1$. \textit{Time dilation analogue demonstrated.}

\textbf{Relative sampling rate.} $\rho(O_1, O_2; C) = 5/1 = 5$. Observer $O_1$ accumulates 5 times more elapsed time than $O_2$ along this chain.
\end{example}

%==================================================================
\section{Relation to Special Relativity}\label{sec:SR}
%==================================================================

\begin{table}[ht]
\centering
\caption{Structural comparison: VFD event-order framework vs.\ special relativity.}
\label{tab:SR}
\begin{tabular}{@{}lll@{}}
\toprule
Feature & VFD (this paper) & Special relativity \\
\midrule
Events & Admissible vertex pairs & Spacetime events \\
Ordering & Birth-angle partial order & Causal structure \\
Observer & Linear extension + measure & Inertial frame \\
Time function & Cumulative sampling measure & Proper time \\
Simultaneity & Observer-dependent (proved) & Frame-dependent \\
Time dilation & Sampling-rate differences (exhibited) & Velocity-dependent \\
Worldline & Maximal chain in poset & Timelike curve \\
Velocity & Proto-concept (sampling ratio) & $dx/dt$ \\
\bottomrule
\end{tabular}
\end{table}

\begin{remark}[Epistemic status]
The structural analogies in Table~\ref{tab:SR} are \textit{exhibited}, not derived from Lorentz invariance. The VFD framework produces the same qualitative phenomena (observer-dependent simultaneity, time dilation, worldline structure) from a different starting point (event poset with sampling measures). Whether this structural analogy can be promoted to full Lorentz covariance requires metric structure not yet available.
\end{remark}

%==================================================================
\section{Discussion}\label{sec:discussion}
%==================================================================

\subsection{What is established}

\begin{itemize}
    \item Observer defined as linear extension + sampling measure (Definition~\ref{def:observer}).
    \item Observer time function constructed explicitly (Definition~\ref{def:time}, Proposition~\ref{prop:preserve}).
    \item Relativity of simultaneity proved (Theorem~\ref{thm:simultaneity}).
    \item Time-dilation analogue exhibited under differing observer measures (Theorem~\ref{thm:dilation}).
    \item Worldlines defined as maximal chains (Definition~\ref{def:worldline}).
    \item Worked example demonstrates all constructions.
\end{itemize}

\subsection{What is not established}

\begin{itemize}
    \item No Lorentz invariance. The framework has discrete ($H_4$) symmetry, not continuous Lorentz symmetry.
    \item No metric. The sampling measure gives duration differences but not distances.
    \item No Einstein equations. The connection between observer structure and gravitational dynamics is not addressed.
    \item No physical velocity. The relative sampling rate is a structural precursor, not a derived physical quantity.
    \item The sampling measure $\mu_O$ is a structural degree of freedom, not yet connected to physical observables (mass, energy, motion).
\end{itemize}

\subsection{Programme position}

\begin{itemize}
    \item Paper~XXIII: event structure and emergent temporal ordering.
    \item Paper~XXIV (this paper): observer frames and relativity-like kinematics.
    \item Paper~XXV (future): metric structure and the route to dynamics.
\end{itemize}

%==================================================================
\section{Conclusion}\label{sec:conclusion}
%==================================================================

The event poset of Paper~XXIII, equipped with observer-dependent sampling measures, supports a kinematic precursor of relativistic structure: observer-dependent simultaneity, time-dilation analogues, and worldline geometry. These features are proved, not assumed. No spacetime manifold, metric, or Lorentz symmetry is postulated.

The key structural insight is that the non-uniqueness of the global clock (Paper~XXIII) is not a deficiency but a feature: it is the geometric origin of observer dependence. Different observers are different ways of completing the partial order into a total order, and different sampling measures produce different experienced durations.

The next step is to equip the event structure with spatial and metric content (Paper~XXV), connecting the sampling measure to physical observables and determining whether the full Lorentz-covariant structure can be recovered from the 600-cell geometry.

%==================================================================
\bibliographystyle{plainnat}
\bibliography{references}

\end{document}
