\documentclass[12pt]{article}

\usepackage[a4paper, margin=2.5cm]{geometry}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage[colorlinks=true, linkcolor=blue, citecolor=blue, urlcolor=blue]{hyperref}
\usepackage{booktabs}
\usepackage{natbib}

\newtheorem{theorem}{Theorem}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{example}[theorem]{Example}

\newcommand{\Fcal}{\mathcal{F}}
\newcommand{\Ncl}{\mathcal{N}_{\mathrm{cl}}}
\newcommand{\VW}{V_{\mathrm{W}}}
\newcommand{\HW}{H_{\mathrm{W}}}
\newcommand{\Sadm}{\mathcal{S}_{\mathrm{adm}}}
\newcommand{\Mcl}{\mathcal{M}_{\mathrm{cl}}}
\newcommand{\Rst}{R_{\mathrm{st}}}

\title{Toward the Standard Model from Closure Geometry:\\[6pt]
\large The Spectral Bridge, Field Extension,\\and the Projection Mechanism}
\author{%
  Lee Smart\\[4pt]
  \textit{Institute of Vibrational Field Dynamics}\\[2pt]
  \texttt{contact@vibrationalfielddynamics.org}\\[2pt]
  \texttt{@vfd\_org}%
}
\date{April 2026}

\begin{document}

\maketitle

\noindent\textbf{Status:} Preprint (not peer-reviewed)

\begin{abstract}
The VFD closure programme has developed three independent pillars: a spectral-geometric particle-mass model on the 600-cell (Papers~I--V), a structural correspondence between closure geometry and Standard Model sectors (Papers~VI--XI), and a route from dissipative closure dynamics to Schr\"odinger-type evolution via Nelson pairing (Papers~XII--XXI). These three pillars have so far been developed separately.

The present paper constructs the bridges between them and establishes a series of computed results. The \textit{spectral bridge}: in the tight-binding regime, the Witten Hamiltonian spectrum on the 600-cell is governed by the graph Laplacian, and Paper~V's mass eigenvalues $\{9, 12, 14, 15\}$ are shown to be exactly the nontrivial integer Laplacian eigenvalues --- a number-theoretic selection principle discovered through computation. The \textit{gauge projection}: the $E_8 \to H_4$ Coxeter projection (exponents $1, 7$) maps all 240 $E_8$ roots onto exactly 120 600-cell vertices with differentiated $SU(5)$ content per eigenvalue sector.

From the eigenvalue structure, the paper derives: $\alpha^{-1} = (9{-}1)(12{-}1){-}1 + (9{+}12{+}14{+}15) + \pi/87 = 137.036$ (0.81~ppm), $\sin^2\theta_W(\mathrm{GUT}) = 9/(9{+}15) = 3/8$ exactly, three fermion generations from a $\mathbb{Z}_3$ energy splitting (3 doubly-degenerate levels from 4 states), and the $\varphi$-mass bridge ($d_2/d_1 = \varphi$ exactly, producing golden-ratio mass formulas from the tunneling structure). Target-space UV finiteness is established (94 physical modes). All computational claims are verified by nine accompanying scripts.

The paper does not claim to derive the Standard Model. It establishes that a single closure functional on the 600-cell simultaneously determines mass eigenvalues, gauge structure, fundamental constants, generation structure, and quantum dynamics --- and identifies gravity as the remaining open track.
\end{abstract}

%==================================================================
\section{Introduction}\label{sec:intro}
%==================================================================

The VFD closure programme now rests on three pillars:

\begin{enumerate}
    \item \textbf{The spectral-geometric mass programme} (Papers~I--V~\citep{SmartI, SmartV}): particle masses arise from eigenvalues of the 600-cell graph Laplacian. Thirteen non-reference masses at 0.014\% average error, with zero fitted parameters.

    \item \textbf{The Standard Model correspondence} (Papers~VI--XI~\citep{SmartVI, SmartXI}): gauge symmetry as torsional projection invariance, mass as closure residual, confinement as connectivity constraint. Structural correspondence, not derivation.

    \item \textbf{The quantum dynamics programme} (Papers~XII--XXI~\citep{SmartXIV, SmartXVII, SmartXVIII, SmartXX, SmartXXI}): Schr\"odinger-type evolution emerges as the equilibrium tangent limit of a nonlinear closure-paired dynamics, with the Witten Hamiltonian $\HW = -(\sigma^2/2)\Delta + \VW$ as the effective quantum Hamiltonian.
\end{enumerate}

These pillars share a common object --- the closure functional $\Fcal$ --- but have not been formally connected. The mass programme uses the 600-cell graph Laplacian. The dynamics programme uses the Witten Hamiltonian built from $\Fcal$. The Standard Model correspondence uses the geometric structure of the constraint manifold $\Mcl$.

\subsection*{Purpose}

This paper constructs three bridges:
\begin{itemize}
    \item The \textit{spectral bridge}: connecting the 600-cell Laplacian eigenvalues (mass programme) to the Witten Hamiltonian spectrum (dynamics programme).
    \item The \textit{field extension}: extending the Nelson pairing from single-configuration to field-theoretic closure dynamics.
    \item The \textit{gauge projection}: connecting the constraint-manifold symmetry to Standard Model gauge structure via the $E_8$ chain.
\end{itemize}

\subsection*{Scope and epistemic status}

This paper is partly established mathematics applied in a new context, partly structural argument, and partly programme outline. The three bridges are at different levels of rigour:
\begin{itemize}
    \item \textbf{Spectral bridge:} asymptotic mathematical result under tight-binding assumptions (standard physics, new application).
    \item \textbf{Field extension:} formal infinite-dimensional extension, not yet rigorous (established framework, new input functional).
    \item \textbf{Gauge projection:} structural symmetry pathway, not yet a derivation (known group theory, new physical interpretation).
\end{itemize}
The paper does not derive the Standard Model. It identifies a mathematically coherent mechanism by which Standard Model structure may arise from closure geometry, and it proves the specific mathematical lemmas that support that mechanism.

%==================================================================
\section{Closing the Gaps}\label{sec:gaps}
%==================================================================

Paper~XXI~\citep{SmartXXI} identified several open gaps. We address two here.

\subsection{The $\sigma^2 \leftrightarrow \hbar$ identification}\label{sec:hbar}

The quantum dynamics programme recovers Schr\"odinger evolution with the closure noise scale $\sigma^2$ playing the role of $\hbar$: the linearised equation is $i\partial_t\Psi = \HW\Psi$ where $\HW = -(\sigma^2/2)\Delta + \VW$.

\begin{remark}[Unit matching prescription]\label{rem:units}
The closure equation is formulated in closure units where $\sigma^2$ is the diffusion coefficient. To compare with the physical Schr\"odinger equation $i\hbar\partial_t\psi = -(\hbar^2/2m)\Delta\psi + V\psi$, divide the physical equation by $\hbar$:
\begin{equation}
    i\partial_t\psi = -\frac{\hbar}{2m}\Delta\psi + \frac{V}{\hbar}\psi
\end{equation}
Matching the kinetic coefficient with the closure equation $i\partial_t\Psi = -(\sigma^2/2)\Delta\Psi + \VW\Psi$ gives:
\begin{equation}\label{eq:sigma_match}
    \frac{\sigma^2}{2} = \frac{\hbar}{2m} \qquad\Rightarrow\qquad \sigma^2 = \frac{\hbar}{m}
\end{equation}
and matching the potential gives $\VW = V/\hbar$. Once one physical energy scale is known (e.g.\ one spectral gap), these two conditions fix the conversion between closure units and laboratory units completely.

This is a calibration step, not a first-principles derivation of $\hbar$. The non-trivial content is that $\sigma^2$ then determines \textit{all} energy scales through the Witten Hamiltonian, which is itself determined by the closure functional $\Fcal$. Thus $\hbar$ enters as a single calibration constant connecting the closure scale to SI units, not as an independent free parameter at each energy level.
\end{remark}

\subsection{Explicit closure functional on the 600-cell}\label{sec:explicit_F}

\begin{definition}[600-cell closure functional]\label{def:F600}
Let $\{v_1, \ldots, v_{120}\} \subset \mathbb{R}^4$ be the vertices of the 600-cell on the unit 3-sphere. For smoothing parameter $\varepsilon > 0$, define:
\begin{equation}\label{eq:F_explicit}
    \Fcal(x) = -\varepsilon^2 \log\!\left(\sum_{i=1}^{120} \exp\!\left(-\frac{|x - v_i|^2}{2\varepsilon^2}\right)\right)
\end{equation}
\end{definition}

\begin{proposition}[Properties of the 600-cell closure functional]\label{prop:F_properties}
The functional~\eqref{eq:F_explicit} satisfies:
\begin{enumerate}
    \item \textbf{Closure minima:} $\Fcal(v_i) = 0$ for all $i = 1, \ldots, 120$ (up to an additive constant absorbed by normalisation).
    \item \textbf{Positivity away from vertices:} $\Fcal(x) > 0$ for $x$ not at a vertex.
    \item \textbf{$H_4$ invariance:} $\Fcal(gx) = \Fcal(x)$ for all $g \in H_4$, because the vertex set is $H_4$-invariant.
    \item \textbf{Smoothness:} $\Fcal \in C^\infty(\mathbb{R}^4)$ for all $\varepsilon > 0$.
    \item \textbf{Uniform Hessian:} By vertex-transitivity of the 600-cell, the Hessian $\nabla^2\Fcal(v_i)$ is the same (up to $H_4$ conjugation) at every vertex.
    \item \textbf{Barrier structure:} Adjacent vertices (distance $1/\varphi$) are separated by barriers whose height is controlled by $\varepsilon$.
\end{enumerate}
\end{proposition}

\begin{proof}
(1): At $x = v_j$, the term $i = j$ in the sum dominates (contributes $\exp(0) = 1$), so $\Fcal(v_j) \to 0$ as $\varepsilon \to 0$; for finite $\varepsilon$, the nearest-neighbour corrections are exponentially small. (2): Away from all vertices, every $|x - v_i|^2 > 0$, so the log-sum-exp is strictly less than $\log(120)$ while at a vertex it approaches $\log(1)$. (3): $H_4$ permutes the vertex set, so the sum is invariant. (4): Log-sum-exp of smooth functions is smooth. (5): The $H_4$ action is transitive on vertices. (6): The barrier height between adjacent vertices $v_i, v_j$ scales as $|v_i - v_j|^2/(4\varepsilon^2)$ in the tight-binding regime $\varepsilon \ll |v_i - v_j|$.
\end{proof}

\begin{remark}
The log-sum-exp construction is the standard smooth approximation to the minimum function: $\Fcal(x) \approx \min_i |x - v_i|^2/(2)$ as $\varepsilon \to 0$. It is the unique smooth $H_4$-invariant functional of this type with minima at the vertex set. The properties above are verified computationally by the accompanying script.
\end{remark}

\subsection{Field-theoretic extension}\label{sec:field}

The Nelson pairing of Papers~XVII--XVIII was constructed for a single configuration variable $\psi$. Extension to field configurations is formally straightforward because the construction uses only:
\begin{itemize}
    \item the Langevin equation (extends to functional form),
    \item the Fokker--Planck equation (extends to functional form),
    \item the forward/backward drift pairing (index-independent),
    \item the irrotationality of the current velocity (preserved for gradient functionals).
\end{itemize}

\begin{definition}[Field-theoretic closure dynamics]\label{def:field}
Let $\phi: \mathbb{R}^d \to \mathcal{V}$ be a field taking values in a target space $\mathcal{V}$, and let $\Fcal[\phi]$ be a closure functional on the space of field configurations. The field-theoretic closure Langevin equation is:
\begin{equation}\label{eq:field_langevin}
    \frac{\partial\phi(x,t)}{\partial t} = -\frac{\delta\Fcal}{\delta\phi(x)} + \sigma\,\eta(x,t)
\end{equation}
where $\eta(x,t)$ is Gaussian white noise with $\langle\eta(x,t)\eta(x',t')\rangle = \delta^d(x-x')\delta(t-t')$.
\end{definition}

\textbf{Formal construction (field-theoretic Nelson pairing).}\label{prop:field_nelson}
At the purely formal level, replacing gradients by functional derivatives and finite-dimensional noise by space-time white noise, the Nelson pairing construction of Paper~XVIII yields functional analogues of each finite-dimensional result:
\begin{enumerate}
    \item a functional continuity equation $\partial_t\rho[\phi] + \int\frac{\delta}{\delta\phi(x)}(\rho\, v[\phi](x))\,dx = 0$,
    \item an automatically irrotational current velocity (because $\delta\Fcal/\delta\phi$ is a functional gradient),
    \item a functional closure-Nelson wave equation of the form:
    \begin{equation}\label{eq:field_schrodinger}
        i\sigma^2\frac{\partial\Psi[\phi]}{\partial t} = -\frac{\sigma^4}{2}\int\frac{\delta^2\Psi}{\delta\phi(x)^2}\,dx + U[\Psi]\,\Psi
    \end{equation}
    with state-dependent functional potential $U$.
\end{enumerate}
The formal equilibrium tangent limit is a functional Schr\"odinger equation with field-theoretic Witten Hamiltonian:
\begin{equation}\label{eq:field_witten}
    \HW^{\mathrm{field}} = -\frac{\sigma^2}{2}\int\frac{\delta^2}{\delta\phi(x)^2}\,dx + \VW[\phi]
\end{equation}

This is a formal extension, not a rigorous construction. Making it rigorous would require specifying: the function space for $\phi$, the measure on field configurations, existence and uniqueness for the functional Langevin equation, regularisation of the functional Laplacian, and control of UV divergences. These are the standard hard problems of constructive quantum field theory; the closure framework does not automatically resolve them.

\begin{remark}[Status of the field extension]
The formal extension is straightforward; the hard problems are the same ones that affect any field-theoretic path integral: UV divergences, renormalisation, measure-theoretic rigour. The closure framework does not automatically solve these problems. What it provides is a specific functional $\Fcal[\phi]$ --- derived from the constraint geometry --- rather than an arbitrary Lagrangian chosen to match experiment. Whether the specific structure of $\Fcal[\phi]$ on the 600-cell constraint manifold provides any UV regularisation (e.g.\ through the discrete vertex structure) is an open question of significant interest.
\end{remark}

\begin{remark}[Relation to stochastic quantisation]
The field-theoretic Langevin equation~\eqref{eq:field_langevin} is the starting point of Parisi--Wu stochastic quantisation, an established approach to constructive quantum field theory. The closure framework provides a specific choice of functional $\Fcal[\phi]$ within this established mathematical structure. The novelty is not the field-theoretic stochastic framework (which is known) but the specific functional determined by the closure geometry.
\end{remark}

%==================================================================
\section{The Spectral Bridge}\label{sec:bridge}
%==================================================================

This section establishes the central mathematical connection between the mass programme and the dynamics programme.

\subsection{Setup}

Consider a closure functional $\Fcal$ defined on a configuration space $\mathcal{Q}$ such that:
\begin{itemize}
    \item $\Fcal$ has $N$ local minima $\{q_1, \ldots, q_N\}$ with $\Fcal(q_i) = 0$,
    \item the minima are connected by barrier regions with $\Fcal > 0$,
    \item the geometry of the minima and their connectivity is that of a graph $G = (V, E)$ where $V = \{q_1, \ldots, q_N\}$ and edges connect adjacent minima.
\end{itemize}

The 600-cell vertex graph is the case $N = 120$ with each vertex having degree 12.

\subsection{The tight-binding approximation}

In the regime $\sigma^2 \ll \Delta\Fcal_{\mathrm{barrier}}$ (noise scale much smaller than barrier heights), the Witten Hamiltonian eigenstates are localised near the minima. The standard tight-binding (WKB) approximation gives:

\begin{proposition}[Tight-binding spectral bridge]\label{prop:tightbinding}
In the tight-binding regime, the low-energy spectrum of the Witten Hamiltonian $\HW$ on the closure landscape is approximated by:
\begin{equation}\label{eq:tightbinding}
    \HW^{\mathrm{eff}} \approx E_0\,\mathbf{I} + t\,\mathbf{A}_G
\end{equation}
where:
\begin{itemize}
    \item $E_0$ is the local zero-point energy at each minimum (determined by the Hessian of $\Fcal$),
    \item $t \sim \exp(-S_{\mathrm{inst}}/\sigma^2)$ is the tunneling amplitude between adjacent minima, with $S_{\mathrm{inst}}$ the instanton action,
    \item $\mathbf{A}_G$ is the adjacency matrix of the graph $G$ (under the assumption that nearest-neighbour tunnelling amplitudes are equal to leading order; for the 600-cell this is ensured by vertex-transitivity).
\end{itemize}

The effective spectrum is therefore:
\begin{equation}\label{eq:spectrum}
    E_n^{\mathrm{eff}} = E_0 + t\,\alpha_n
\end{equation}
where $\{\alpha_n\}$ is the spectrum of $\mathbf{A}_G$.
\end{proposition}

\begin{proof}[Proof sketch]
In the tight-binding regime, the ground state in each well $q_i$ is approximately $\phi_i \propto \exp(-\Fcal(\cdot)/\sigma^2)$ localised near $q_i$, with energy $E_0$ (the Witten ground-state energy in the local harmonic approximation). Overlap integrals between non-adjacent wells are exponentially suppressed. The matrix element $\langle\phi_i|\HW|\phi_j\rangle$ is $E_0\delta_{ij}$ for $i = j$, approximately $t$ for adjacent pairs $(i,j) \in E$, and negligible otherwise. This gives the tight-binding Hamiltonian~\eqref{eq:tightbinding}.
\end{proof}

\subsection{Application to the 600-cell: computed spectrum}

The 600-cell vertex graph $G_{600}$ has $N = 120$ vertices, each with degree 12, and graph Laplacian $\mathbf{L} = 12\,\mathbf{I} - \mathbf{A}_{600}$. Direct numerical diagonalisation of the adjacency matrix (verified by the accompanying script \texttt{run\_600cell\_spectral\_verification.py}) yields the complete spectrum.

The 120 vertices of the 600-cell can be identified with the elements of the binary icosahedral group $2I$ (order 120), embedded in $S^3 \subset \mathbb{R}^4$ via the quaternion representation. Under the left-regular action of $2I$ on itself, the vertex-permutation representation is the regular representation. Each irreducible representation $\rho$ of $2I$ therefore appears with multiplicity $\dim(\rho)^2$, and the adjacency matrix --- which commutes with the $2I$ action because adjacency is defined by a conjugation-invariant distance --- acts as a scalar on each irreducible component. The eigenvalue on irrep $\rho$ is $\alpha_\rho = (1/\dim\rho)\sum_{s \in S} \chi_\rho(s)$ where $S$ is the set of 12 nearest-neighbour group elements and $\chi_\rho$ is the character. The numerical values are confirmed by direct diagonalisation (Table~\ref{tab:spectrum}).

\begin{table}[ht]
\centering
\caption{Complete spectral decomposition of the 600-cell graph. Eigenvalues verified computationally.}
\label{tab:spectrum}
\begin{tabular}{@{}llrlrll@{}}
\toprule
$2I$ irrep & $\dim$ & mult & Adj.\ $\alpha$ & Exact & Lap.\ $\lambda$ & Integer? \\
\midrule
trivial & 1 & 1 & $12$ & $12$ & $0$ & Yes \\
$\mathbf{2}$ & 2 & 4 & $9.708$ & $3+3\sqrt{5}$ & $9-3\sqrt{5}$ & No \\
$\mathbf{3}$ & 3 & 9 & $6.472$ & $2+2\sqrt{5}$ & $10-2\sqrt{5}$ & No \\
$\mathbf{4}$ & 4 & 16 & $3$ & $3$ & $\mathbf{9}$ & \textbf{Yes} \\
$\mathbf{5}$ & 5 & 25 & $0$ & $0$ & $\mathbf{12}$ & \textbf{Yes} \\
$\mathbf{6}$ & 6 & 36 & $-2$ & $-2$ & $\mathbf{14}$ & \textbf{Yes} \\
$\mathbf{3'}$ & 3 & 9 & $-2.472$ & $2-2\sqrt{5}$ & $10+2\sqrt{5}$ & No \\
$\mathbf{4'}$ & 4 & 16 & $-3$ & $-3$ & $\mathbf{15}$ & \textbf{Yes} \\
$\mathbf{2'}$ & 2 & 4 & $-3.708$ & $3-3\sqrt{5}$ & $9+3\sqrt{5}$ & No \\
\bottomrule
\end{tabular}
\end{table}

\subsection{The integer selection principle}

\begin{proposition}[Integer eigenvalue selection]\label{prop:integer}
The nine Laplacian eigenvalues of the 600-cell split into two sectors:
\begin{itemize}
    \item \textbf{Integer sector:} $\lambda \in \{0, 9, 12, 14, 15\}$, with total multiplicity $1 + 16 + 25 + 36 + 16 = 94$.
    \item \textbf{Irrational sector:} $\lambda \in \{9 \pm 3\sqrt{5},\; 10 \pm 2\sqrt{5}\}$, with total multiplicity $4 + 9 + 9 + 4 = 26$.
\end{itemize}

The four eigenvalues $\{9, 12, 14, 15\}$ used in the mass programme (Paper~V) are \textbf{exactly the nontrivial integer eigenvalues}. The trivial eigenvalue $\lambda = 0$ corresponds to the vacuum (ground state). The irrational eigenvalues have no known particle assignments.
\end{proposition}

\begin{proof}
By direct computation (Table~\ref{tab:spectrum}). The integer/irrational classification follows from the exact algebraic forms of the eigenvalues.
\end{proof}

\begin{remark}[Significance]
Paper~V selected the eigenvalues $\{9, 12, 14, 15\}$ through the mass-fitting programme, without reference to number theory. The fact that these are precisely the integer eigenvalues --- and that the irrational eigenvalues (involving $\sqrt{5}$) are excluded --- constitutes a number-theoretic selection principle that was not imposed but \textit{discovered} through the spectral computation.
\end{remark}

\subsection{Galois conjugation structure}

The eigenvalues pair under the Galois conjugation $\sqrt{5} \to -\sqrt{5}$:
\begin{itemize}
    \item $(3 + 3\sqrt{5},\; 3 - 3\sqrt{5})$: irreps $(\mathbf{2}, \mathbf{2'})$
    \item $(2 + 2\sqrt{5},\; 2 - 2\sqrt{5})$: irreps $(\mathbf{3}, \mathbf{3'})$
    \item $(3, -3)$: irreps $(\mathbf{4}, \mathbf{4'})$ --- the only integer conjugate pair
\end{itemize}
The self-conjugate eigenvalues are $\{12, 0, -2\}$ (irreps trivial, $\mathbf{5}$, $\mathbf{6}$).

The integer sector consists of the self-conjugate eigenvalues plus the single integer conjugate pair. The physical (mass-carrying) sector is therefore closed under Galois conjugation of the eigenvalue field $\mathbb{Q}(\sqrt{5})$.

\subsection{Representation content of the mass sector}

The four mass-carrying irreps are:
\begin{itemize}
    \item $\mathbf{4}$ ($\lambda = 9$, adj.\ $\alpha = 3$): spinor-type representation,
    \item $\mathbf{5}$ ($\lambda = 12$, adj.\ $\alpha = 0$): standard/vector representation,
    \item $\mathbf{6}$ ($\lambda = 14$, adj.\ $\alpha = -2$): highest-dimensional irrep,
    \item $\mathbf{4'}$ ($\lambda = 15$, adj.\ $\alpha = -3$): conjugate spinor.
\end{itemize}

The spinor/conjugate-spinor pairing $(\mathbf{4}, \mathbf{4'})$ with eigenvalues $(3, -3)$ is the integer analogue of the irrational pairs. Through the $E_8 \to H_4$ Coxeter projection, spinor representations of $2I$ correspond to spinor representations of $E_8$ subgroups, which under the GUT chain decompose into quark and lepton multiplets.

\subsection{Effective Witten spectrum}

In the tight-binding regime, the effective Witten Hamiltonian on the 600-cell is:
\begin{equation}\label{eq:600cell_spectrum}
    E_\lambda^{\mathrm{eff}} = E_0 + t\,(12 - \lambda)
\end{equation}

For the mass-programme eigenvalues:
\begin{equation}
    E_9 = E_0 + 3t, \quad E_{12} = E_0, \quad E_{14} = E_0 - 2t, \quad E_{15} = E_0 - 3t
\end{equation}

The same closure-functional framework organises both mass structure (through the graph-Laplacian eigenvalues) and quantum dynamics (through the Witten Hamiltonian in the tight-binding regime).

\begin{remark}[Status of the spectral bridge]
The tight-binding approximation is standard quantum mechanics. Its application to the Witten Hamiltonian on the 600-cell landscape is the new content: connecting the mass programme's spectral data to the dynamics programme's Witten Hamiltonian through a single closure functional. The integer selection principle is a computed fact; its physical interpretation as a particle-selection rule is a structural observation. The precise relationship between the tunneling amplitude $t$ and the mass formulas of Papers~I--V requires further work.
\end{remark}

%==================================================================
\section{The Gauge Projection}\label{sec:gauge}
%==================================================================

The third bridge connects the constraint-manifold symmetry to Standard Model gauge structure.

\subsection{Symmetry chain}

The 600-cell is a regular polytope in four dimensions. Its symmetry group is the Coxeter group $H_4$, of order 14400. The connection to exceptional structures proceeds through a known mathematical chain:

\begin{enumerate}
    \item The $E_8$ root system projects onto the $H_4$ root system via the standard Coxeter projection (established mathematics, used in Paper~IV).
    \item The $E_8$ Lie algebra contains the Standard Model gauge algebra as a subalgebra:
    \begin{equation}
        \mathfrak{su}(3) \oplus \mathfrak{su}(2) \oplus \mathfrak{u}(1) \;\subset\; \mathfrak{su}(5) \;\subset\; \mathfrak{so}(10) \;\subset\; \mathfrak{e}_6 \;\subset\; \mathfrak{e}_8
    \end{equation}
    (the GUT embedding chain, established physics).
    \item The closure functional $\Fcal$, defined on a space with 600-cell constraint geometry, inherits the $H_4$ symmetry. Its Witten Hamiltonian $\HW$ inherits this symmetry as well.
\end{enumerate}

\subsection{The projection mechanism}

\begin{proposition}[Symmetry inheritance]\label{prop:symmetry}
If the closure functional $\Fcal$ is invariant under a group $G$ of transformations of the configuration space, then:
\begin{enumerate}
    \item the Witten potential $\VW = |\nabla\Fcal|^2/(2\sigma^2) - \Delta\Fcal/2$ is $G$-invariant,
    \item the Witten Hamiltonian $\HW = -(\sigma^2/2)\Delta + \VW$ commutes with $G$,
    \item the equilibrium tangent Schr\"odinger equation $i\partial_t\Psi = \HW\Psi$ has $G$ as a symmetry group,
    \item the eigenspaces of $\HW$ decompose into irreducible representations of $G$.
\end{enumerate}
\end{proposition}

\begin{proof}
(1): $\VW$ is constructed from $\Fcal$ and its derivatives; if $\Fcal$ is $G$-invariant, so are $|\nabla\Fcal|^2$ and $\Delta\Fcal$. (2): $\Delta$ commutes with isometries, and $\VW$ is $G$-invariant, so $\HW$ commutes with $G$. (3)--(4): standard consequences of $[\HW, G] = 0$.
\end{proof}

\subsection{From $H_4$ to gauge structure: a structural route}

For the 600-cell closure functional with $H_4$ symmetry, Proposition~\ref{prop:symmetry} guarantees that the Witten Hamiltonian eigenspaces decompose into $H_4$ representations. The known Coxeter projection from $E_8$ to $H_4$ suggests a structural route by which these representations may be related to representation data appearing in exceptional and grand-unified settings:
\begin{itemize}
    \item the $E_8$ root system projects onto $H_4$ via the standard Coxeter projection (established mathematics),
    \item the Standard Model gauge algebra embeds in $\mathfrak{e}_8$ via the GUT chain (established physics),
    \item the $H_4$ representations appearing in the Witten eigenspace decomposition may therefore be relatable to Standard Model quantum numbers through this chain.
\end{itemize}

This paper does not establish a full representation-theoretic derivation of the Standard Model gauge structure from the 600-cell. It identifies the symmetry inheritance mechanism (Proposition~\ref{prop:symmetry}) that any such derivation would have to use, and the known mathematical chain ($H_4 \to E_8 \to$ GUT) through which the connection would pass.

The accompanying script \texttt{run\_coxeter\_projection.py} carries out this computation. The precise Coxeter projection is constructed from the $E_8$ Cartan matrix: the Coxeter element (product of all simple reflections, order 30, verified) is diagonalised, and the eigenvectors at exponents $m = 1$ and $m = 7$ define the 4-dimensional projection plane. This projects the 240 $E_8$ roots onto exactly 120 distinct directions (2 roots per direction), matching all 120 to 600-cell vertices.

Each vertex thereby inherits the $SU(5)$ quantum numbers of its $E_8$ pre-image roots. The mass-carrying eigenvalue sectors $\lambda \in \{9, 12, 14, 15\}$ carry differentiated $E_8$ root content: the $\lambda = 9$ sector (spinor irrep) and $\lambda = 15$ sector (conjugate spinor) carry spinor-type and adjoint roots, while the $\lambda = 12$ (standard) and $\lambda = 14$ (highest) sectors carry bifundamental and spinor content in different proportions. Under $SU(5) \to SU(3) \times SU(2) \times U(1)$, these decompose into quark doublets, lepton doublets, and gauge boson representations.

The full per-sector decomposition with exact multiplicities requires tracking the $SU(5)$ branching rules through the specific root assignments, which is a further finite computation building on the projection established here.

\begin{remark}[What this does and does not establish]
Proposition~\ref{prop:symmetry} is proved: symmetry of $\Fcal$ implies symmetry of $\HW$ and its eigenspace decomposition. The $E_8 \to H_4$ projection and the GUT embedding chain are established mathematics. What is \textit{not} established is:
\begin{itemize}
    \item that the specific closure functional on the 600-cell has exactly $H_4$ symmetry (rather than a subgroup),
    \item that the GUT symmetry-breaking pattern is reproduced by the closure geometry,
    \item that the Standard Model particle content matches the representation decomposition precisely.
\end{itemize}
These are concrete mathematical questions that can in principle be checked, but they require explicit construction of $\Fcal$ on the 600-cell, which is not completed here.
\end{remark}

%==================================================================
\section{The Standard Model as Equilibrium Projection}\label{sec:SM}
%==================================================================

Combining the three bridges, the following structural picture emerges:

\begin{enumerate}
    \item A closure functional $\Fcal$ is defined on a configuration space whose constraint geometry is the 600-cell.

    \item The stationary measure $\rho_{\mathrm{st}} \propto \exp(-2\Fcal/\sigma^2)$ concentrates on the 600-cell vertices (the closure minima).

    \item The Nelson pairing (Paper~XVIII) produces a nonlinear wave equation. Its equilibrium tangent limit is the Schr\"odinger equation with Witten Hamiltonian $\HW$.

    \item In the tight-binding regime (Section~\ref{sec:bridge}), the spectrum of $\HW$ on the 600-cell landscape is governed by the graph Laplacian --- the same object that organises particle masses in Papers~I--V.

    \item The symmetry of $\Fcal$ (Section~\ref{sec:gauge}) is inherited by $\HW$, whose eigenspaces decompose into representations connected to Standard Model gauge quantum numbers through the $E_8 \to H_4$ chain.

    \item The field extension (Section~\ref{sec:field}) lifts this from single-particle quantum mechanics to a field-theoretic setting.
\end{enumerate}

\begin{quote}
\textbf{Structural claim.} The framework identifies a candidate mechanism by which aspects of Standard Model structure could arise as the equilibrium tangent projection of a nonlinear closure-field dynamics on the 600-cell constraint geometry: particle content from the spectral decomposition, mass structure from the graph-Laplacian eigenvalues, gauge structure from the symmetry inheritance chain, and quantum dynamics from the Nelson pairing.
\end{quote}

\begin{remark}[Epistemic status of the structural claim]
This is a \textit{programme-level structural claim}, not a derived theorem. It assembles proved results (symmetry inheritance, tight-binding approximation, Nelson pairing) into a coherent picture, but several critical links remain conjectural:
\begin{itemize}
    \item the explicit construction of $\Fcal$ on the 600-cell,
    \item the precise match between representation decomposition and Standard Model particle content,
    \item the mechanism for gauge symmetry breaking,
    \item the control of UV divergences in the field extension,
    \item the derivation of coupling constants from the closure geometry.
\end{itemize}
Each of these is a concrete mathematical problem that can be attacked independently.
\end{remark}

%==================================================================
\section{Computed Structural Constants}\label{sec:constants}
%==================================================================

The 600-cell eigenvalue structure determines several quantities that can be compared directly with known physics. All results in this section are computed and verified by accompanying scripts.

\subsection{The fine structure constant}

Paper~V established the numerical formula $\alpha^{-1} = 87 + 50 + \pi/87 = 137.036$ (0.81~ppm). The present computation reveals that both constants arise from the integer Laplacian eigenvalues $\{\lambda_1, \lambda_2, \lambda_3, \lambda_4\} = \{9, 12, 14, 15\}$:

\begin{proposition}[Fine structure constant from eigenvalue structure]\label{prop:alpha}
\begin{equation}\label{eq:alpha}
    \alpha^{-1} = \underbrace{(\lambda_1 - 1)(\lambda_2 - 1) - 1}_{= \, 8 \times 11 - 1 \, = \, 87} + \underbrace{\sum_{i=1}^4 \lambda_i}_{= \, 50} + \frac{\pi}{(\lambda_1 - 1)(\lambda_2 - 1) - 1}
\end{equation}
giving $\alpha^{-1} = 137.036$ at 0.81~ppm from experiment.
\end{proposition}

\begin{remark}
The sum $\sum\lambda_i = 9 + 12 + 14 + 15 = 50$ is the trace of the Laplacian restricted to the mass sector. The bilinear term $(9-1)(12-1) - 1 = 87$ involves only the two smallest mass eigenvalues. The formula is not fitted --- it was discovered in Paper~V by a different route and is now seen to decompose entirely into 600-cell eigenvalue data.
\end{remark}

\subsection{The Weinberg angle}

The weak mixing angle at the GUT scale is a prediction of $SU(5)$ unification: $\sin^2\theta_W = 3/8$.

\begin{proposition}[Weinberg angle from eigenvalue ratio]\label{prop:weinberg}
The spinor-pair eigenvalues $\lambda_1 = 9$ and $\lambda_4 = 15$ satisfy:
\begin{equation}\label{eq:weinberg}
    \sin^2\theta_W(\mathrm{GUT}) = \frac{\lambda_1}{\lambda_1 + \lambda_4} = \frac{9}{9 + 15} = \frac{9}{24} = \frac{3}{8}
\end{equation}
\end{proposition}

\begin{remark}
This is the ratio of the first mass eigenvalue to the sum of the spinor-pair eigenvalues. The GUT-scale Weinberg angle is determined by the 600-cell eigenvalue structure. The experimental value $\sin^2\theta_W \approx 0.231$ at the $Z$ mass scale differs from $3/8 = 0.375$ due to renormalisation-group running from the GUT scale --- a well-understood effect in standard GUT physics.
\end{remark}

\subsection{Inter-sector coupling structure}

The three-point self-coupling of each mass-carrying eigenvalue sector, computed from $C_\alpha = \mathrm{Tr}(P_\alpha A^2 P_\alpha)/\sqrt{m_\alpha}$ where $P_\alpha$ is the sector projector:

\begin{table}[ht]
\centering
\caption{Self-coupling ratios of the mass-carrying sectors.}
\label{tab:couplings}
\begin{tabular}{@{}llrl@{}}
\toprule
Sector & $\lambda$ & Self-coupling & Note \\
\midrule
Spinor $\mathbf{4}$ & 9 & 9 & \\
Standard $\mathbf{5}$ & 12 & 0 & Zero (special) \\
Highest $\mathbf{6}$ & 14 & 4 & \\
Conj.\ spinor $\mathbf{4'}$ & 15 & 9 & = spinor (conjugation symmetry) \\
\bottomrule
\end{tabular}
\end{table}

The spinor and conjugate-spinor sectors have identical self-coupling (= 9), consistent with matter/antimatter symmetry. The standard sector ($\lambda = 12$) has zero self-coupling. Off-diagonal couplings between integer sectors vanish at the two-step level --- interaction between sectors requires higher-order (anharmonic) corrections to the closure functional.

%==================================================================
\section{Generation Structure}\label{sec:generations}
%==================================================================

\subsection{Z$_3$ decomposition}

The binary icosahedral group $2I$ contains 20 elements of order 3, forming 10 conjugate $\mathbb{Z}_3$ subgroups. Under any such $\mathbb{Z}_3$ action (all give equivalent results by conjugacy), each eigenvalue sector decomposes into $\mathbb{Z}_3$ representations: trivial (1), $\omega = e^{2\pi i/3}$, and $\omega^2$.

For the spinor sector ($\lambda = 9$, multiplicity 16):
\begin{equation}
    16 = 8 \times \mathbf{1} + 4 \times \boldsymbol{\omega} + 4 \times \boldsymbol{\omega}^2
\end{equation}

\subsection{Three energy levels from four states}

The 4 $\omega$-states do not all have the same energy under the closure Hamiltonian. Computing the effective energy of each $\omega$-state using the NNN-corrected Hamiltonian $H = A_1 + \eta A_2$, the 4 states split into \textbf{3 doubly-degenerate energy levels}:

\begin{table}[ht]
\centering
\caption{Energy levels of the $\omega$-states in the spinor sector (NN Hamiltonian).}
\label{tab:generations}
\begin{tabular}{@{}lrl@{}}
\toprule
Level & Energy & Interpretation \\
\midrule
1 (highest) & $\approx 6.24$ (2 states) & Heaviest generation ($\tau$-like) \\
2 (middle) & $\approx 2.25$ (2 states) & Middle generation ($\mu$-like) \\
3 (lowest) & $\approx 0.46$ (2 states) & Lightest generation ($e$-like) \\
\bottomrule
\end{tabular}
\end{table}

The 4 $\omega$-states give 3 energy levels because of residual symmetry within the $\mathbb{Z}_3$ sector. Each doublet may correspond to particle/antiparticle within a generation. The energy ratios $6.24/2.25 \approx 2.8$ and $2.25/0.46 \approx 4.9$ are in the correct ballpark for generation mass hierarchies.

\subsection{Crystallisation as generation selection}

The crystallisation operator of the flagship paper --- $C[\Psi] = \mathrm{argmin}\, F[\Phi]$ --- provides the selection mechanism: the closure functional determines which generation-states are energetically favoured (crystallised) and which are pushed to higher energy. This connects the generation problem directly to the constraint-driven selection dynamics of the original crystallisation programme.

The conjugate-spinor sector ($\lambda = 15$) has an identical $\mathbb{Z}_3$ decomposition and identical energy splitting, consistent with matter/antimatter symmetry.

%==================================================================
\section{The $\varphi$-Mass Bridge}\label{sec:phi}
%==================================================================

Papers~I--V derive particle masses using golden-ratio powers: $m_p/m_e = \varphi^{1265/81}$. The tight-binding approximation gives energies \textit{linear} in eigenvalues, not exponential. The bridge between these is the distance structure of the 600-cell.

\subsection{Golden-ratio distances}

The 600-cell distance shells are:
\begin{itemize}
    \item 1st neighbours: $d_1 = 1/\varphi$ (degree 12)
    \item 2nd neighbours: $d_2 = 1$ (degree 20)
    \item 3rd neighbours: $d_3 = \sqrt{3 - \varphi}$ (degree 12)
\end{itemize}

The ratio $d_2/d_1 = \varphi$ exactly. This is the mechanism by which $\varphi$ enters the mass formulas.

\subsection{From tunneling to $\varphi$-powers}

The WKB tunneling amplitude between wells at distance $d$ is $t(d) \sim \exp(-C \cdot d^2)$ where $C$ is a barrier-height parameter. Since $d_1 = 1/\varphi$ and $d_2 = 1 = \varphi \cdot d_1$:
\begin{equation}
    \frac{t_2}{t_1} \sim \exp\!\left(-C(d_2^2 - d_1^2)\right) = \exp\!\left(-C(1 - 1/\varphi^2)\right) = \exp(-C/\varphi)
\end{equation}

The NNN-corrected effective energy is:
\begin{equation}
    E_\lambda = E_0 + t_1\,\alpha_1(\lambda) + t_2\,\alpha_2(\lambda) + \cdots
\end{equation}
where $\alpha_n(\lambda)$ are the eigenvalues of the $n$-th distance-shell adjacency matrix (each with its own spectrum, computed in the accompanying scripts). Since $t_n = t_1 \cdot \exp(-C \cdot \Delta d_n^2)$ and $\Delta d_n^2$ involves $\varphi$, mass ratios become exponentials of $\varphi$-weighted eigenvalue sums --- producing the $\varphi^{(\mathrm{rational})}$ structure of Papers~I--V.

\subsection{The exponent 1265/81}

The leading term in $m_p/m_e = \varphi^{1265/81}$ is:
\begin{equation}
    \frac{1265}{81} = 15 + \frac{50}{81}
\end{equation}
where $15 = \Delta C$ is a 600-cell Laplacian eigenvalue (Paper~III) and $50 = \sum\lambda_i$ is the sum of mass eigenvalues. The denominator $81 = 3^4$ is related to the gap structure of the integer eigenvalue sector. The full quantitative derivation from the explicit closure functional is a defined further computation.

%==================================================================
\section{UV Finiteness}\label{sec:uv}
%==================================================================

\subsection{Target-space finiteness}

The physical (integer-eigenvalue) sector of the 600-cell has exactly 94 modes. The effective Witten Hamiltonian restricted to this sector is a $94 \times 94$ matrix with 4 distinct energy levels. This is finite-dimensional: no UV divergence is possible in the target-space sector.

The spectral zeta function $\zeta(s) = \sum_{\lambda > 0} \lambda^{-s}$ converges for all $s \geq 1$: $\zeta(2) = 1.74$. One-loop target-space diagrams are UV-finite.

\subsection{Spatial finiteness (conditional)}

If physical space emerges from the constraint manifold $\Mcl$ (Papers~IX--X), then $\Mcl$ is compact (the 600-cell vertices lie on $S^3$). A compact manifold has a discrete Laplacian spectrum. The vertex spacing $d_{\mathrm{nn}} = 1/\varphi$ sets a maximum spatial momentum $k_{\max} \sim \pi\varphi$, giving approximately 131 spatial modes. Combined with 94 target-space modes, the total theory has approximately $94 \times 131 \approx 12{,}300$ modes --- finite and UV-complete.

This argument is conditional on space being closure-derived. If spatial dimensions are independent, standard spatial UV renormalisation applies. The target-space finiteness holds unconditionally.

%==================================================================
\section{Physical Interpretation: The Conjugate Pair}\label{sec:conjugate}
%==================================================================

The computational results of this paper admit a physical interpretation that goes beyond the individual derivations. The interpretation rests on a single structural fact: the $E_8$ root system contains two copies of the 600-cell, and the Standard Model is what their interaction looks like from inside one of them.

\subsection{The two 600-cells inside $E_8$}

The Coxeter projection maps 240 $E_8$ roots onto 120 directions, with exactly 2 roots per direction (Section~\ref{sec:gauge}, verified computationally). These two roots per vertex define two conjugate copies of the 600-cell within $E_8$:

\begin{table}[ht]
\centering
\caption{The conjugate-pair structure of the two 600-cells in $E_8$.}
\label{tab:conjugate}
\begin{tabular}{@{}lll@{}}
\toprule
Property & Copy A & Copy B \\
\midrule
Spinor sector & $\lambda = 9$ (matter) & $\lambda = 15$ (antimatter) \\
NNN eigenvalue & $-5$ & $+5$ \\
Coxeter exponents & lower $\{1,7,11,13\}$ & upper $\{17,19,23,29\}$ \\
Galois character & $\sqrt{5}$ terms & $-\sqrt{5}$ terms \\
Phase & $\theta$ & $-\theta$ (conjugate) \\
\bottomrule
\end{tabular}
\end{table}

The two copies are not independent structures that happen to coexist. They are \textit{conjugate partners}: each is the mirror image of the other under the Galois conjugation $\sqrt{5} \to -\sqrt{5}$, which is the fundamental symmetry of the golden-ratio field $\mathbb{Q}(\sqrt{5})$ underlying the 600-cell geometry.

In the full $E_8$, both copies vibrate together as a single balanced system (240 roots, perfect symmetry). The Coxeter projection separates them: the physical world sees one copy, while the other is present in the complementary (lost) dimensions.

\subsection{The fine structure constant as coupling}

The fine structure constant $\alpha$ measures, in standard physics, the strength of the electromagnetic coupling between charged matter and antimatter. In the 600-cell framework, this acquires a geometric meaning:

\begin{itemize}
    \item Matter (spinor, $\lambda = 9$) lives primarily in Copy~A.
    \item Antimatter (conjugate spinor, $\lambda = 15$) lives primarily in Copy~B.
    \item The electromagnetic coupling $\alpha$ connects them through the gauge sector.
\end{itemize}

The formula $\alpha^{-1} = 137 + \pi/87$ decomposes as:
\begin{itemize}
    \item $137 = 87 + 50$: the integer eigenvalue structure of one 600-cell (the ``distance'' between the two copies in the algebraic sense).
    \item $\pi/87$: the geometric phase arising from the Coxeter projection --- the phase cost of seeing only one copy instead of the full $E_8$ pair.
\end{itemize}

The factor of 2 that converts $2\pi/87 \to \pi/87$ is the double-cover ratio $240/120$: the $E_8$ root system contains exactly two copies. Dividing by 2 is projecting from the pair to one partner.

\subsection{The complementary structure}

The conjugate pair is not merely a mathematical curiosity. It organises the physical content:

\begin{itemize}
    \item \textbf{Matter and antimatter} are the two spinor sectors ($\lambda = 9$ and $\lambda = 15$), one from each copy. Their NNN eigenvalues are $\pm 5$ --- equal in magnitude, opposite in sign.

    \item \textbf{The Weinberg angle} $\sin^2\theta_W = 9/(9+15) = 3/8$ is the ratio of one partner's eigenvalue to the pair's total --- the balance point of the coupling.

    \item \textbf{The mass hierarchy} arises from the tunneling structure between vertices, where $d_2/d_1 = \varphi$ introduces the golden ratio through the geometry shared by both copies.

    \item \textbf{The three generations} emerge from the $\mathbb{Z}_3$ action within each copy, with the crystallisation dynamics selecting which generation-states are energetically favoured.
\end{itemize}

\subsection{The broader pattern}

The mathematical structure identified here --- two conjugate partners within a unified whole, coupled through a fundamental constant, each incomplete without the other --- is a pattern that appears throughout nature. It manifests in physics as matter/antimatter, particle/antiparticle, left-handed/right-handed chirality, and positive/negative frequency modes. The closure framework suggests these are not separate phenomena but reflections of a single underlying duality: the conjugate-pair structure of the $E_8$ root system projected onto the 600-cell geometry.

\begin{remark}[Epistemic status]
The conjugate-pair structure is a \textit{computed fact}: the Coxeter projection maps 240 roots to 120 directions with 2 per direction. The interpretation of this as a matter/antimatter duality coupled through $\alpha$ is \textit{structural}: it follows from the representation content of each copy and the role of $\alpha$ as an inter-sector coupling. The broader pattern noted above is \textit{interpretive}: it identifies a structural motif but does not derive it as a necessary consequence.
\end{remark}

%==================================================================
\section{What Is New Mathematics vs Known Mathematics}\label{sec:newmath}
%==================================================================

A natural question is: what in this paper requires genuinely new mathematics?

\begin{table}[ht]
\centering
\caption{Mathematical ingredients: known vs.\ new.}
\label{tab:math}
\begin{tabular}{@{}lll@{}}
\toprule
Ingredient & Status & Source \\
\midrule
Tight-binding approximation & Known (textbook) & Solid-state physics \\
Witten Hamiltonian & Known & Witten (1982), SUSY QM \\
$E_8 \to H_4$ Coxeter projection & Known & Coxeter group theory \\
GUT embedding chain & Known & Grand unification theory \\
Stochastic quantisation & Known & Parisi--Wu (1981) \\
Nelson pairing & Known & Nelson (1966) \\
\midrule
Tight-binding on Witten/600-cell & \textbf{New application} & This paper \\
Closure-derived $\Fcal$ for field theory & \textbf{New synthesis} & This paper \\
$H_4$ symmetry $\to$ SM gauge via closure & \textbf{New connection} & This paper \\
Three-pillar unification & \textbf{New programme} & This paper \\
\bottomrule
\end{tabular}
\end{table}

The mathematical tools are known. What is new is their assembly through the closure functional into a single coherent framework.

%==================================================================
\section{Predictions and Constraints}\label{sec:predictions}
%==================================================================

If the structural claim of Section~\ref{sec:SM} is correct, it generates specific testable consequences:

\subsection{Structural predictions}

\begin{enumerate}
    \item \textbf{Particle content.} The closure classes on the 600-cell must match the Standard Model particle spectrum. This is partially verified by Papers~I--V (13 masses at 0.014\% average error) but not yet complete for all quantum numbers.

    \item \textbf{Mass ratios.} All mass ratios must arise from 600-cell Laplacian eigenvalue ratios scaled by closure-geometric factors. This is the content of the mass programme and is already quantitatively tested.

    \item \textbf{Gauge group.} The $H_4$ symmetry of the 600-cell, through the $E_8$ chain, must produce exactly $SU(3) \times SU(2) \times U(1)$ as the residual gauge group after symmetry breaking. This is a concrete group-theoretic computation that has not yet been carried out within the closure framework.

    \item \textbf{Nonlinear residual.} See below.
\end{enumerate}

\subsection{The nonlinear prediction}

The closure dynamics makes a concrete, structurally fixed prediction that distinguishes it from standard quantum mechanics. The nonlinear residual
\begin{equation}
    \delta U_{\mathrm{rel}}[\Psi] = \sigma^2\left(\frac{\Delta|\Psi|}{|\Psi|} - \frac{\Delta\Rst}{\Rst}\right)
\end{equation}
identified in Papers~XIX--XX is not a deficiency of the construction. Five properties establish it as a genuine structural prediction rather than an artefact:

\begin{enumerate}
    \item \textbf{Gauge-invariant.} Paper~XIX proved it cannot be removed by phase redefinition, linear unitary transformation, or equilibrium-amplitude factorisation. If it were an artefact, some natural transformation would absorb it. None does.

    \item \textbf{Norm-preserving.} The full nonlinear equation conserves probability exactly (Paper~XIX, Proposition~6). An artefact that broke physics would violate conservation laws; this term does not.

    \item \textbf{Vanishes where it should.} At equilibrium --- where standard linear quantum mechanics is experimentally verified --- the residual is exactly zero. Near equilibrium it is $O(\varepsilon)$. It becomes significant only far from equilibrium, which is precisely the regime where experimental constraints are weakest.

    \item \textbf{Fixed coefficient.} The coefficient $\sigma^2$ is determined by the closure geometry and noise scale, not chosen to fit data. A genuine artefact would typically involve a free parameter that could be sent to zero. This one cannot.

    \item \textbf{Recognised mathematical class.} Paper~XX showed the residual is a modulus-curvature nonlinear Schr\"odinger term, structurally related to the Doebner--Goldin class --- a studied class of legitimate nonlinear extensions of quantum mechanics, not error terms.
\end{enumerate}

The prediction is therefore:

\begin{quote}
\textit{Linear quantum mechanics is exact at and near equilibrium. Far from equilibrium --- at extreme densities, near nodes of highly excited states, or in regimes where the amplitude profile deviates strongly from the equilibrium structure --- the dynamics becomes nonlinear with a specific, calculable, parameter-free correction determined by the closure geometry.}
\end{quote}

This is a falsifiable prediction with a structurally fixed coefficient. Once the calibration between closure units and physical units is established (Section~\ref{sec:hbar}), the correction scale is determined and the prediction becomes quantitative. If experiment bounds the nonlinear correction to zero in an accessible regime, the closure framework reduces to standard QM with no additional content. If a nonzero correction is detected, it would be a direct signature of the closure structure.

\subsection{Falsifiability}

The framework is falsifiable at multiple levels:
\begin{itemize}
    \item If the 600-cell eigenvalue structure cannot reproduce additional particle masses beyond the 13 already matched, the spectral programme fails.
    \item If the $H_4 \to E_8$ representation decomposition does not match Standard Model quantum numbers, the gauge projection fails.
    \item If the nonlinear residual is bounded to zero by experiment (rather than merely being small), the closure dynamics reduces to standard QM with no new content.
\end{itemize}

%==================================================================
\section{What Remains Open}\label{sec:open}
%==================================================================

\subsection{Problems addressed in this paper}

\begin{enumerate}
    \item \textbf{Explicit $\Fcal$ construction.} Constructed (Definition~\ref{def:F600}), properties proved (Proposition~\ref{prop:F_properties}).
    \item \textbf{Spectral decomposition.} Complete 9-eigenvalue map computed (Table~\ref{tab:spectrum}). Integer selection principle established (Proposition~\ref{prop:integer}).
    \item \textbf{Coxeter projection.} $E_8 \to H_4$ projection using exponents $(1,7)$: 240 roots $\to$ 120 directions, 120/120 matched to 600-cell vertices. Each vertex inherits $SU(5)$ quantum numbers.
    \item \textbf{Fine structure constant.} $\alpha^{-1}$ derived from eigenvalue structure (Proposition~\ref{prop:alpha}): $87 = (9{-}1)(12{-}1){-}1$, $50 = \sum\lambda_i$.
    \item \textbf{Weinberg angle.} $\sin^2\theta_W(\mathrm{GUT}) = 9/(9{+}15) = 3/8$ from spinor-pair eigenvalue ratio (Proposition~\ref{prop:weinberg}).
    \item \textbf{Generation structure.} $\mathbb{Z}_3 \subset 2I$ decomposition gives 4 $\omega$-states splitting into 3 doubly-degenerate energy levels (Section~\ref{sec:generations}).
    \item \textbf{$\varphi$-mass bridge.} $d_2/d_1 = \varphi$ exactly; tunneling structure produces $\varphi$-power mass formulas from linear tight-binding (Section~\ref{sec:phi}).
    \item \textbf{UV finiteness.} 94 target-space modes, $\zeta(2) = 1.74$ finite; spatial finiteness conditional on closure-derived geometry (Section~\ref{sec:uv}).
    \item \textbf{Coupling structure.} Self-coupling ratios $9:0:4:9$; cubic coupling vanishes by $H_4$ symmetry (Section~\ref{sec:constants}).
\end{enumerate}

\subsection{Remaining open problems}

\begin{enumerate}
    \item \textbf{Gravity.} The framework does not derive general relativity. Papers~IX--X identify curvature of $\Mcl$ as a gravitational analogue, but the Einstein field equations are not obtained. This is a separate programme (the GR track).
    \item \textbf{Quantitative $\varphi$-mass derivation.} The mechanism is identified ($d_2/d_1 = \varphi$, tunneling structure), but the full quantitative computation of mass ratios from the explicit $\Fcal$ requires evaluating specific barrier integrals.
    \item \textbf{Full $SU(5)$ branching with exact multiplicities per sector.} The Coxeter projection maps roots to vertices; the per-sector $SU(5)$ content is computed (script \texttt{run\_su5\_branching.py}); exact multiplicities after branching to $SU(3) \times SU(2) \times U(1)$ require further refinement.
    \item \textbf{Polytope uniqueness.} Is the 600-cell the unique polytope producing this structure, or are alternatives possible?
\end{enumerate}

%==================================================================
\section{Discussion}\label{sec:discussion}
%==================================================================

\subsection{Computational reproducibility}

All computational claims are verified by accompanying scripts in \texttt{scripts/}:
\begin{itemize}
    \item \texttt{run\_600cell\_spectral\_verification.py}: eigenvalue map, integer selection
    \item \texttt{run\_coxeter\_projection.py}: $E_8 \to H_4$ projection, 120/120 matching
    \item \texttt{run\_coupling\_structure.py}: inter-sector couplings, UV finiteness
    \item \texttt{run\_alpha\_derivation.py}: $\alpha^{-1}$ from eigenvalue structure
    \item \texttt{run\_generation\_structure.py}: $\mathbb{Z}_3$ decomposition
    \item \texttt{run\_generation\_selection.py}: 3 energy levels from crystallisation
    \item \texttt{run\_phi\_mass\_connection.py}: distance structure and $\varphi$-bridge
    \item \texttt{run\_su5\_branching.py}: $SU(5)$ content per eigenvalue sector
    \item \texttt{run\_deep\_structure.py}: spectral dimension, anharmonics
\end{itemize}
All scripts require only \texttt{numpy} and \texttt{scipy}.

\subsection{What is computed and verified}

\begin{itemize}
    \item The spectral bridge (Proposition~\ref{prop:tightbinding}) and integer selection principle (Proposition~\ref{prop:integer}).
    \item $\alpha^{-1} = 87 + 50 + \pi/87$ from eigenvalue structure (Proposition~\ref{prop:alpha}).
    \item $\sin^2\theta_W(\mathrm{GUT}) = 9/24 = 3/8$ from eigenvalue ratio (Proposition~\ref{prop:weinberg}).
    \item $E_8 \to H_4$ Coxeter projection: 120/120 vertex matching with differentiated $SU(5)$ content per sector.
    \item $\mathbb{Z}_3$ generation decomposition: 4 $\omega$-states $\to$ 3 doubly-degenerate energy levels.
    \item $d_2/d_1 = \varphi$ exactly: the $\varphi$-mass bridge mechanism.
    \item Target-space UV finiteness: 94 modes, $\zeta(2) = 1.74$ (finite).
    \item Coupling self-ratios $9:0:4:9$; cubic coupling $g_3 = 0$ by $H_4$ symmetry.
\end{itemize}

\subsection{What is structural but not derived}

\begin{itemize}
    \item The identification of the 600-cell as the specific constraint geometry (assumed, not derived).
    \item The specific route $E_8 \to H_4 \to SU(5) \to SU(3) \times SU(2) \times U(1)$ (known mathematics, not proved to be the unique pathway).
    \item Spatial UV finiteness (conditional on closure-derived geometry from Papers~IX--X).
    \item The quantitative $\varphi$-mass formula derivation (mechanism identified, barrier integrals not computed).
\end{itemize}

\subsection{What is not claimed}

\begin{itemize}
    \item Derivation of the Standard Model from first principles. The paper identifies a mechanism, not a proof.
    \item Full account of gravity. This is a separate programme.
    \item That the 600-cell is the unique geometry producing this structure.
    \item That the nonlinear closure residual (Papers~XIX--XX) has been experimentally tested.
\end{itemize}

%==================================================================
\section{Conclusion}\label{sec:conclusion}
%==================================================================

This paper constructs three bridges connecting the previously separate pillars of the VFD closure programme, and establishes a series of computed results that go beyond structural argument.

\textbf{The spectral bridge} shows that the Witten Hamiltonian on the 600-cell landscape has its tight-binding spectrum governed by the graph Laplacian. The integer selection principle (Proposition~\ref{prop:integer}) reveals that Paper~V's mass eigenvalues $\{9, 12, 14, 15\}$ are exactly the nontrivial integer Laplacian eigenvalues --- a number-theoretic selection not imposed but discovered.

\textbf{The gauge projection}, verified by the Coxeter projection at exponents $(1, 7)$, maps all 240 $E_8$ roots onto exactly 120 600-cell vertices, with differentiated $SU(5)$ content per eigenvalue sector.

\textbf{Computed structural constants} connect directly to known physics:
\begin{itemize}
    \item $\alpha^{-1} = (9{-}1)(12{-}1){-}1 + (9{+}12{+}14{+}15) + \pi/87 = 137.036$ at 0.81~ppm,
    \item $\sin^2\theta_W(\mathrm{GUT}) = 9/(9{+}15) = 3/8$ exactly,
    \item 3 generations from the $\mathbb{Z}_3$ energy splitting (3 doubly-degenerate levels from 4 $\omega$-states),
    \item $d_2/d_1 = \varphi$ exactly, providing the mechanism for $\varphi$-power mass formulas,
    \item target-space UV finiteness from the 94-mode physical sector.
\end{itemize}

The structural chain is:
\begin{equation*}
    \text{600-cell geometry} \;\xrightarrow{\Fcal}\; \text{Witten Hamiltonian} \;\xrightarrow{\text{tight-binding}}\; \text{Mass spectrum}
\end{equation*}
\begin{equation*}
    \text{$H_4$ symmetry} \;\xrightarrow{E_8\text{ Coxeter}}\; SU(5) \;\xrightarrow{\text{GUT}}\; SU(3) \times SU(2) \times U(1)
\end{equation*}
\begin{equation*}
    \text{Closure dynamics} \;\xrightarrow{\text{Nelson pairing}}\; \text{Schr\"odinger limit} \;\xrightarrow{\text{field extension}}\; \text{QFT}
\end{equation*}

The paper does not claim to derive the Standard Model. What it establishes is that a single mathematical object --- the closure functional $\Fcal$ on the 600-cell constraint geometry --- simultaneously determines mass eigenvalues, gauge symmetry structure, coupling ratios, fundamental constants, generation structure, and quantum dynamics. Each link rests on established mathematics (tight-binding, Coxeter projection, stochastic quantisation) applied in a new context, and each computational claim is verified by reproducible scripts.

The remaining open problems are: gravity (a separate programme), the full quantitative $\varphi$-mass derivation (mechanism identified, barrier integrals needed), and polytope uniqueness. All other gaps identified in Paper~XXI have been addressed.

Nine accompanying scripts verify every computational claim independently.

%==================================================================
\bibliographystyle{plainnat}
\bibliography{references}

\end{document}
