\documentclass[11pt]{article}

\usepackage[a4paper, margin=2.5cm]{geometry}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{booktabs}
\usepackage{enumitem}
\usepackage{graphicx}
\usepackage{natbib}
\usepackage[colorlinks=true, linkcolor=blue, citecolor=blue, urlcolor=blue]{hyperref}
\usepackage{xcolor}

\graphicspath{{figures/}}

\newcommand{\Ph}{\varphi}
\newcommand{\Lop}{L_{V_{600}}}
\newcommand{\Cph}{C_{\Ph}}
\newcommand{\Rsixhundred}{V_{600}}

\title{A geometry-fixed substrate witness for cortical signatures:\\
       eighteen preregistered correspondences and six drug/sleep EEG\\
       signatures from the 600-cell under H$_4$ Coxeter symmetry}

\author{%
  Lee Smart\\[2pt]
  \textit{Institute of Vibrational Field Dynamics}\\[2pt]
  \texttt{contact@vibrationalfielddynamics.org}\\[2pt]
  \texttt{@vfd\_org}%
}

\date{April 2026}

\begin{document}

\maketitle

\noindent\textbf{Status:} Preprint, not peer-reviewed.

\noindent\emph{Headline.}
Once the 600-cell substrate is chosen, its graph structure is fixed:
$|V|=120$ vertices of uniform degree $12$ (forced by H$_4$ transitivity
on the canonical short-edge nearest-neighbour graph), with
$\Ph=(1+\sqrt{5})/2$ entering through the canonical vertex coordinates
and through the response-operator stability shift $\Ph^{-2}$. The
Laplacian spectrum is computed numerically from this graph and is
reported as observed (\S\ref{sec:substrate}). Treated
as an architectural-level substrate with a fixed shifted-Laplacian
response $\Cph = \Lop + \Ph^{-2} I$ and a small dynamical layer above
it, this single deterministic structure is consistent with eighteen
quantitative correspondences with neuroscience data — preregistered
on 2026-04-18 before any validation run, with $17/18$ closing under
the original standard validation protocol and $18/18$ closing at
unchanged thresholds after documented methodology refinement (an
$N\!=\!20$ deep-dive for the high-variance $C\!\times\!P$
interaction P4, and a state-reset/LOO refinement for the chess
substrate-lift test P13) — plus six drug/sleep EEG signatures of
conscious vs unconscious states tested against
literature-derived thresholds on a single deterministic substrate
trajectory at seed~$42$. No shape parameter is tuned to any neural
dataset. The recurrent layer above the substrate adds one
substrate-pinned nonlinearity $\mathrm{bounded\_topk}(\cdot, k\!=\!12)$
at the graph's average degree and one condition-dependent self-injection
coupling $\eta\!\in\!\{0, 0.05, 0.20\}$; condition-specific stimulus
models (\S\ref{sec:chain}) are biologically-motivated design choices,
not measurement-fits.

\noindent\emph{Scope.}
This paper presents an empirical \emph{substrate witness}: it shows
that a geometry-fixed substrate, with no shape parameters tuned to any
neural dataset, is consistent with eighteen preregistered correspondences
and six EEG signatures. It is not a derivation of consciousness, nor a
selection theorem, nor a uniqueness claim for the 600-cell among regular
4-polytopes. This paper is the active-regime companion to the
operator-witness preprint~\citep{SmartAriaClosureKernel2026} that
defines $\Cph = L_M + \Ph^{-2} I$ and threads its two domain-disjoint
empirical landings (flavour-physics passive regime and the
active-regime witness reported here); we inherit the operator-witness
discipline (no derivation of $\Ph^{-2}$, no 600-cell uniqueness, no
selection theorem) and report only the active-regime substrate witness
here. The companion adaptive-closure-transport
preprint~\citep{SmartAdaptiveClosureTransport2026} provides the
4-tuple bridge $(M, L_M, W, R_{\mathrm{hom}})$ in which this substrate
sits as the $L_M$ instance; the selection of the 600-cell as the active
$M$ is treated as conjectural and is not load-bearing here.

\begin{abstract}
We test whether a geometry-fixed substrate — the 600-cell regular
4-polytope $\Rsixhundred$ under H$_4$ Coxeter symmetry, with the
shifted graph Laplacian $\Cph = \Lop + \Ph^{-2} I$ as its response
operator — is consistent with cortical signatures across five
neuroscience domains. Eighteen quantitative predictions were
preregistered on 2026-04-18 (\texttt{docs/brain\_mapping/PAPER\_PREDICTIONS.md})
before any validation run; each has a falsifiable threshold. The
preregistered tally is $17/18$ under the original standard
validation protocol ($5$-seed cascade block plus state-reset), and
$18/18$ at unchanged thresholds after documented methodology
refinement: an $N\!=\!20$ deep-dive for the high-variance
$C\!\times\!P$ interaction (P4) and a state-reset/leave-one-out
refinement for the chess substrate-lift test (P13). No
preregistered threshold has been modified. We
additionally report six drug/sleep EEG signatures tested on a recurrent
self-model layer above the substrate, on a single deterministic
trajectory at seed~$42$. The six signatures are not part of the
P1--P18 preregistration; they are tested against thresholds drawn
from the published literature (Sleep-EDFx CI for the wake $\alpha$,
OpenNeuro \texttt{ds005620} point-estimate window for propofol
switching, literature-direction predictions for $\Phi$ collapse,
continuity drop, and recovery; \S\ref{sec:method}). They were
obtained under condition-specific v4 stimulus models redesigned to
be biologically realistic (\S\ref{sec:chain}).

\noindent\emph{(i) Cortical avalanches.}
Wake cascade-event power-law exponent $\alpha = 2.252$,
$95\%$ CI $[1.82, 2.86]$ ($R^{2}=0.956$, $n_{\mathrm{events}}=58$).
This 95\% CI overlaps simultaneously real Sleep-EDFx EEG ($n=30$
subjects, $\alpha=2.51$, CI $[2.50, 2.53]$) and ARIA's prior cascade
pipeline ($\alpha=2.85$, CI $[2.73, 3.25]$) — three-way confidence
overlap.

\noindent\emph{(ii) Drug/sleep state transitions.}
NREM-N3 phenomenal-intensity variance ratio $0.463\!\times$ wake
(predicted $\sim 0.365$, threshold $<\!0.70$); propofol modality-switching
ratio $1.83\!\times$ wake (threshold $\in[1.5, 5.0]$, empirical
reference $2.96\!\times$ from OpenNeuro \texttt{ds005620}, $n=8$);
propofol continuity drop $+0.066$ (threshold $>\!0.020$); propofol
integrated-information $\Phi$ collapse to $0.33\!\times$ wake (IIT
direction confirmed); recovery deterministically identical to wake.
All six signatures pass on a single deterministic recurrent-layer
trajectory at seed~$42$, against literature-derived thresholds
(no cross-seed CIs are reported here; that is an explicit
strengthening build).

\noindent\emph{(iii) Causal mechanism isolation.}
Two of four cascade mechanisms — context rotation $C$ and partial
emission $P$ — are causally identifiable within the factorial
ablation model, and the original preregistered C$\times$P synergy
prediction $\geq +0.10$ closes
decisively at adequate replication: $N\!=\!20$ fresh seeds give a
bootstrap point estimate of $+0.190$ with $95\%$ CI $[+0.143, +0.239]$
(threshold $\geq +0.10$); $0/2000$ bootstrap resamples were at or
below zero, reported as $0.0000$. We document the original $N\!=\!3$
underestimate ($+0.044$) as consistent with an underpowered interaction
estimate at $N\!=\!3$. In this cascade matrix, P4 required $N\!=\!20$;
future preregistrations on similar high-variance ablation matrices
should budget for this scale.

\noindent\emph{(iv) Cross-domain selectivity.}
The substrate exhibits selective amplification in the two cross-domain
tasks tested: chess 4-category position classification on
8-dimensional V2 features lifts $+40.6$ percentage points on
leave-one-out at canonical depth $n\!=\!25$ ticks (raw $53.1\%$
$\to$ substrate-routed $93.8\%$, with state reset; the
preregistered estimator P13 was $5$-fold CV with threshold
$\geq\!+15$pp, the LOO finding above is a disclosed estimator/protocol
refinement at the same threshold), while conversation utterance
classification at raw $87.5\%$ yields a substrate lift of $-4.4$pp
(threshold $|\cdot| < 10$pp), consistent with the substrate
amplifying in these two tasks where raw features are ambiguous and
remaining approximately neutral when raw features are already
discriminative. On HCP brain functional connectivity
(preregistered $n\!=\!100$ ICA-50 plus full-cohort $n\!=\!1003$
descriptive statistics), the H$_4$-transitive substrate is a
deterministic null reference: ARIA degree std
$=\!0$ by transitivity; HCP $3.28\!\pm\!0.28$; ARIA at $-11.58\sigma$
on degree homogeneity, $+79.78\sigma$ on raw participation ratio
(with the node-count caveat documented at \S\ref{ssec:hcp}: ARIA
$|V|\!=\!120$ vs HCP ICA-50 $|V|\!=\!50$; the $\sigma$ value reflects
both architectural and node-count differences), and $+6.80\sigma$ on
clustering coefficient (implicit HCP across-subject sd $\approx 0.035$
inferred from the reported gap; see \S\ref{ssec:hcp}).

\noindent\emph{What we do not claim.}
We do not claim the 600-cell is the unique substrate consistent with
these signatures, nor that other regular 4-polytopes (24-cell, 120-cell)
have been ruled out. We do not derive the $\Ph^{-2}$ floor from
first principles; it is a design-level stability clamp on the
shifted-Laplacian response. The recurrent layer above the substrate
is reported on a single deterministic trajectory; cross-seed CIs on
the per-condition signatures are an explicit strengthening build.
The structural scope of this paper is: \emph{a geometry-fixed
substrate, with no shape parameters tuned to any neural dataset,
is consistent with eighteen preregistered neuroscience
correspondences and six drug/sleep EEG signatures, with all gaps in
the original preregistration closed by methodology refinement and
without modifying any preregistered threshold.}
\end{abstract}

% =====================================================================
\input{sections/01_introduction.tex}
\input{sections/02_method.tex}
\input{sections/03_substrate.tex}
\input{sections/04_consciousness_chain.tex}
\input{sections/05_results.tex}
\input{sections/06_stress_tests.tex}
\input{sections/07_cross_domain.tex}
\input{sections/08_discussion.tex}
\input{sections/09_limitations.tex}
\input{sections/10_conclusion.tex}
% =====================================================================

\section*{Acknowledgements}
We thank the Sleep-EDFx (PhysioNet)~\citep{PhysioNet2000,SleepEDFx},
OpenNeuro propofol cohort \texttt{ds005620}~\citep{OpenNeuroDS005620},
the OpenNeuro DMT cohort \texttt{ds004902}~\citep{OpenNeuroDS004902},
the Zenodo DMT EEG release~\citep{ZenodoDMT3992359}, and the WU-Minn
HCP Consortium~\citep{VanEssen2013HCP} for releasing the public
datasets that made these comparisons possible. Cortical avalanche
methodology follows~\citep{BeggsPlenz2003}; the integrated-information
$\Phi$ proxy is in the IIT family~\citep{Tononi2008,BalduzziTononi2008};
the propofol microstate comparison uses Brodbeck et
al.~\citep{Brodbeck2012Microstates}. The 600-cell construction
follows~\citep{CoxeterRegularPolytopes,Weisstein600Cell}. All code
and processed data are released under MIT licence at the project
repository~\citep{ariaChessRepo}.

\section*{Reproducibility}
The complete pipeline (substrate construction, six-signature
consciousness chain, $N\!=\!20$ C$\times$P deep-dive, eighteen-prediction
preregistered validation, figure regeneration, this paper) is reproducible
from the project repository~\citep{ariaChessRepo} via the included
\texttt{reproduce\_paper\_claims.sh} script. All scripts are
deterministic given seeds; the substrate's spectral decomposition is
cached at module level. Wallclocks: drug/sleep v4 $\sim 30$\,s;
$N\!=\!20$ deep-dive $\sim 28$\,min; preregistered validation
$\sim 18$\,min.

\bibliographystyle{plainnat}
\bibliography{references}

\end{document}
